### 2024-05-30 noam_data_exploration_v2.Rmd
Add files via upload

---
title: "noam_data_exploration_v1"
author: "Noam Altman-Kurosaki"
date: "2024-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
```

```{r }
data <- read.csv(here::here("data", "algaeonly_cleaned.csv"), stringsAsFactors = F) # replacements don't work when strings are factors

str(data)

#pivot long

long_data <- pivot_longer(
  data = data,
  cols = 8:77,
  names_to = "Taxa",
  values_to = "Percent_cover")

transect_level <- long_data %>%
  dplyr::group_by(Year, Site, Habitat, Transect, Taxa) %>%
  summarise(Mean_cover = mean(Percent_cover), 
            SE = sd(Percent_cover)/sqrt(n()))

site_level <- transect_level %>%
  dplyr::group_by(Year, Site, Habitat, Taxa) %>%
  summarise(Mean_site_cover = mean(Mean_cover), 
            Site_se = sd(Mean_cover)/sqrt(n()))
```

```{r diversity indicies, fig.height = 15, fig.width=20}
meta_cols <- colnames(data[1:7]) 
species_cols <- colnames(data[8:77])

wide_transect_level <- data %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat, Transect) %>%
  summarize(across(all_of(species_cols), mean, na.rm = TRUE))

alpha_diversity <-
  data.frame(
    Year = as.factor(wide_transect_level$Year),
    Site = as.factor(wide_transect_level$Site),
    Habitat = as.factor(wide_transect_level$Habitat),
    Site_habitat = as.factor(wide_transect_level$Site_habitat),
    Transect = as.factor(wide_transect_level$Transect),
    Site_habitat_transect = as.factor(paste(wide_transect_level$Site_habitat, wide_transect_level$Transect)),
    Shannon = diversity(wide_transect_level[5:74], "shannon"),
    Richness = rowSums(wide_transect_level[5:74] > 0)
  )

alpha_diversity_means <- alpha_diversity %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()))

site_colours <- c("LTER 1" = "cornflowerblue",
               "LTER 2" = "darkblue",
               "LTER 3" = "yellowgreen",
               "LTER 4" = "forestgreen",
               "LTER 5" = "mediumorchid",
               "LTER 6" = "orange")
habitat_linetypes <- c("Fringing" = "solid",
                       "Backreef" = "longdash",
                       "Outer 10" = "twodash",
                       "Outer 17" = "dotted")
habitat_shapes <- c("Fringing" = 16,
                     "Backreef" = 17,
                     "Outer 10" = 1,
                     "Outer 17" = 2)

(richness_plot <-
  ggplot(data = alpha_diversity_means) +
  geom_point(aes(x = Year, y = Richness_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
  geom_line(aes(x = Year, y = Richness_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat)+
  labs(title = "a. Richness\n", y = "Richness\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

(shannon_plot <-
  ggplot(data = alpha_diversity_means) +
  geom_point(aes(x = Year, y = Shannon_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
  geom_line(aes(x = Year, y = Shannon_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat)+
  labs(title = "b. Shannon\n", y = "shannon\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

```

```{r alpha diversity panel graph, fig.height = 30, fig.width= 20}

ggarrange(richness_plot + rremove("x.text"), shannon_plot, common.legend = T,
          nrow = 2)

# seems like patterns in alpha diversity different depending on if you look at richness or shannon diversity 
```

```{r basic alpha diversity stats}
# these violate all sorts of assumptions I'm just messing around
richness_mod <- glmmTMB(Richness ~ Year*Site*Habitat + (1|Site_habitat_transect), data = alpha_diversity, family = poisson())
Anova(richness_mod)

# shannon_mod <- glmmTMB(Shannon ~ Year*Site*Habitat + (1|Site_habitat_transect), data = alpha_diversity, family = Gamma(link = "log")) # 0 values fuck this up. Return to this when not tired. 
```

```{r nmds, fig.height = 15, fig.width = 15}
species_matrix <- as.matrix(wide_transect_level[6:74])
rownames(species_matrix) <- paste(wide_transect_level$Year, wide_transect_level$Site_habitat, wide_transect_level$Transect)
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
wide_transect_level_subset <- wide_transect_level[!rows_to_remove,] # make sure rows are also removed from dataframe
wide_transect_level_subset$Stratum <- paste(wide_transect_level_subset$Year, "-", wide_transect_level_subset$Site, "-", wide_transect_level_subset$Habitat )

# identifying species w/ > 10% cover for plotting later
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 10))
abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
abundant_species_names <- colnames(abundant_species_matrix)

# WARNING - Before you run this, know that it will take a long time
set.seed(1032) 
# ord2 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 2) # looks like it automatically wisconsin transforms now?
# ord2 # stress = 0.22
# stressplot(ord2)

set.seed(1032)
ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
ord3 # 0.166
stressplot(ord3)

# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation

# plot(pro, cex=1.5)

site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
# add back in metadata
site.scores$Year <- as.factor(wide_transect_level_subset$Year)
site.scores$Site <- as.factor(wide_transect_level_subset$Site)
site.scores$Habitat <- as.factor(wide_transect_level_subset$Habitat)
site.scores$Site_habitat <- as.factor(wide_transect_level_subset$Site_habitat)
site.scores$Transect <- as.factor(wide_transect_level_subset$Transect)
site.scores$Stratum <- as.factor(wide_transect_level_subset$Stratum)

cent <- aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2

# Need to split stratum up for plotting later
cent<- 
  cent %>%
    separate(Stratum, into = c("Year", "Site", "Habitat"), sep = "-")
cent$Site_habitat <- paste(cent$Site, cent$Habitat)
# Add back in stratum column
cent$Stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean)$Stratum
# A lot of extra spaces

cent <- cent %>%
  mutate(across(everything(), str_trim))
cent$Year <- as.factor(cent$Year)
cent$Habitat <- as.factor(cent$Habitat)
cent$Site <- as.factor(cent$Site)
cent$Site_habitat <- as.factor(cent$Site_habitat)
cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
cent$NMDS2 <- as.numeric(cent$NMDS2)
species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
species.scores$Species <- rownames(species.scores)

# subset only the species w/ > 10% cover
abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]

ggplot() + 
  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = Habitat, colour = Site),
             alpha = 0.25, size=2) + # add the point markers
  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels

  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = Site_habitat, colour = Site),
            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
  facet_wrap(~Habitat) +
  scale_linetype_manual(values = habitat_linetypes) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        strip.text.x = element_text(size = 25),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(1.5)), # increase legend text
        legend.title = element_text(size = rel(1.5)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) + # add box around legend
  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
                                     title.hjust =0.5, nrow = 2),
         shape=guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5,
                            nrow = 2),
         path = guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5)
         )

# looks like stochasticity increases from outer 17 --> Fringing
# Even in lagoon though it looks like communities return to initial state... except for North Shore LTER sites
# ways to test stochasticity vs. determinism?
# seems like deterministic trajectories, although we're not starting from a clean slate
# accounting for founder effects?

# Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
bray_matrix <- vegdist(species_matrix, method = "bray")
year_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Year)
year_beta_disper #

# plot of dispersion test
plot(year_beta_disper, sub = "") # lol
boxplot(year_beta_disper, main = "Year", xlab = "")

site_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site)
plot(site_beta_disper, sub = "") # site five stands out
boxplot(site_beta_disper, main = "Site", xlab = "")

habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Habitat)
plot(habitat_beta_disper, sub = "") 
boxplot(habitat_beta_disper, main = "Habitat", xlab = "") # greatest variance in fringing it seems (as expected)

site_habitat_beta_disper <- vegan::betadisper(bray_matrix, wide_transect_level_subset$Site_habitat)
plot(site_habitat_beta_disper, sub = "") 
boxplot(site_habitat_beta_disper, main = "Site_habitat", xlab = "")

```

```{r CV, fig.height = 8, fig.width = 10}
# coefficient of variation
CV<-function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}

alpha_diversity$Cover <- rowSums(wide_transect_level[5:74])/100

diversity_stability <- alpha_diversity %>%
  dplyr::group_by(Site, Habitat, Site_habitat, Transect) %>%
  summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Richness_cv = CV(Richness),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()),
            Shannon_cv = CV(Shannon),
            Cover_mean = mean(Cover),
            Cover_se = sd(Cover)/sqrt(n()),
            Cover_cv = CV(Cover))

diversity_mod_1 <- glmmTMB(Cover_cv ~ Richness_mean, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_1)
plot(allEffects(diversity_mod_1)) # richness reduces cv (i.e. enhances stability!)

diversity_mod_2 <- glmmTMB(Cover_cv ~ Shannon_mean, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_2) # similar results

diversity_mod_3 <- glmmTMB(Cover_cv ~ Habitat*Site, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_3) # christ 
Anova(diversity_mod_3) # all highly sig but what the fuck am I doing here

diversity_mod_4 <- glmmTMB(Cover_cv ~ Habitat, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_4)
Anova(diversity_mod_4)
plot(allEffects(diversity_mod_4)) # greatest variance in the lagoon (makes sense given beta diversity trends above)

diversity_mod_5 <- glmmTMB(Cover_cv ~ Site, family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_5)
Anova(diversity_mod_5)

diversity_mod_6 <- glmmTMB(Cover_cv ~ Shannon_mean*Habitat + (1|Site), family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_6) # negligible variance in random effects
Anova(diversity_mod_6) # Diversity the strongest predictor of "stability" by far!

diversity_mod_7 <- glmmTMB(Cover_cv ~ Shannon_mean*Site + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_mod_7)
Anova(diversity_mod_7)

diversity_mod_universal <- glmmTMB(Cover_cv ~ Shannon_mean*Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
diversity_dredge <- dredge(diversity_mod_universal)
model.sel(diversity_dredge)

# best model based on dredge:
diversity_best <-  glmmTMB(Cover_cv ~ Shannon_mean*Site + Site*Habitat + (1|Site_habitat), family = Gamma(link = "log"), data = diversity_stability)
summary(diversity_best) # very low variance in random effect, which is great
Anova(diversity_best) # Interactions all significant
plot(dp <- simulateResiduals(diversity_best)) # some outliers, but otherwise good model fit. I am fine with this.

(richness_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Richness_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(shannon_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# maybe some heteroskedasticity?
# just eyeballing but it looks like "stability" is highest in LTER 2 Fringe, which also has low diversity
# Functional form vs. portfolio effects?
# Either way looks sick as shit. Great clustering by habitat within site, and diversity trend is strong in LTER 3 4 5
# LTER2 the trend is reversed
# LTER 1 is mostly driven by habitat
# LTER 6 potentially a wash

# GOALS: test portfolio effects, look at turnover and temporal beta diversity stats, ways to test beta dispersion

```

```{r variance vs mean diversity plot, fig.height = 8, fig.width = 10}
(diversity_mean_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
# More diverse regions are also those that fluctuate the least in their diversity
```

```{r general cover diversity, fig.height = 30, fig.width = 30}
(diversity_cover_plot = 
  ggplot(data = alpha_diversity) +
  geom_point(aes(x = Cover, y = Shannon, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Year) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         ) # ABORT ABORT ABORT
```