### 2024-07-02 mcr_algal_data_cleaning_final.Rmd
use this cleaning script with the raw data

---
title: "MCR_Algae_Data_Cleaning"
author: "Noam Altman-Kurosaki and Lauren Enright"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls()) # clean up
library(knitr)
library(dplyr)
library(plyr)
library(car)
library(reshape)
library(reshape2)
library(tidyr)
library(effects)
library(devtools)
opts_chunk$set(comment="  ",
               collapse=TRUE, 
               echo=FALSE,
               #fig.asp=1/gr,
               fig.height=8,
               fig.width = 10,
               dev="png",
               warning=TRUE
               )
opts_knit$set(eval.after = "fig.cap")

```

```{r data cleaning}
library(here)

data <- read.csv(here("data", "MCR_LTER_Annual_Survey_Benthic_Cover_20231211.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
str(data)
levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 86 levels
data$Site_habitat <- as.factor( paste(data$Site, data$Habitat))

#### REPLACE OUTDATED TAXA

# Caulerpa peltata is outdated --> Caulerpa racemosa
data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Caulerpa peltata"] <- "Caulerpa racemosa"

# Boodlea kaeneana --> Cladophoropsis membranacea
data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Boodlea kaeneana"] <- 
  "Cladophoropsis membranacea"

# Cladophoropsis luxurians --> Cladophora fuliginosa
data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Cladophoropsis luxurians"] <-
  "Cladophora fuliginosa"

# Neogoniolithon frutescens --> Neogoniolithon brassica-florida
data$Taxonomy_Substrate_Functional_Group[data$Taxonomy_Substrate_Functional_Group == "Neogoniolithon frutescens"] <- 
  "Neogoniolithon brassica-florida"

# remove old levels
data <- droplevels(data)
levels(as.factor(data$Taxonomy_Substrate_Functional_Group)) # 84 levels

```

```{r QA QC - checking quadrats}

quad_level_summary <- data %>%
  dplyr::group_by(Year, Location, Site, Habitat, Site_habitat, Transect, Quadrat) %>%
  dplyr::summarize(sum = sum(Percent_Cover))

non100_quad_level <- quad_level_summary %>%
  dplyr::filter(sum != 100)

data %>% 
  filter(Taxonomy_Substrate_Functional_Group == "No data")

#99 observations that do not equal 100

#how many are no data (-1)
quad_level_summary %>%
  dplyr::filter(sum == -1)

#14

mistakes_quad_level <- quad_level_summary %>%
  dplyr::filter(sum != 100) %>%
  dplyr::filter(sum != -1)

#remaining 85 have something weird happening! Will send to Bob for updates. For now, 
# will filter these out

unique(quad_level_summary$sum)

#write out these 85

#write.csv(mistakes_quad_level, "mistakes_quad_level.csv", row.names = FALSE)
```

```{r subsetting the data }

### CREATE DATAFRAME THAT IS JUST ALGAE

not_algae <- c("Ascidian", "Bare Space", "Coral", "Coral Rubble", "Corallimorpharia", "Heteractis sp.", "Millepora platyphylla", "No data", "Sand", "Sacrophyton sp.", "Shell Debris", "Soft Coral", "Sponge", "Tridacna sp.", "Zooxanthellae")

### CREATE DATAFRAME THAT IS JUST MACROALGAE
# Maintaining mat-forming cyanos even though they're not macroalgae in the traditional sense
# "Coelothrix irregularis", "Symploca hydnoides"

not_macro <- c("Algal Turf", "Crustose Corallines", "Cyanophyta", "Damselfish Turf", "Lithophyllum kotschyanum", "Neogoniolithon brassica-florida", "Sporolithon sp." )

```

```{r wide - with all data}

comb_wide <- cast(data, Year + Location + Site + Habitat + Site_habitat + Transect + Quadrat ~ Taxonomy_Substrate_Functional_Group, fun.aggregate = mean, value = "Percent_Cover") # this is the correct number of rows given the missing data

#21000 obs of 91 variables

#change NAs to 0s

is.nan.data.frame <- function(x){
  do.call(cbind, lapply(x, is.nan))}

comb_wide[is.nan.data.frame(comb_wide)] <- 0

#21000 obs of 91 variables

```

```{r QAQC for the wide dataset}

comb_wide_quad_sum <- comb_wide %>%
  mutate(row_sums = rowSums(dplyr::select(., 8:91)))

#do we have data for all transects?

comb_wide_quad_sum %>%
  dplyr::group_by(Year, Site) %>%
  dplyr::summarise(count = n())

#should be 200 per year per site (50 at each habitat * 4 habitats)
#2005 and 2006 are one large collective survey
#2021 sites are at 100 (did not do forereef habitats due to COVID)

#do all quads = 100 ?

non100_wide_quad_level <- comb_wide_quad_sum %>%
  dplyr::filter(row_sums != 100)

#also 99 observations that do not equal 100, which matches what we found in the long data

#how many are no data (-1)
comb_wide_quad_sum %>%
  dplyr::filter(row_sums == -1)

#14

mistakes_wide_quad_level <- comb_wide_quad_sum %>%
  dplyr::filter(row_sums != 100) %>%
  dplyr::filter(row_sums != -1)

comb_wide_quad_sum %>%
 dplyr::filter(row_sums != 100) %>% dplyr::filter(Year == 2020)

#remaining 85 have something weird happening! Will send to Bob for updates. For now, 
# will filter these out

#want to make sure these 85 are the same we found using the long data

setdiff(mistakes_wide_quad_level$Location, mistakes_quad_level$Location)

setdiff(mistakes_quad_level$Location, mistakes_wide_quad_level$Location)

#yay, they are the same 85

```

```{r dealing with the 99s in the wide dataset}
#LE talked to Hillary and we found that of the 85 rows that have sums that do not = 100, 49 of them are because a "-1" was accidentally added to the no data column, even though data was taken. (100 percent cover - 1 = 99). This happened only in 2017. Hillary will fix this eventually on EDI, but for now I am going to manually remove these -1 from the data set so we can continue with our analyses

#show this error in the data

no_data_mistakes <- mistakes_wide_quad_level %>%
  filter(row_sums == 99)

unique(no_data_mistakes$`No data`)

#remove the -1 from the No data column

comb_wide_quad_sum$`No data`[comb_wide_quad_sum$row_sums == 99 & comb_wide_quad_sum$`No data` == -1] <- 0

#check to see if it worked

cleaned_sumrows99 <- comb_wide_quad_sum %>%
  filter(row_sums == 99)

unique(cleaned_sumrows99$`No data`)

#yay it works

```

#of our 85 quads that do not equal 100, 49 have been accounted for. 23 of them are from 2020, either Backreef or Fringing, which we know were averaged. This leaves only 13 quads with questions remaining.

```{r make new row sums column}
#need to update row sum now that fixes have been made
comb_wide_quad_sum_final <- comb_wide_quad_sum %>%
  dplyr::rename(row_sum_old = row_sums) %>%
  mutate(row_sums = rowSums(dplyr::select(., 8:91))) %>%
  select(-row_sum_old)

#confirm 99 is gone 
unique(comb_wide_quad_sum_final$row_sums)

#whooo

```

```{r filter the wide dataset}

#we want to filter out 2005 and 2006 - these were "ramp up years"
#also want to filter out rows where the sum does not = 0 
wide_alldata_filtered <- comb_wide_quad_sum_final %>%
  dplyr::filter(Year != 2005 & Year != 2006) %>%
  dplyr::filter(row_sums == 100)

#should be 19750 obs

#writing one file so column names will work with Noam's code
write.csv(wide_alldata_filtered, "allbenthicdata_cleaned.csv", row.names = FALSE)

#clean names so column names are not so hard to work with

wide_alldata_filtered_cleaned_column_names <- 
  wide_alldata_filtered %>%
  janitor::clean_names()

write.csv(wide_alldata_filtered_cleaned_column_names, "allbenthicdata_cleaned_colnames.csv", row.names = FALSE)

#we also want a copy that only includes taxa we care about 
#not_algae character string created by Noam in Chunk 4 (Subsetting the data) and includes 14 
#categories  -- these categories will be listed in a separate document as well

wide_algaeonly_filtered <- wide_alldata_filtered[ , -which(names(wide_alldata_filtered) 
                                                           %in% not_algae)]

#need to still remove the row_sums column for algae only

wide_algaeonly_filtered <- wide_algaeonly_filtered %>%
  dplyr::select(-row_sums)

write.csv(wide_algaeonly_filtered, "algaeonly_cleaned.csv", row.names = FALSE)

#clean names so column names are not so hard to work with

wide_algaeonly_cleaned_column_names <- 
  wide_algaeonly_filtered %>%
  janitor::clean_names()

write.csv(wide_algaeonly_cleaned_column_names, "algaeonly_cleaned_colnames.csv", row.names = FALSE)

```

```{r filtered data QA QC}
# you can use this to see which sites lost quads during filtering process 
#should be 50 per year/site/habitat combo

wide_alldata_filtered %>%
  dplyr::group_by(Year, Site, Habitat) %>%
  dplyr::summarise(count = n())

wide_alldata_filtered %>%
  dplyr::group_by(Year, Site) %>%
  dplyr::summarise(count = n())

```

## Anything below this line was to help identify missing data, overall taxa levels, etc. and does not necessarily need to be run to get a cleaned dataframe

```{r subsetting the data }

### CREATE DATAFRAME THAT IS JUST ALGAE
not_algae <- c("Ascidian", "Bare Space", "Coral", "Coral Rubble", "Corallimorpharia", "Heteractis sp.", "Millepora platyphylla", "No data", "Sand", "Sacrophyton sp.", "Shell Debris", "Soft Coral", "Sponge", "Tridacna sp.", "Zooxanthellae")

algae_df <- data[ ! data$Taxonomy_Substrate_Functional_Group %in% not_algae, ]
algae_df <- droplevels(algae_df)
levels(as.factor(algae_df$Taxonomy_Substrate_Functional_Group )) # 70 levels

### CREATE DATAFRAME THAT IS JUST MACROALGAE
# Maintaining mat-forming cyanos even though they're not macroalgae in the traditional sense
# "Coelothrix irregularis", "Symploca hydnoides" 
not_macro <- c("Algal Turf", "Crustose Corallines", "Cyanophyta", "Damselfish Turf", "Lithophyllum kotschyanum", "Neogoniolithon brassica-florida", "Sporolithon sp." )

macro_df <- algae_df[! algae_df$Taxonomy_Substrate_Functional_Group %in% not_macro, ]
macro_df <- droplevels(macro_df)
levels(as.factor(macro_df$Taxonomy_Substrate_Functional_Group )) # 63 levels

# NOTE: WHICH DO WE WANT TO WORK WITH?
# MY VOTE IS "algae_df"

```

```{r create wide df - algae only}
# using algae data
# I don't think the date of the transect matters. Dropping that.

algae_wide <- cast(algae_df, Year + Location + Site + Habitat + Site_habitat + Transect + Quadrat ~ Taxonomy_Substrate_Functional_Group, fun.aggregate = mean, value = "Percent_Cover")
# replace NaN in dataframe
# function from https://stackoverflow.com/questions/18142117/how-to-replace-nan-value-with-zero-in-a-huge-data-frame

is.nan.data.frame <- function(x){
  do.call(cbind, lapply(x, is.nan))}

algae_wide[is.nan.data.frame(algae_wide)] <- 0

length(levels(as.factor(algae_wide$Year))) # 19 years of data collection
length(levels(algae_wide$Site_habitat)) # 24 sites
# 5 transects and 10 quads per transect
# should be 22800 rows, so clearly some years are missing sites

```

```{r missing data part 1}

### NOTE - USED CHAT GPT FOR THIS NEXT PART
# Step 1: Create a contingency table of counts for each combination of year and site
year_site_counts <- table(comb_wide$Year, comb_wide$Site_habitat)

# Step 2: Identify rows with zero counts (indicating missing sites)
missing_years <- rownames(year_site_counts)[rowSums(year_site_counts == 0) > 0] # 2005, 2006, 2021 missing data

# should be 19200 rows if we remove those years
nrow(algae_wide[! algae_wide$Year %in% missing_years,]) # 19186... what are those 14 datapoints??

# Step 3: Extract the years corresponding to missing sites
for (year in missing_years) {
  # Filter the dataframe for the current missing year
  missing_year_data <- subset(algae_wide, Year == year)

  # Get the unique sites present in the missing year data
  present_sites <- unique(missing_year_data$Site_habitat)

  # Get all possible sites
  all_sites <- unique(algae_wide$Site_habitat)

  # Identify missing sites
  missing_sites <- setdiff(all_sites, present_sites)

  # Print missing sites for the current year
  cat("Year", year, "is missing data from sites:", paste(missing_sites, collapse = ", "), "\n")
}

# try to find what else is missing
algae_subset <- algae_wide[! algae_wide$Year %in% missing_years,]

site_transect_counts <- table(algae_subset$Site_habitat, algae_subset$Transect)
site_quadrat_counts <- table(algae_subset$Site_habitat, algae_subset$Quadrat)

# Step 2: Identify sites with missing transects or quadrats
missing_transect_sites <- rownames(site_transect_counts)[rowSums(site_transect_counts == 0) > 0]
missing_quadrat_sites <- rownames(site_quadrat_counts)[rowSums(site_quadrat_counts == 0) > 0]

# Step 3: Print out sites with missing transects or quadrats
print(paste("Sites with missing transects:", missing_transect_sites))
print(paste("Sites with missing quadrats:", missing_quadrat_sites))
 # Hm, nothing missing
```

```{r missing data part 2}
# create a dataframe with all possible combinations of year, site, habitat, transect, quadrat
year <- unique(data$Year)
sites <- unique(data$Site)
habitat <- unique(data$Habitat)
transect <- unique(data$Transect)
quadrat <- unique(data$Quadrat)
all_comb <- expand_grid(year, sites, habitat, transect, quadrat)
all_comb$comb <- paste(all_comb$year, all_comb$sites, all_comb$habitat, all_comb$transect, all_comb$quadrat)

# create a vector with the combinations that are actually present in the data
data$comb <- paste(data$Year, data$Site, data$Habitat, data$Transect, data$Quadrat) 
data_comb <- unique(data$comb)

missing_quadrats <- setdiff(all_comb$comb, data_comb)
length(missing_quadrats) # 1800 missing quadrats, consistent with the missing sites/years above
# extra missing values might be from quadrats with no algae in wide dataframe?

```