### 2024-07-03 julianna_data_exploration.Rmd
Exploratory diversity analyses with coral cover

---
title: "Pre working group data exploration: Mainly wrt coral cover"
output: html_document
date: "2024-07-02"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
require(lme4) # glmer()
require(emmeans) # emmeans()
```

# Bring in the data

These are the most recent file versions done by Lauren as of July 2, 2024

```{r}
allbenthos <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) 
algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv"))
```

# Algal alpha diversity

Make a spp matrix to use with vegan

```{r}
# matrix of just species and their relative percent covers in the same order as the algae tibble
algae %>% 
  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
```

```{r}
algae %>% 
  # grab the metadata info that is in the same order (can re-join after)
  select(year, location, site, habitat, site_habitat, transect, quadrat) %>% 
  mutate(Richness = specnumber(sppMat),
         Shannon = diversity(sppMat, "shannon"),
         Richness = specnumber(sppMat)) -> algDiv
```

## Plot

Broken out by year

```{r}
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
  geom_boxplot() +
  facet_grid(site~year) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```

Or without:

```{r}
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
  geom_boxplot() +
  facet_wrap(~site) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Even simpler:

```{r}
# shannon
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

# richness
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = Richness, group = habitat, fill = habitat)) +
  geom_histogram() +
  theme_bw() +
  facet_wrap(~habitat) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

```

Richness on top of each other

```{r}
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = Richness, group = habitat, fill = habitat)) +
  geom_bar(position = "dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

# Compare algal alpha diversity and coral cover

Make a joined dataframe for comparisons

```{r}
allbenthos %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, coral) %>% 
  full_join(algDiv) %>% 
  # add column for each transect ID
  rowwise() %>% 
  mutate(transect_id = paste0(site, habitat, year, transect)) -> coralAlgDiv  
```

## Plot shannon

Plot Shannon diversity across sites and habitats--what's up with this structure? Emailed Jamie and asked Lauren. TBD

```{r}
coralAlgDiv %>% 
  ggplot(aes(x = coral, y = Shannon, color = year)) +
  geom_point() + 
  facet_grid(habitat~site) +
  theme_classic()
```

Try by year:

```{r}
coralAlgDiv %>% 
  filter(habitat == "backreef") %>% 
  ggplot(aes(x = coral, y = Shannon, color = as.factor(transect))) +
  geom_point() + 
  facet_grid(year~site) +
  theme_classic()
```

## Plot richness

Plot richness across habitats and years

As points:

```{r}
coralAlgDiv %>% 
  ggplot(aes(x = coral, y = Richness, color = year, group = Richness)) +
  geom_jitter() + 
  facet_grid(habitat~site) +
  theme_classic()
```

```{r}
coralAlgDiv %>% 
  ggplot(aes(x = Richness, y = coral, color = year, group = Richness)) +
  geom_boxplot() + 
  facet_grid(year ~ habitat) +
  theme_classic()
```

Plot richness across sites and habitats

```{r}
coralAlgDiv %>% 
  ggplot(aes(x = Richness, y = coral, color = year, group = Richness)) +
  geom_boxplot() + 
  facet_grid(site ~ habitat) +
  theme_classic()
```

## Make a bad richness model

Dumb model--need to think more about random effects and figure out time component, but instructive to see there may be a relationship

```{r}
m1 <- glmer(Richness ~ scale(coral) + habitat + (1|site) + (1|transect_id) + (1|year), 
            family = poisson(link = "log"), data = coralAlgDiv)
  plot(m1)
  summary(m1)
  emmeans(m1, ~ habitat)
```

# Beta diversity

## Distance matrix

We get an error message if we run vegdist right away saying that we have empty rows and that their dissimilarities may be meaningless. Let's look to see if any rows have no algae at all:

```{r}
sppMat %>% 
  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
  filter(AlgSum == 0) %>% dim() # 672 with no algae at all
```

Need to run this without zero observations, so we'll exclude quadrats with no algae and make a species matrix

```{r}
# spp matrix
algae %>% 
  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
  filter(AlgSum != 0) %>% 
  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0

# metadata matrix
algae %>% 
  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
  filter(AlgSum != 0) %>%
  select(year, location, site, habitat, site_habitat, transect, quadrat) -> sppMeta0

```

Calculate Bray Curtis distances

```{r}
brayMat <- vegdist(sppMat0, "bray") # 1 is maximum dissimilarity
```

Calculate dispersion (this is using Noam's code)

```{r}
siteHabBetaDisper <- vegan::betadisper(bray_matrix, sppMeta0$site_habitat)

# plot of dispersion test
plot(year_beta_disper, sub = "") # lol
boxplot(year_beta_disper, main = "Year", xlab = "") 
```