### 2024-07-04 julianna_data_exploration.Rmd
Did a bunch of exploration to figure out what the bottom bound structure is in the Shannon algal vs. coral/turf/algae plots. Believe it has to do with instances when you have one dominant thing (e.g., turf algae) and then increase the proportion of a second alga incrementally at 4% intervals

---
title: "Pre working group data exploration: Mainly wrt coral cover"
output: html_document
date: "2024-07-02"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
require(lme4) # glmer()
require(emmeans) # emmeans()
```

# Bring in the data

These are the most recent file versions done by Lauren as of July 2, 2024

```{r}
allbenthos <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) 
algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
  # going to run everything without algal turf
  select(-algal_turf)
```

# Algal alpha diversity

Make a spp matrix to use with vegan

```{r}
# matrix of just species and their relative percent covers in the same order as the algae tibble
algae %>% 
  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
```

```{r}
algae %>% 
  # grab the metadata info that is in the same order (can re-join after)
  select(year, location, site, habitat, site_habitat, transect, quadrat) %>% 
  mutate(Richness = specnumber(sppMat),
         Shannon = diversity(sppMat, "shannon"),
        # Evenness = diversity(sppMat, "ev")
         Richness = specnumber(sppMat)) -> algDiv
```

## Plot

Broken out by year

```{r}
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
  geom_boxplot() +
  facet_grid(site~year) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```

Or without:

```{r}
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
  geom_boxplot() +
  facet_wrap(~site) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Even simpler:

```{r}
# shannon
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = habitat, y = Shannon, group = habitat)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

# richness
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = Richness, group = habitat, fill = habitat)) +
  geom_histogram() +
  theme_bw() +
  facet_wrap(~habitat) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

```

Richness on top of each other

```{r}
algDiv %>% 
  mutate(habitat = factor(habitat, levels = c("fringing", "backreef", "outer_10", "outer_17"))) %>% 
  ggplot(aes(x = Richness, group = habitat, fill = habitat)) +
  geom_bar(position = "dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

# Compare algal and coral cover

```{r}
# get ranges to calculate over
startAlg <- grep("acanthophora_spicifera", colnames(algae))
endAlg <- grep("valonia_ventricosa", colnames(algae))

algae %>% 
  mutate(AlgSum = rowSums(algae[startAlg:endAlg])) -> algSum

algSum %>% 
  full_join(allbenthos) %>% 
  ggplot(aes(x = coral, y = AlgSum, color = year)) +
  geom_point() +
  facet_grid(site~habitat)

```

# Compare algal alpha diversity and coral cover

Make a joined dataframe for comparisons

```{r}
allbenthos %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, coral) %>% 
  full_join(algDiv) %>% 
  # add column for each transect ID
  rowwise() %>% 
  mutate(transect_id = paste0(site, habitat, "_", year, "_", transect)) -> coralAlgDiv  
```

## Plot Shannon

Plot Shannon diversity across sites and habitats--what's up with this structure? Emailed Jamie and asked Lauren. TBD

```{r}
coralAlgDiv %>% 
  ggplot(aes(x = coral, y = Shannon, color = as.factor(Richness))) +
  geom_point(alpha = 0.5) + 
  facet_grid(habitat~site) +
  theme_classic() + 
  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed") # this is when two species are equally distributed (half half)--not sure why that would be
```

Try by year:

```{r}
coralAlgDiv %>% 
  filter(habitat == "backreef") %>% 
  ggplot(aes(x = coral, y = Shannon, color = as.factor(transect))) +
  geom_point() + 
  facet_grid(year~site) +
  theme_classic()
```

Or look at Shannon vs. the percent cover of algae in the plot

```{r}
coralAlgDiv %>% 
  full_join(algSum) %>% 
  ggplot(aes(x = AlgSum, y = Shannon, color = as.factor(Richness))) +
  geom_point(alpha = 0.5) + 
  facet_grid(habitat~site) +
  theme_classic() + 
  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
```

What about this when Richness == 2?

```{r}
coralAlgDiv %>% 
  full_join(algSum) %>% view()
  filter(Richness == 2) %>% 
  ggplot(aes(x = AlgSum, y = Shannon, color = as.factor(Richness))) +
  geom_point(alpha = 0.5) + 
  facet_grid(habitat~site) +
  theme_classic() + 
  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
```

Could this be because the only way to get a high enough imbalance/unevenness is having a large enough algal N

What does it look like relative to turf algae?

```{r}
coralAlgDiv %>% 
  full_join(algSum) %>% 
  filter(Richness == 2) %>% 
  ggplot(aes(x = algal_turf, y = Shannon, color = as.factor(Richness))) +
  geom_point(alpha = 0.5) + 
  facet_grid(habitat~site) +
  theme_classic() + 
  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
```

Nic: are these lines all turf but 1 line, all turf but 2 points line, all turf but 3 points line?

```{r}
coralAlgDiv %>% 
  full_join(algSum) %>% 
  filter(Richness == 2) %>% 
  mutate(Nonturf = AlgSum - algal_turf) %>% select(Nonturf) %>% unique() %>% view()
  ggplot(aes(x = algal_turf, y = Shannon, color = as.factor(Nonturf))) +
  geom_point(alpha = 0.5) + 
  facet_grid(habitat~site) +
  theme_classic() + 
  geom_hline(yintercept = 0.693, color = "red", linetype = "dashed")
```

^This seems to be true here. It seems like lines have to do with how much of the non-turf you have. They only do 25 points and multiply by 4 so everything is a multiple of 4. But then how can we have a 2 there???

## Problem solve Shannon

Trying to instead calculate Shannon manually with 100 as the denominator?

```{r}
shannon100 <- function(x, equalize.groups = FALSE, base = exp(1)) {
  x <- drop(as.matrix(x))
  if (length(dim(x)) > 1) {
        x <- sweep(x, MARGIN = 1, 100, FUN = "/") # MARGIN = 1 I think means to do this rowwise; FUN = "/" means to divide by 100 (replaced the row sum with 100 here)
    }
    else {
        x <- x/(100 <- sum(x)) # replaced total with 100 here
    }
  x <- -x * log(x, base)
  if (length(dim(x)) > 1) 
        H <- apply(x, MARGIN = 1, sum, na.rm = TRUE)
    else H <- sum(x, na.rm = TRUE)
  H
}

# add this to the df
algae %>% 
  # grab the metadata info that is in the same order (can re-join after)
  select(year, location, site, habitat, site_habitat, transect, quadrat) %>% 
  mutate(Richness = specnumber(sppMat),
         Shannon = diversity(sppMat, "shannon"),
         Shannon100 = shannon100(sppMat),
         Richness = specnumber(sppMat)) -> algDiv100
```

Plot

```{r}
algDiv100 %>% 
  ggplot(aes(x = Shannon100, y = Shannon)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") 
```

Re-make old plot

```{r}
# re-join with coral
allbenthos %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, coral) %>% 
  full_join(algDiv100) %>% 
  # add column for each transect ID
  rowwise() %>% 
  mutate(transect_id = paste0(site, habitat, "_", year, "_", transect)) -> coralAlgDiv100

# plot  
coralAlgDiv100 %>% 
  ggplot(aes(x = coral, y = Shannon, color = year)) +
  geom_point() + 
  facet_grid(habitat~site) +
  theme_classic() # NOPE very much still there 
```

## Plot richness

Plot richness across habitats and years

As points:

```{r}
coralAlgDiv %>% 
  ggplot(aes(x = coral, y = Richness, color = year, group = Richness)) +
  geom_jitter() + 
  facet_grid(habitat~site) +
  theme_classic()
```

```{r}
coralAlgDiv %>% 
  ggplot(aes(x = Richness, y = coral, color = year, group = Richness)) +
  geom_boxplot() + 
  facet_grid(year ~ habitat) +
  theme_classic()
```

Plot richness across sites and habitats

```{r}
coralAlgDiv %>% 
  ggplot(aes(x = Richness, y = coral, color = year, group = Richness)) +
  geom_boxplot() + 
  facet_grid(site ~ habitat) +
  theme_classic()
```

## Make a bad richness model

Dumb model--need to think more about random effects and figure out time component, but instructive to see there may be a relationship

```{r}
m1 <- glmer(Richness ~ scale(coral) + habitat + (1|site) + (1|transect_id) + (1|year), 
            family = poisson(link = "log"), data = coralAlgDiv)
  plot(m1) # I think this looks pretty weird right?
  summary(m1)
  emmeans(m1, ~ habitat)
```

# Beta diversity

## Distance matrix

We get an error message if we run vegdist right away saying that we have empty rows and that their dissimilarities may be meaningless. Let's look to see if any rows have no algae at all:

```{r}
sppMat %>% 
  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
  filter(AlgSum == 0) %>% dim() # 672 with no algae at all
```

Need to run this without zero observations, so we'll exclude quadrats with no algae and make a species matrix

```{r}
# spp matrix
algae %>% 
  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
  filter(AlgSum != 0) %>% 
  select(-year, -location, -site, -habitat, -site_habitat, -transect, -quadrat, -AlgSum) -> sppMat0

# metadata matrix
algae %>% 
  mutate(AlgSum = rowSums(across(where(is.numeric)))) %>% 
  filter(AlgSum != 0) %>%
  select(year, location, site, habitat, site_habitat, transect, quadrat) -> sppMeta0

```

Calculate Bray Curtis distances

```{r}
brayMat <- vegdist(sppMat0, "bray") # 1 is maximum dissimilarity
```

Calculate dispersion (this is using Noam's code)

```{r}
siteHabBetaDisper <- vegan::betadisper(bray_matrix, sppMeta0$site_habitat)

# plot of dispersion test
plot(year_beta_disper, sub = "") # lol
boxplot(year_beta_disper, main = "Year", xlab = "") 
```