### 2024-07-23 noam_data_exploration_v3.Rmd
Added a cover vs cv figure, but I don't like it. Played around with how to calculate portfolio effects some more.

---
title: "noam_data_exploration_v1"
author: "Noam Altman-Kurosaki"
date: "2024-05-09"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(ecofolio) # possible other method for examining portfolio effects 
```

```{r import and format data}
# starting with all data now because of 0 algae transects/quadrats
data <- read.csv(here::here("data", "allbenthicdata_cleaned.csv"), stringsAsFactors = F)

not_algae <- c("Ascidian", "Bare.Space", "Coral", "Coral.Rubble", "Corallimorpharia", "Heteractis.sp.", "Millepora.platyphylla", "No.data", "Sand", "Sacrophyton.sp.", "Shell.Debris", "Soft.Coral", "Sponge", "Tridacna.sp.", "Zooxanthellae", "row_sums")

data <- data[ , -which(names(data) %in% not_algae)]

# adding island side - adapted from Julianna's code
data <- data %>% mutate(Island_side = case_when(Site == "LTER 1" ~ "North",
                                 Site == "LTER 2" ~ "North",
                                 Site == "LTER 3" ~ "East",
                                 Site == "LTER 4" ~ "East",
                                 Site == "LTER 5" ~ "West",
                                 Site == "LTER 6" ~ "West"))
data$Island_side_habitat <- paste(data$Island_side, data$Habitat)

str(data)

meta_cols <- c(colnames(data[1:7] ), "Island_side", "Island_side_habitat", "Stratum")
species_cols <- colnames(data[8:77])

#pivot long 
long_data <- pivot_longer(
  data = data,
  cols = 8:77,
  names_to = "Taxa",
  values_to = "Percent_cover")

 transect_level <- long_data %>%
   dplyr::group_by(Year, Site, Habitat, Transect, Taxa) %>%
   dplyr::summarise(Mean_cover = mean(Percent_cover), 
             SE = sd(Percent_cover)/sqrt(n()))

# site_level <- transect_level %>%
#   dplyr::group_by(Year, Site, Habitat, Taxa) %>%
#   dplyr::summarise(Mean_site_cover = mean(Mean_cover), 
#             Site_se = sd(Mean_cover)/sqrt(n()))
```

```{r diversity indicies, fig.height = 15, fig.width=20}

wide_transect_level <- data %>%
  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat, Transect) %>%
  dplyr::summarize(across(all_of(species_cols), mean, na.rm = TRUE))

alpha_diversity <-
  data.frame(
    Year = as.factor(wide_transect_level$Year),
    Island_side = as.factor(wide_transect_level$Island_side),
    Island_side_habitat = as.factor(wide_transect_level$Island_side_habitat),
    Site = as.factor(wide_transect_level$Site),
    Habitat = as.factor(wide_transect_level$Habitat),
    Site_habitat = as.factor(wide_transect_level$Site_habitat),
    Transect = as.factor(wide_transect_level$Transect),
    Site_habitat_transect = as.factor(paste(wide_transect_level$Site_habitat, wide_transect_level$Transect)),
    Shannon = diversity(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)], "shannon"),
    Richness = rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)] > 0)
  )

alpha_diversity_means <- alpha_diversity %>%
  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()))

site_colours <- c("LTER 1" = "cornflowerblue",
               "LTER 2" = "darkblue",
               "LTER 3" = "yellowgreen",
               "LTER 4" = "forestgreen",
               "LTER 5" = "mediumorchid",
               "LTER 6" = "orange")
island_side_colours <- c("North" = "darkblue",
                         "East" = "mediumorchid",
                         "West" = "orange")
habitat_linetypes <- c("Fringing" = "solid",
                       "Backreef" = "longdash",
                       "Outer 10" = "twodash",
                       "Outer 17" = "dotted")
habitat_shapes <- c("Fringing" = 16,
                     "Backreef" = 17,
                     "Outer 10" = 1,
                     "Outer 17" = 2)

(richness_plot <-
  ggplot(data = alpha_diversity_means) +
  geom_point(aes(x = Year, y = Richness_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
  geom_line(aes(x = Year, y = Richness_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Island_side,
                    group = Island_side_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = island_side_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat)+
  labs(title = "a. Richness\n", y = "Richness\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

(shannon_plot <-
  ggplot(data = alpha_diversity_means) +
  geom_point(aes(x = Year, y = Shannon_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
  geom_line(aes(x = Year, y = Shannon_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Island_side,
                    group = Island_side_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = island_side_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat)+
  labs(title = "b. Shannon\n", y = "shannon\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

```

```{r alpha diversity panel graph, fig.height = 30, fig.width= 20}

ggarrange(richness_plot + rremove("x.text"), shannon_plot, common.legend = T,
          nrow = 2)

# seems like patterns in alpha diversity different depending on if you look at richness or shannon diversity 
```

```{r alpha diversity quadrat level, fig.height = 15, fig.width=20}
alpha_diversity_quad <-
  data.frame(
    Year = as.numeric(data$Year),
    Location = as.factor(data$Location),
    Island_side = as.factor(data$Island_side),
    Island_side_habitat = as.factor(data$Island_side_habitat),
    Site = as.factor(data$Site),
    Habitat = as.factor(data$Habitat),
    Site_habitat = as.factor(data$Site_habitat),
    Transect = as.factor(data$Transect),
    Site_habitat_transect = as.factor(paste(data$Site_habitat, data$Transect)),
    Shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
    Richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
  )

alpha_diversity_quad_means <- alpha_diversity_quad %>%
  dplyr::group_by(Year, Island_side, Island_side_habitat, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()))

(richness_quad_plot <-
  ggplot(data = alpha_diversity_quad_means) +
  geom_point(aes(x = Year, y = Richness_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8) +
  geom_line(aes(x = Year, y = Richness_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Richness_mean - Richness_se , ymax = Richness_mean + Richness_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat)+
  labs(title = "a. Richness quad level\n", y = "Richness\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

(shannon_quad_plot <-
  ggplot(data = alpha_diversity_quad_means) +
  geom_point(aes(x = Year, y = Shannon_mean, colour = Island_side, shape = Habitat, group = Island_side_habitat), size = 8) +
  geom_line(aes(x = Year, y = Shannon_mean, colour = Island_side, group = Island_side_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Shannon_mean - Shannon_se , ymax = Shannon_mean + Shannon_se, colour = Island_side,
                    group = Island_side_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = island_side_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat)+
  labs(title = "b. Shannon quad level\n", y = "shannon\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

```
```{r compare quad and transect level plots, fig.height = 30, fig.width= 40}
ggarrange(richness_plot + rremove("x.text"), richness_quad_plot + rremove("x.text"),
          shannon_plot,shannon_quad_plot,
          common.legend = T,
          nrow = 2, ncol = 2)

```
```{r gam approach}
# NOTE: Not sure where I'm going with this yet
# Idea is to look at smoother estimates for each time series to get an idea for the "shape" of the time series
# Not sure if worthwhile, might skip

# The below takes 17 years to run in and of itself
# gam_model <- gam(Richness ~ s(Year) + s(Year, by = Habitat) + 
#                  s(Island_side, bs = "re") + s(Transect, bs = "re") + s(Location, bs = "re"),
#                  data = alpha_diversity_quad, family = poisson())

gam_columns <- c("Location", "Intercept", "edf") # names of columns for dataframe
gam_coefficients <- # new dataframe name - same workflow as above
  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$Location)),
                    ncol = length(gam_columns))
             )
colnames(gam_coefficients) <- gam_columns # change column names to what was specified
gam_coefficients$Location <- levels(alpha_diversity_quad$Location) # populate Location column with each quad/time series

# same workflow as *below (I did gls first originally)
for(i in 1:length(unique(alpha_diversity_quad$Location)) ){ # for each  quadrat
  # create a new dataframe for the quadrat of the current iteration
  newData <- subset(alpha_diversity_quad, Location == levels(alpha_diversity_quad$Location)[i])

  # fit gam for current time series
  loop_mod <- gamm(Richness ~ s(Year), data = newData, correlation = corCAR1(form = ~ Year), method = "REML", family = poisson)
  loop_sum <- summary(loop_mod$gam) # store summary data
  # populate appropriate row and column with the relevant coeffcients from gls
  gam_coefficients[i, 2] <- unname(loop_sum$p.coeff)
  gam_coefficients[i, 3] <- unname(loop_sum$edf)

}

hist(gam_coefficients$Intercept)
hist(gam_coefficients$edf) # generally around 1 except for an outlier
# LTER 5 Fringing Reef Algae Transect 3 Quad 6 - a quadrat that has almost no algae the entire time series
# issues with quadrat - smaller sample = more 0s, makes modeling trickier
hist(subset(gam_coefficients, Location != "LTER 5 Fringing Reef Algae Transect 3 Quad 6")$edf)
sum(gam_coefficients$edf > 1.01) # all but 18/1200 quadrats can be fit by a smoother with edf < 1.01
# trends more straightforward than I anticipated?
# argument for just looking at mean richness and slope?
```

```{r gls approach}

# FRAMING: What is the general trend in algal alpha diversity over the course of the time series? Increasing, decreasing, neutral?

columns <- c("Location", "Intercept", "Slope") # names of columns for dataframe
gls_coefficients <- # new dataframe name
  data.frame(matrix(nrow = length(unique(alpha_diversity_quad$Location)), # 1200 quadrats = 1200 different time series
                    ncol = length(columns)) # specified to make sure number of columns doesn't exceed what was specified
             )
colnames(gls_coefficients) <- columns # change column names to what was specified
gls_coefficients$Location <- levels(alpha_diversity_quad$Location) # populate Location column with each quad/time series

for(i in 1:length(unique(alpha_diversity_quad$Location)) ){ # for each  quadrat
  # create a new dataframe for the quadrat of the current iteration
  newData <- subset(alpha_diversity_quad, Location == levels(alpha_diversity_quad$Location)[i])

  # fit gls for current time series
  loop_mod <- gls(Richness ~ Year, correlation =corAR1(form=~Year), data = newData)

  # populate appropriate row and column with the relevant coeffcients from gls
  gls_coefficients[i, 2] <- unname(coef(loop_mod)[1])
  gls_coefficients[i, 3] <- unname(coef(loop_mod)[2])

}

hist(gls_coefficients$Slope) # looks basically normal, bounded -Inf, Inf

hist(gls_coefficients$Intercept) # same with intercept - although it probably shouldn't be

# Start adding back in relevant metadata - probably a more elegant way to do this!!!
habitat_vector <- c(rep("Backreef", 50), rep("Fringing", 50), rep("Outer 10", 50), rep("Outer 17", 50))
transect_sequence <- c(rep(1,10),
                       rep(2,10),
                       rep(3,10),
                       rep(4,10),
                       rep(5,10))
extra_meta <- 
  data.frame(Site = as.factor(c(rep("LTER 1", 200),
                                rep("LTER 2", 200),
                                rep("LTER 3", 200),
                                rep("LTER 4", 200),
                                rep("LTER 5", 200),
                                rep("LTER 6", 200)) ),
             Island_side = as.factor(c(rep("North", 400),
                                     rep("East", 400),
                                     rep("West", 400))
                                     ),
             Habitat = as.factor(rep(habitat_vector, 6 )),
             Transect = as.factor(rep(transect_sequence, 240)),
             Location = levels(alpha_diversity_quad$Location) )

gls_coefficients_merged <- merge(gls_coefficients, extra_meta, by = "Location")

slope_mod <- glmmTMB(Slope ~ Habitat*Site + (1|Site/Transect), data = gls_coefficients_merged)
# model convergence problem
summary(slope_mod) # pretty low variance in random effects
plot(gls_res <- simulateResiduals(slope_mod)) # there are issues, but I can't figure out what they are...
# I think this is due to 1200 obs in model
hist(residuals(slope_mod))
plot(predict(slope_mod) ~ residuals(slope_mod))
Anova(slope_mod)
# Response: Slope
#                 Chisq Df Pr(>Chisq)    
# Habitat      1397.999  3  < 2.2e-16 ***
# Site           39.095  5  2.273e-07 ***
# Habitat:Site 1738.155 15  < 2.2e-16 ***
r.squaredGLMM(slope_mod) # 0.22, 0.24
plot(allEffects(slope_mod))
# Backreef: richness increases on East and North shore, but relatively constant west side
# Fringe: richness decreasing on east and west side, but relatvely constant on north shore
# Outer 10: Richness neutral/decreasing on east side, but neutral/increasing on north and west side
# Outer 17: Richness decreasing north/east side, but neutral west side
# Worth noting trends are slight - arguable whether they are biologically meaningful

slope_emm <- emmeans(slope_mod, ~ Habitat*Site) # create emmeans object
emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
# Note: not sure I want to include this
emmeans(slope_mod, pairwise ~ Site | Habitat)
```

```{r gls plot, fig.height=20, fig.width=20}
# not sure how to visualize here... thinking something like Dornelas et al. 2014 Fig 2a
# I reached out to them about getting their code for this

ggplot(data = gls_coefficients_merged) +
  geom_histogram(aes(x = Slope), bins = 30) +
  geom_vline(xintercept = 0, size = 2, linetype = 2) +
  facet_grid(Habitat ~ Site) +
#  theme_classic() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain")
  )

```

```{r gls intercept, fig.height=20, fig.width=20}
intercept_mod <- glmmTMB(Intercept ~ Habitat*Island_side + (1|Site/Transect), data = gls_coefficients_merged)
summary(intercept_mod) # a lot of variance w/i transect, less so w/i site
plot(gls_int_res <- simulateResiduals(intercept_mod)) # same issues as above

hist(residuals(intercept_mod))
plot(predict(intercept_mod) ~ residuals(intercept_mod))
Anova(intercept_mod)
#                        Chisq Df Pr(>Chisq)    
# Habitat             1307.838  3  < 2.2e-16 ***
# Island_side           26.358  2  1.889e-06 ***
# Habitat:Island_side  789.627  6  < 2.2e-16 ***
r.squaredGLMM(intercept_mod) # 0.16, 0.18
plot(allEffects(intercept_mod))

ggplot(data = gls_coefficients_merged) +
  geom_histogram(aes(x = Intercept), bins = 30) +
  geom_vline(xintercept = 0, size = 2, linetype = 2) +
  facet_grid(Habitat ~ Island_side) +
#  theme_classic() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain")
  )

```

```{r glmm approach}
alpha_diversity_quad$Year_factor <- as.factor(alpha_diversity_quad$Year) # for glmmTMB ar1 structure
# Framing - is adiv generally higher in certain habitats/regions?

# richness_mod <- glmmTMB(
#   Richness ~ Island_side*Habitat + (1 | Island_side/Site/Transect/Location) + ar1(Year_factor + 0 | Location),
#   family = poisson,
#   data = alpha_diversity_quad,
#   control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000))
# ) # overspecified but going with it for now
# summary(richness_mod) # looks like strong autocorrelation, which makes sense to me
# Anova(richness_mod)
# plot(allEffects(richness_mod))

# probably the closest to what others have done:
richness_mod_2 <- glmmTMB(
  Richness ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
  family = poisson,
  data = alpha_diversity_quad,
  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)), na.action = na.pass
) # overspecified but going with it for now

summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
Anova(richness_mod_2)
plot(allEffects(richness_mod_2))
# emmeans(richness_mod_2, pairwise ~ Habitat | Year_factor) # if we want to compare habitats year to year
# model.sel(dredge(richness_mod_2))

top_rich_mod <- glmmTMB(Richness ~ Year_factor*Habitat + Year_factor*Site + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),   family = poisson, data = alpha_diversity_quad, control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)))
summary(top_rich_mod)
Anova(top_rich_mod)
```
```{r turnover troubleshooting, fig.height = 30, fig.width=25}

transect_level$Site_habitat <-  paste(transect_level$Site, transect_level$Habitat)
transect_level$Stratum <- paste(transect_level$Site, "-", transect_level$Habitat, "-", transect_level$Transect)

# long_data$Site_habitat <-  paste(long_data$Site, long_data$Habitat)
# long_data$Stratum <- paste(long_data$Site, "-", long_data$Habitat, "-", long_data$Transect)

# turnover analyses break when I try to do it at the quadrat level
# checking if it's because there are quadrats that have effecively 0% algae
total_cover_per_quadrat_year <- long_data %>%
  group_by(Location, Year) %>%
  summarize(total_cover = sum(Percent_cover), .groups = 'drop')

# Filter quadrats where the total percent cover is 0%
zero_cover_quadrats <- total_cover_per_quadrat_year %>%
  filter(total_cover == 0)

# Count the number of years each quadrat had 0% cover
zero_cover_count <- zero_cover_quadrats %>%
  group_by(Location) %>%
  summarize(years_with_zero_cover = n())

# Filter quadrats that had 0% cover for multiple years
quadrats_multiple_years_zero_cover <- zero_cover_count %>%
  filter(years_with_zero_cover > 1)

length(unique(quadrats_multiple_years_zero_cover$Location)) # 144/1200 quadrats have multiple years with no algal cover

# compare this to the number of transects with 0 algae cover
total_cover_per_transect_year <- transect_level %>%
  group_by(Stratum, Year) %>%
  summarize(total_cover = sum(Mean_cover), .groups = 'drop')

# Filter quadrats where the total percent cover is 0%
zero_cover_transects <- total_cover_per_transect_year %>%
  filter(total_cover == 0)

# Count the number of years each quadrat had 0% cover
zero_cover_transect_count <- zero_cover_transects %>%
  group_by(Stratum) %>%
  summarize(years_with_zero_cover = n())

# Filter quadrats that had 0% cover for multiple years
transects_multiple_years_zero_cover <- zero_cover_transect_count %>%
  filter(years_with_zero_cover > 1)

# NO transects ever have more than one year with 0 algal cover, as opposed to 144 quadrats
# Fairly certain this is why the turnover analyses are breaking

# identify quadrats missing timepoints
years <- unique(long_data$Year)
quadrats <- unique(long_data$Location)

# Create a dataframe with all combinations of quadrats and years
all_combinations <- expand.grid(Location = quadrats, Year = years)

# Identify the missing combinations by performing an anti-join with your actual data
missing_timepoints <- anti_join(all_combinations, long_data, by = c("Location", "Year"))

# Display the missing timepoints
print(missing_timepoints)

problem_children <- c(
                      "LTER 5 Fringing Reef Algae Transect 3 Quad 9",
                      "LTER 5 Fringing Reef Algae Transect 3 Quad 8",
                      "LTER 5 Fringing Reef Algae Transect 3 Quad 6",
                      "LTER 5 Fringing Reef Algae Transect 3 Quad 3",
                      "LTER 5 Fringing Reef Algae Transect 3 Quad 10")

long_data_sub <- long_data[!long_data$Location %in% problem_children, ] # remove problem children

data_sub <- data[!data$Location %in% problem_children, ]

```

```{r total turnover, fig.height = 24, fig.width=16}
# the turnover function calculates turnover as the proportion or species that either appear or disappear between timepoints
total_turnover<-
  turnover(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location",
  metric = "total"
)

# get relevant metadata for turnover metrics
data_sub <- data[paste(data$Year, data$Location) %in% paste(total_turnover$Year, total_turnover$Location), ]

# check to make sure everything is the same length
# nrow(data_sub) # 17829
# nrow(total_turnover) # 17829

# extract metadata
turnover_meta <- data_sub[ ,which(names(data_sub) %in% meta_cols)]

# merge with turnover df
total_turnover <- merge(total_turnover, turnover_meta, by = c("Year", "Location"))
total_turnover$Year_factor <- as.factor(total_turnover$Year)

# turnover bounded between 0 and 1 --> beta regression
# need to transform due to presence of 0s and 1s
# Smithson verkuilen transformation for beta-regression
sv_trans = function(prop, N, s = 0.000005){
  (prop*(nrow(N) - 1) + s)/nrow(N)
  # where prop is the proportional value you're transforming, 
  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
  # and s is a small offset 
}
total_turnover$Turnover_trans <- sv_trans(total_turnover$total, total_turnover)

# model
total_turnover_mod1 <- glmmTMB(Turnover_trans ~ Year_factor*Habitat*Site +
                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
                               family = beta_family(), data = total_turnover)
summary(total_turnover_mod1) # low temporal autocorrelation... maybe don't need ar1 term?
plot(total_turnover_res <- simulateResiduals(total_turnover_mod1)) # residuals look very weird
Anova(total_turnover_mod1) # obviously everything sig

total_turnover_mod2 <- glmmTMB(Turnover_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
                               family = beta_family(), data = total_turnover)
summary(total_turnover_mod2) # 
plot(total_turnover_res2 <- simulateResiduals(total_turnover_mod2)) # residuals still look like shit
Anova(total_turnover_mod2) 
anova(total_turnover_mod2, total_turnover_mod1) # models are sig different from each other
AICc(total_turnover_mod2, total_turnover_mod1) # first model is better
# need to figure out residuals problem
plot(allEffects(total_turnover_mod1))

# plot
total_turnover_means <-
  total_turnover %>%
  dplyr::group_by(Year, Island_side, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Turnover_mean = mean(total), 
            Turnover_se = sd(total)/sqrt(n()))

(turnover_plot <-
  ggplot(data = total_turnover_means) +
  geom_point(aes(x = Year, y = Turnover_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Turnover_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Turnover_mean - Turnover_se , ymax = Turnover_mean + Turnover_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "a. Turnover\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
#
```

```{r appearances, fig.height = 24, fig.width=16}
# examine appearance and disappearances independently 
# refers to the proportion of species that appeared or disappeared relative to total species pool between two time points

appearance <-
  turnover(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location",
  metric = "appearance"
)

appearance <- merge(turnover_meta, appearance, by = c("Year", "Location"))
appearance$Appearance_trans <- sv_trans(appearance$appearance, appearance)
appearance$Year_factor <- as.factor(appearance$Year)

# model
appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ Year_factor*Habitat*Site +
                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
                               family = beta_family(), data = appearance)
summary(appearance_turnover_mod1) # low temporal autocorrelation and it's negative
plot(appearance_turnover_res <- simulateResiduals(appearance_turnover_mod1)) # similar residuals problem
Anova(appearance_turnover_mod1) # obviously everything sig

appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
                               family = beta_family(), data = appearance)
summary(appearance_turnover_mod2) # 
plot(appearance_turnover_res2 <- simulateResiduals(appearance_turnover_mod2)) #
Anova(appearance_turnover_mod2) 
anova(appearance_turnover_mod2, appearance_turnover_mod1) # models are sig different from each other
AICc(appearance_turnover_mod2, appearance_turnover_mod1) # first model is better but not as drastic as total

plot(allEffects(appearance_turnover_mod1))

appearance_means <-
  appearance %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Appearance_mean = mean(appearance), 
            Appearance_se = sd(appearance)/sqrt(n()))

(appearance_plot <-
  ggplot(data = appearance_means) +
  geom_point(aes(x = Year, y = Appearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Appearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "b. Appearances\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
#
```

```{r disappearances, fig.height = 24, fig.width=16}
disappearance <-
  turnover(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location",
  metric = "disappearance"
)

disappearance <- merge(turnover_meta, disappearance, by = c("Year", "Location"))
disappearance$Disappearance_trans <- sv_trans(disappearance$disappearance, disappearance)
disappearance$Year_factor <- as.factor(disappearance$Year)

# model
disappearance_turnover_mod1 <- glmmTMB(Disappearance_trans ~ Year_factor*Habitat*Site +
                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
                               family = beta_family(), data = disappearance)
summary(disappearance_turnover_mod1) # strong autocorrelation here
plot(disappearance_turnover_res <- simulateResiduals(disappearance_turnover_mod1)) # 
Anova(disappearance_turnover_mod1) #

disappearance_turnover_mod2 <- glmmTMB(Disappearance_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
                               family = beta_family(), data = disappearance)
summary(disappearance_turnover_mod2) # 
plot(disappearance_turnover_res2 <- simulateResiduals(disappearance_turnover_mod2)) #
Anova(disappearance_turnover_mod2) 
anova(disappearance_turnover_mod2, disappearance_turnover_mod1) # only case where models are not sig different from each other
AICc(disappearance_turnover_mod2, disappearance_turnover_mod1) # models almost equivalent

plot(allEffects(disappearance_turnover_mod1))

disappearance_means <-
  disappearance %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Disappearance_mean = mean(disappearance), 
            Disappearance_se = sd(disappearance)/sqrt(n()))

(disappearance_plot <-
  ggplot(data = disappearance_means) +
  geom_point(aes(x = Year, y = Disappearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Disappearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "c. Disappearance\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
```

```{r mean rank shift, fig.height = 24, fig.width=16}
# rank shift

rank_shift<-
  rank_shift(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location"
)

rank_shift <- merge(turnover_meta, rank_shift, by = c("Location"))

rank_shift_means <-
  rank_shift %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Rank_shift_mean = mean(MRS), 
            Rank_shift_se = sd(MRS)/sqrt(n()))
# Looks very weird - not sure what to do with this
(rank_shift_plot <-
  ggplot(data = rank_shift_means) +
  geom_point(aes(x = Year, y = Rank_shift_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Rank_shift_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Rank_shift_mean - Rank_shift_se , ymax = Rank_shift_mean + Rank_shift_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "a. rank_shift\n", y = "Total rank_shift\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
```

```{r rate of change, fig.height=15, fig.width=15}
# rate change
long_data_sub$Percent_cover_transformed <- wisconsin(long_data_sub$Percent_cover) # wisconsin transform abundance data

# NOTE: the results here vary considerably if you used transformed vs untransformed data
# Looking into way to change distance matrix from Euclidiean to something else
rate_change_table_transformed<-
  rate_change(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover_transformed",
  replicate.var = "Location"
) # summarizes rate of change for each quadrat over the course of the time series

rate_change_table_transformed <- merge(turnover_meta, rate_change_table_transformed, by = "Location")
hist(rate_change_table_transformed$rate_change)

rate_change_mod1 <- glmmTMB(rate_change ~ Habitat*Site + (1|Habitat/Site/Transect), data = rate_change_table_transformed)
summary(rate_change_mod1)
plot(rc_res <- simulateResiduals(rate_change_mod1)) # some deviations, but sig likely due to large n
Anova(rate_change_mod1)
#                Chisq Df Pr(>Chisq)    
# Habitat      187.988  3  < 2.2e-16 ***
# Site          81.209  5  4.686e-16 ***
# Habitat:Site 143.638 15  < 2.2e-16 ***
plot(allEffects(rate_change_mod1))
# Outer 17 has fewest differences
# LTER 1 2 3 changing more than 4 5 6 at outer10
# Backreef change highest at LTER 1 2
# Fringing highest at LTER 1 2 6
emmeans(rate_change_mod1, pairwise ~ Site | Habitat)

# Untransformed
rate_change_table_untransformed<-
  rate_change(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location"
) # summarizes rate of change for each quadrat over the course of the time series

rate_change_table_untransformed <- merge(turnover_meta, rate_change_table_untransformed, by = "Location")
hist(rate_change_table_untransformed$rate_change)

rate_change_mod2 <- glmmTMB(rate_change ~ Habitat*Site + (1|Habitat/Site/Transect), data = rate_change_table_untransformed)
summary(rate_change_mod2)
plot(rc_res <- simulateResiduals(rate_change_mod2)) # as above
Anova(rate_change_mod2)
#                Chisq Df Pr(>Chisq)    
# Habitat       39.224  3  1.556e-08 ***
# Site          17.882  5   0.003098 ** 
# Habitat:Site 123.014 15  < 2.2e-16 ***
plot(allEffects(rate_change_mod1))
# eyeballing it
# Outer 17 scarcely changing regardless of site
# Rate of change similar in backreef between sites
# big variance between sites in outer10 and fringing
# Outer 10: As you go around the island, sites become increasing more "stable" over time (i.e. less shift toward dissimilarity)
# Effectively the opposite trend in Fringing - more dissimilar over time as you go around the island
emmeans(rate_change_mod1, pairwise ~ Site | Habitat)

ggplot(data = rate_change_table_transformed) +
  geom_boxplot(aes(x = Site, y = rate_change, colour = Site) ) +
  geom_jitter(aes(x = Site, y = rate_change, colour = Site), width = 0.2) +
  facet_wrap(~Habitat) +
  scale_colour_manual(values = site_colours) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )

ggplot(data = rate_change_table_untransformed) +
  geom_boxplot(aes(x = Site, y = rate_change, colour = Site) ) +
  geom_jitter(aes(x = Site, y = rate_change, colour = Site), width = 0.2) +
  facet_wrap(~Habitat) +
  scale_colour_manual(values = site_colours) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )

```

```{r rate change vis method 2, fig.height=15, fig.width=15}
rate_change_untransformed<-
  rate_change_interval(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location"
)

rate_change_untransformed_distances <- merge(rate_change_untransformed, turnover_meta, by = "Location")

rate_change_transformed<-
  rate_change_interval(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover_transformed",
  replicate.var = "Location"
)

rate_change_transformed_distances <- merge(rate_change_transformed, turnover_meta, by = "Location")

 ggplot(rate_change_untransformed_distances,
        aes(x = interval, y = distance, colour = Site, shape = Habitat) )+
   facet_wrap(~Habitat) + 
   geom_point() +
   scale_colour_manual(values = site_colours) +
   scale_linetype_manual(values = habitat_linetypes) +
   stat_smooth(method = "lm", se = T, size = 2) +
   scale_shape_manual(values = habitat_shapes) +
   labs(y = "Untransformed distance\n") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") # make legend horizontal
         )

 ggplot(rate_change_transformed_distances,
        aes(x = interval, y = distance, colour = Site, shape = Habitat) )+
   facet_wrap(~Habitat) + 
   geom_point() +
   scale_colour_manual(values = site_colours) +
   scale_linetype_manual(values = habitat_linetypes) +
   stat_smooth(method = "lm", se = T, size = 2) +
   scale_shape_manual(values = habitat_shapes) +
   labs(y = "Transformed distance\n") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") # make legend horizontal
         ) 
```

```{r multivariate change}
## Ran this for 13 hours and it didn't finish (obviously)
## Need to subset or specify questions or try remote access on more powerful computer
# multi_change_untransformed <-
#   multivariate_change(
#    df = long_data_sub,
#    time.var = "Year",
#    species.var = "Taxa",
#    abundance.var = "Percent_cover",
#    replicate.var = "Location"
#   )
#

```

```{r transect level portfolio effect}
# using Anderson et al. 2013 definitions of portfolio effects - relevant fxns in ecofolio packages
wide_transect_level$Stratum <- as.factor(paste(wide_transect_level$Site, "-", wide_transect_level$Habitat, "-", wide_transect_level$Transect))

transect_portfolio_columns <- c("Stratum", "Portfolio_effect") # names of columns for dataframe
transect_level_portfolio <- 
  data.frame(matrix(nrow = length(unique(wide_transect_level$Stratum)), 
                    ncol = length(transect_portfolio_columns))
             )
colnames(transect_level_portfolio) <- transect_portfolio_columns # change column 
transect_level_portfolio$Stratum <- levels(wide_transect_level$Stratum) # populate

for(i in 1:length(unique(unique(wide_transect_level$Stratum))) ){ # for each  transect
  # create a new dataframe for current transect
  newData <- subset(wide_transect_level, Stratum == levels(wide_transect_level$Stratum)[i])
  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
  if(sum(current_matrix) == 0){ # if this was a 0 algae transect
    transect_level_portfolio[i, 2] <- NA # set portfolio effects to 0
  }else{ # otherwise
    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae

    transect_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
  }

}

```

```{r CV, fig.height = 8, fig.width = 10}
# coefficient of variation
CV<-function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}

alpha_diversity$Cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
alpha_diversity$Stratum <- paste(alpha_diversity$Site, "-", alpha_diversity$Habitat, "-", alpha_diversity$Transect)
lm_synchrony_transect <- codyn::synchrony(df = transect_level,
                                   time.var = "Year",
                                   species.var = "Taxa",
                                   abundance.var = "Mean_cover",
                                   metric = "Loreau",
                                   replicate.var = "Stratum") # Loreau and Mazancourt synchrony
colnames(lm_synchrony_transect) = c("Stratum", "Loreau_synchrony") # rename cols for merge later

# gross_synchrony_transect <- codyn::synchrony(df = transect_level,
#                                    time.var = "Year",
#                                    species.var = "Taxa",
#                                    abundance.var = "Mean_cover",
#                                    metric = "Gross",
#                                    replicate.var = "Stratum") # lots of warnings in this one - examine further later
# colnames(gross_synchrony_transect) = c("Stratum", "Gross_synchrony") # rename cols for merge later

# stability_transect <- community_stability(df = transect_level,
#                                           time.var = "Year",
#                                           abundance.var = "Mean_cover",
#                                           replicate.var = "Stratum") # using codyn measure of stability
# NOTE: codyn community_stability() is just the inverse of CV
diversity_stability <- alpha_diversity %>%
  dplyr::group_by(Site, Habitat, Site_habitat, Transect, Stratum) %>%
  dplyr::summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Richness_cv = CV(Richness),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()),
            Shannon_cv = CV(Shannon),
            Shannon_stability = 1/CV(Shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
            Cover_mean = mean(Cover),
            Cover_se = sd(Cover)/sqrt(n()),
            Cover_cv = CV(Cover),
            Cover_stability = 1/CV(Cover)) # as above

diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "Stratum") # add in loreau and de Mazancourt synchrony
# diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "Stratum")
diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "Stratum") # add in portfolio effects
pairs(~ Cover_mean + Cover_cv + Cover_stability + 
        Shannon_mean + Shannon_cv + Shannon_stability + 
        Loreau_synchrony + Portfolio_effect, data = diversity_stability)
```

```{r basic diversity stability plots transect level, fig.height=8, fig.width=10}
(shannon_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Cover_stability, colour = Site, shape = Habitat), size = 8) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
#    stat_smooth(aes(x = Shannon_mean, y = Shannon_cv),method = "lm", se = T, size = 2) + 
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(shannon_cv_plot_2 = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# maybe some heteroskedasticity?
# just eyeballing but it looks like "stability" is highest in LTER 2 Fringe, which also has low diversity
# Functional form vs. portfolio effects?
# Either way looks sick as shit. Great clustering by habitat within site, and diversity trend is strong in LTER 3 4 5
# LTER2 the trend is reversed
# LTER 1 is mostly driven by habitat
# LTER 6 potentially a wash

```

```{r variance vs mean diversity plot, fig.height = 8, fig.width = 10}
(diversity_mean_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
# More diverse regions are also those that fluctuate the least in their diversity

(diversity_mean_cv_plot_2 = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Shannon_cv, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +

#  facet_wrap(~Site) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         ) # awesome relationship
```

```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Loreau_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# (gross_stability_plot = 
#   ggplot(data = diversity_stability) +
#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~Site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )
# 
(portfolio_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects
```

```{r diversity, portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_diversity_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# (gross_diversity_plot = 
#   ggplot(data = diversity_stability) +
#   geom_point(aes(x = Shannon_mean, y = Gross_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~Site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )

(portfolio_diversity_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Portfolio_effect, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
```

```{r model selection}
diversity_mod_universal <- glmmTMB(Cover_stability ~ Loreau_synchrony + Shannon_mean*Site*Habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
diversity_dredge <- dredge(diversity_mod_universal)
model.sel(diversity_dredge)

# top model is as follows (by a decent margin):
diversity_top <- glmmTMB(Cover_stability ~ Loreau_synchrony + Shannon_mean*Site + Site*Habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
summary(diversity_top)
Anova(diversity_top)
```

```{r general cover diversity, fig.height = 30, fig.width = 30}
(diversity_cover_plot = 
  ggplot(data = alpha_diversity) +
  geom_point(aes(x = Cover, y = Shannon, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Year) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         ) # ABORT ABORT ABORT
```

```{r diversity stability quadrat level setup, fig.height = 20, fig.width=20}
data_sub$Stratum <- as.factor(paste(data_sub$Site, "-", data_sub$Habitat, "-", data_sub$Transect))
data_sub$Location <- as.factor(data_sub$Location)

quadrat_portfolio_columns <- c("Location", "Portfolio_effect") # names of columns for dataframe
quadrat_level_portfolio <- 
  data.frame(matrix(nrow = length(unique(data_sub$Location)), 
                    ncol = length(quadrat_portfolio_columns))
             )
colnames(quadrat_level_portfolio) <- quadrat_portfolio_columns # change column 
quadrat_level_portfolio$Location <- unique(data_sub$Location) # populate

for(i in 1:length(unique(unique(data_sub$Location))) ){ # for each  quadrat
  # create a new dataframe for current quadrat
  newData <- subset(data, Location == levels(data_sub$Location)[i])
  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
  if(sum(current_matrix) == 0 | # if this was a 0 algae quadrat
     # or if there was only 1 column after removing all irrelevant algae
     is.null(ncol(current_matrix[, which(colSums(current_matrix) != 0)]))){ 
    quadrat_level_portfolio[i, 2] <- NA # set portfolio effects to 0
  }else{ # otherwise
    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae

    quadrat_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
  }

}

# sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble

hist(quadrat_level_portfolio$Portfolio_effect) # some crazy outliers due to low richness

alpha_diversity_quad$Cover <- rowSums(data[ , -which(names(data) %in% meta_cols)])/100
alpha_diversity_quad_sub <- alpha_diversity_quad[!alpha_diversity_quad$Location %in% problem_children, ]

# alpha_diversity_quad_sub$Cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
alpha_diversity_quad_sub$Stratum <- 
  paste(alpha_diversity_quad_sub$Site, "-", alpha_diversity_quad_sub$Habitat, "-", alpha_diversity_quad_sub$Transect)
lm_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
                                   time.var = "Year",
                                   species.var = "Taxa",
                                   abundance.var = "Percent_cover",
                                   metric = "Loreau",
                                   replicate.var = "Location") # Loreau and Mazancourt synchrony
colnames(lm_synchrony_quadrat) = c("Location", "Loreau_synchrony") # rename cols for merge later

## gross_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
##                                    time.var = "Year",
##                                    species.var = "Taxa",
##                                    abundance.var = "Percent_cover",
##                                    metric = "Gross",
##                                    replicate.var = "Location") # lots of warnings in this one - examine further later
## colnames(gross_synchrony_transect) = c("Location", "Gross_synchrony") # rename cols for merge later

variance_ratio_quadrat <- variance_ratio(df = long_data_sub,
                                   time.var = "Year",
                                   species.var = "Taxa",
                                   abundance.var = "Percent_cover",
                                   average.replicates = F,
                                   bootnumber = 1,
                                   replicate.var = "Location")

diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
  dplyr::group_by(Location, Site, Habitat, Site_habitat, Transect, Stratum) %>%
  dplyr::summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Richness_cv = CV(Richness),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()),
            Shannon_cv = CV(Shannon),
            Shannon_stability = 1/CV(Shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
            Cover_mean = mean(Cover),
            Cover_se = sd(Cover)/sqrt(n()),
            Cover_cv = CV(Cover),
            Cover_stability = 1/CV(Cover)) # as above

diversity_stability_quadrat <- 
  merge(diversity_stability_quadrat, lm_synchrony_quadrat, by = "Location") # add in loreau and de Mazancourt synchrony
diversity_stability_quadrat <- merge(diversity_stability_quadrat, gross_synchrony_quadrat, by = "Location")
diversity_stability_quadrat <- merge(diversity_stability_quadrat, quadrat_level_portfolio, by = "Location") # add in portfolio effects
diversity_stability_quadrat <- merge(diversity_stability_quadrat, variance_ratio_quadrat, by = "Location")
pairs(~ Cover_mean + Cover_cv + Cover_stability + 
        Shannon_mean + Shannon_cv + Shannon_stability + 
        Loreau_synchrony +  Portfolio_effect + VR, data = diversity_stability_quadrat)

```

```{r cover and cv, fig.height = 8, fig.width = 10}
(cover_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Cover_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(cover_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Cover_mean, y = Shannon_mean, colour = Site, shape = Habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

```

```{r basic diversity stability plots quadrat level, fig.height=8, fig.width=10}
(shannon_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(shannon_cv_plot_2 = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Shannon_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Habitat) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# LTER 2 obscures a lot of the trends

```

```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_stability_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# (gross_stability_plot_quad = 
#   ggplot(data = diversity_stability_quadrat) +
#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~Site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )

(portfolio_stability_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Portfolio_effect, y = Cover_cv, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
 # facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         ) # almost a perfect line, which makes sense given that CV is a component of portfolio effects

```

```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
(loreau_diversity_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Shannon_mean, y = Loreau_synchrony, colour = Site, shape = Habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(portfolio_diversity_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Shannon_mean, y = Portfolio_effect, colour = Site, shape = Habitat), size = 4, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
```

```{r diversity models quad level}
diversity_mod_universal <- glmmTMB(Cover_cv ~Shannon_mean*Site + Shannon_mean*Habitat + Site*Habitat + (1|Transect/Location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
diversity_dredge <- dredge(diversity_mod_universal)
model.sel(diversity_dredge) # best model is as specified
summary(diversity_mod_universal) # low variance in random effects
Anova(diversity_mod_universal) # LR Tests are kinda useless
r.squaredGLMM(diversity_mod_universal) # 0.653, 0.655

synchrony_mod_universal <- glmmTMB(Cover_cv ~Loreau_synchrony*Site + Loreau_synchrony*Habitat + Site*Habitat + (1|Transect/Location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
synchrony_dredge <- dredge(synchrony_mod_universal)
model.sel(synchrony_dredge) # best model is as specified
summary(synchrony_mod_universal)
r.squaredGLMM(synchrony_mod_universal) # 0.800, 0.802

diversity_stability_quadrat$Loreau_synchrony_trans <- sv_trans(diversity_stability_quadrat$Loreau_synchrony, diversity_stability_quadrat) # for modelling on beta
synchrony_diversity_universal <- glmmTMB(Loreau_synchrony_trans ~Shannon_mean*Site + Shannon_mean*Habitat + Site*Habitat + (1|Transect/Location), family = beta_family(), data = diversity_stability_quadrat, na.action = na.pass)
model.sel(dredge(synchrony_diversity_universal)) # best model as specified
summary(synchrony_diversity_universal)
Anova(synchrony_diversity_universal)
r.squaredGLMM(synchrony_diversity_universal) # Model's distribution-specific variance is negative. Results are not reliable.

```

```{r nmds, fig.height = 15, fig.width = 15}
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- paste(data$Year, data$Site_habitat, data$Transect)
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
data_subset$Stratum <- paste(data_subset$Year, "-", data_subset$Site, "-", data_subset$Habitat )

# identifying species w/ > 10% cover for plotting later
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 0.1))
abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
abundant_species_names <- colnames(abundant_species_matrix)

# My computer can't run this at the quadrat level
set.seed(1032) 
 ord2 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 2) 
# ord2 # stress = 0.22
# stressplot(ord2)

set.seed(1032)
ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
ord3 # 0.166
stressplot(ord3)

# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation

# plot(pro, cex=1.5)

site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
# add back in metadata
site.scores$Year <- as.factor(data_subset$Year)
site.scores$Site <- as.factor(data_subset$Site)
site.scores$Habitat <- as.factor(data_subset$Habitat)
site.scores$Site_habitat <- as.factor(data_subset$Site_habitat)
site.scores$Transect <- as.factor(data_subset$Transect)
site.scores$Stratum <- as.factor(data_subset$Stratum)

cent <- aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2

# Need to split stratum up for plotting later
cent<- 
  cent %>%
    separate(Stratum, into = c("Year", "Site", "Habitat"), sep = "-")
cent$Site_habitat <- paste(cent$Site, cent$Habitat)
# Add back in stratum column
cent$Stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ Stratum, data = site.scores, FUN = mean)$Stratum
# A lot of extra spaces

cent <- cent %>%
  mutate(across(everything(), str_trim))
cent$Year <- as.factor(cent$Year)
cent$Habitat <- as.factor(cent$Habitat)
cent$Site <- as.factor(cent$Site)
cent$Site_habitat <- as.factor(cent$Site_habitat)
cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
cent$NMDS2 <- as.numeric(cent$NMDS2)
species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
species.scores$Species <- rownames(species.scores)

# subset only the species w/ > 10% cover
abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]

ggplot() + 
  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = Habitat, colour = Site),
             alpha = 0.25, size=2) + # add the point markers
  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels

  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = Site_habitat, colour = Site),
            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17'))) +
  scale_linetype_manual(values = habitat_linetypes) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        strip.text.x = element_text(size = 25),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(1.5)), # increase legend text
        legend.title = element_text(size = rel(1.5)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) + # add box around legend
  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
                                     title.hjust =0.5, nrow = 2),
         shape=guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5,
                            nrow = 2),
         path = guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5)
         )

# looks like stochasticity increases from outer 17 --> Fringing
# Even in lagoon though it looks like communities return to initial state... except for North Shore LTER sites
# ways to test stochasticity vs. determinism?
# seems like deterministic trajectories, although we're not starting from a clean slate
# accounting for founder effects?

```

```{r beta dispersion, fig.height = 15, fig.width=20}
# Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
 bray_matrix <- vegdist(species_matrix, method = "bray")
# year_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Year)
# year_beta_disper #

# plot of dispersion test
# plot(year_beta_disper, sub = "") # lol
# boxplot(year_beta_disper, main = "Year", xlab = "") 
# 
# site_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Site)
# plot(site_beta_disper, sub = "") # site five stands out
# boxplot(site_beta_disper, main = "Site", xlab = "") 
# 
# 
# habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Habitat)
# plot(habitat_beta_disper, sub = "") 
# boxplot(habitat_beta_disper, main = "Habitat", xlab = "") # greatest variance in fringing it seems (as expected)
# 
# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Site_habitat)
# plot(site_habitat_beta_disper, sub = "") 
# boxplot(site_habitat_beta_disper, main = "Site_habitat", xlab = "") 
# 
stratum_beta_disper <- vegan::betadisper(bray_matrix, data_subset$Stratum)
# plot(site_habitat_beta_disper, sub = "") 
boxplot(stratum_beta_disper, main = "Site_habitat", xlab = "")
stratum_distances <- as.data.frame(stratum_beta_disper$distances)
colnames(stratum_distances)[1] <- "Distance"
stratum_distances$Stratum <- data_subset$Stratum
stratum_distances$Transect <- data_subset$Transect
stratum_distances$Year <- data_subset$Year
stratum_distances$Site <- data_subset$Site
stratum_distances$Habitat <- data_subset$Habitat
stratum_distances$Site_habitat <- data_subset$Site_habitat

beta_disper_means <- stratum_distances %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Distance_mean = mean(Distance), 
            Distance_se = sd(Distance)/sqrt(n()))

(Distance_plot <-
  ggplot(data = beta_disper_means) +
  geom_point(aes(x = Year, y = Distance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Distance_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~factor(Habitat, levels=c('Fringing','Backreef','Outer 10','Outer 17')))+
  labs(title = "Distance to centroid\n", y = "Distance\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
# beta dispersion increases as we move from forereef to lagoon, lagoon to shore

```