### 2024-07-30 julianna_algalMetric_wrangling.Rmd
Created a fresh Rmd for just the algal metric creation code

---
title: "Julianna's algal metrics wrangling code"
output: html_document
date: "2024-07-30"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
```

This script generates the CSV of algal metrics to regress other variables against

# Bring in the data

```{r}
# algal data
algae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
  mutate(island_side = case_when(site == "lter_1" ~ "north",
                                 site == "lter_2" ~ "north",
                                 site == "lter_3" ~ "east",
                                 site == "lter_4" ~ "east",
                                 site == "lter_5" ~ "west",
                                 site == "lter_6" ~ "west",
                                 TRUE ~ "UdINGLEbUTT"))
```