### 2024-07-30 julianna_richness_model.Rmd
Created the model markdown

---
title: "Richness model"
output: html_document
date: "2024-07-30"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(lme4) # mixed models
```

Make a model for algal richness through time

# Bring in the data

```{r load data}
cleaned_rich <- read_csv(here("data/algal_metrics_quadrat.csv")) # generated using my (Julianna's) algal metric code
```

# Make a model

```{r}
mRich <- glmer(,
               family = Poisson,
               data = cleaned_rich)
```