### 2024-08-10 comparing_datasets.Rmd
Markdown comparing datasets generated by J and N independently

---
title: "comparing noam and j datasets"
author: "Noam Altman-Kurosaki"
date: "2024-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

library(here) 
library(tidyverse)

```

This is a document for comparing the datasets N and J prepared

```{r quad level}
meta_cols <- c("year", "site", "island_side", "habitat")

j_quad <- read_csv(here("data", "j_algal_metrics_quadrat.csv"))
n_quad <- read_csv(here("data", "algae_diversity_quadrat.csv"))

quad_comp <- merge(j_quad, n_quad, by = meta_cols)

# cover - should be effectively 0 differnces
sum(quad_comp$cover - quad_comp$algal_perc_cov/100) # 7.297496e-13 --> some rounding issues, but fine

# richness
sum(quad_comp$richness - quad_comp$algal_richness) # 0

# shannon
sum(quad_comp$shannon - quad_comp$algal_Shann) # -2.218337e-12 --> rounding issues
```

```{r site level}

j_site <- read_csv(here("data", "j_algal_metrics_site.csv"))
n_site <- read_csv(here("data", "algae_diversity_site.csv"))

site_comp <- merge(j_site, n_site, by = meta_cols)

# cover - should be effectively 0 differnces
sum(site_comp$algae_cover_mean - site_comp$mean_alg_perc_cov/100) # 1.11e-15 --> some rounding issues, but fine

# richness
sum(site_comp$mean_alg_richness - site_comp$algae_richness_mean) # sameish here

# shannon
sum(site_comp$algae_shannon_mean - site_comp$mean_alg_Shann) # 0

```

```{r diversity stability}
j_stab <- read_csv(here("data", "j_algal_metrics_all_time.csv"))
n_stab <- read_csv(here("data", "algae_diversity_stability_metrics_site.csv"))

stab_comp <- merge(j_stab, n_stab, by = c("site", "island_side", "habitat"))

# cover 
sum(stab_comp$algae_cover_mean - stab_comp$mean_alg_cov/100) # -3.191891e-15--> rounding issues

# richness
sum(stab_comp$algae_richness_mean - stab_comp$mean_alg_richness) # 5.107026e-15

# shannon
sum(stab_comp$algae_shannon_mean - stab_comp$mean_Shann) # 0

# CV
sum(stab_comp$algae_cover_cv - stab_comp$alg_cov_cv) # -9.15934e-15
```