### 2024-08-15 julianna_coral_macroalg_wrangling.Rmd
Start a cleaned script with just the coral/macroalgae wrangling

---
title: "Wrangling coral and macroalgae data to be joinable"
output: html_document
date: "2024-08-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
```

This script summarizes the coral and macroalgae subsets to be joinable with the full algae dataset