### 2024-08-16 julianna_figs_Aug29.Rmd
Making a clean script for Aug 29 figgies that only uses standardized dataset

---
title: "Juilanna's figures for August 29"
output: html_document
date: "2024-08-16"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load librarian (for repeatable package loading/installation)
if (!require("librarian")) install.packages("librarian")

# load packages
librarian::shelf(
  here, # relative file paths
  tidyverse, # data wrangling
  vegan # community/diversity analyses
                 )
```

# Read in data

Want data at the quadrat level, which is our level of replication

```{r load data}
quads <- read_csv(here("data"))
```