### 2024-08-20 R/Fish_ord.Rmd
committing fishy things :fish:

---
title: "Fish_Stuff"
author: "Callie Stephenson"
date: "2024-03-28"
output: html_document
---
```{r packages, message=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(here)
library(fishualize)
library(ggplot2)
library(vegan)
library(mclust)
library(gridExtra)
library(glue)
library(patchwork) 
library(scales)
```

```{r data, message=FALSE, echo=FALSE}
#algae <- read.csv(here("data", "algaeonly_cleaned.csv"))
fish_fun <- read.csv(here("data", "biomass_functional_summary.csv"))
fish <- read.csv(here("data", "biomass_herb_summary.csv"))
alpha <- read.csv(here("data", "algae_alpha_diversity.csv")) %>% 
  filter(Habitat != "Outer 17")
```

Noam Algae Code:
```{r}
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sacrophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

algae <- data[ , -which(names(data) %in% not_algae)]
```

```{r}
data <- read.csv(here::here("data", "algaeonly_cleaned.csv"), stringsAsFactors = F) # replacements don't work when strings are factors
meta_cols <- colnames(data[1:7]) 
species_cols <- colnames(data[8:77])

wide_site_level <- data %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  summarize(across(all_of(species_cols), mean, na.rm = TRUE))

wide_site_level_merge <- wide_site_level[wide_site_level$Habitat != "Outer 17", ]
```

```{r plot colors from Noam}
#copied from Noam
site_colours <- c("LTER 1" = "cornflowerblue",
               "LTER 2" = "darkblue",
               "LTER 3" = "yellowgreen",
               "LTER 4" = "forestgreen",
               "LTER 5" = "mediumorchid",
               "LTER 6" = "orange")
island_side_colours <- c("North" = "darkblue",
                         "East" = "mediumorchid",
                         "West" = "orange")
habitat_linetypes <- c("Fringing" = "solid",
                       "Backreef" = "longdash",
                       "Outer 10" = "twodash",
                       "Outer 17" = "dotted")
habitat_shapes <- c("Fringing" = 16,
                     "Backreef" = 17,
                     "Outer 10" = 1,
                     "Outer 17" = 2)
```

### Ordination of Algal Communities with Fish Community as Drivers:

Transform the data and remove rows that are 100% coral cover

```{r}
species.data.transform = wisconsin(sqrt(wide_site_level_merge[c(5:74)]))
non_zero_rows <- rowSums(species.data.transform == 0) != ncol(species.data.transform)
species.data.use <- species.data.transform[non_zero_rows, ]
```

```{r}
pcoa = capscale(species.data.use ~ 1, dist = "bray")
pcoa
summary <- summary(pcoa)
summary$cont
```

```{r}
par(mfrow = c(1,2))
plot(pcoa, type = 'n', xlim = c(-1.5, 1.5))
text(pcoa, "species", col = 'blue', cex = 0.5)
plot(pcoa, type = 'n', xlim = c(-1.5, 1.5))
#text(pcoa, "location", col = 'blue', cex = 0.5)
points(pcoa, col = 'black', bg = 'grey', pch = 21)
```

```{r}
# Step 1: Ensure the 5th column is a vector
colors <- as.factor(wide_site_level_merge[non_zero_rows, 4][[1]])

# Step 2: Create the color palette based on the factor levels
color_palette <- rainbow(length(levels(colors)))

# Step 3: Plot the PCoA with colored points
par(mfrow = c(1,3))

# First plot: species plot (same as before)
plot(pcoa, type = 'n', xlim = c(-1.5, 1.5))
text(pcoa, "species", col = 'blue', cex = 1)

# Second plot: location plot with colored points
plot(pcoa, type = 'n', xlim = c(-1.5, 1.5))
points(pcoa, bg = color_palette[colors], pch = 21)

# Optionally, add a legend
plot(1, type = 'n', xlab = '', ylab = '', xaxt = 'n', yaxt = 'n')
legend("center", legend = levels(colors), col = color_palette, pch = 21, pt.bg = color_palette)
```

```{r}
fish_fun1 <- fish_fun
fish_fun1[is.na(fish_fun1)] <- 0
fish_fun1 <- fish_fun1 %>% 
  mutate(Site = paste0("LTER ",Site))
merged <- left_join(wide_site_level_merge, fish_fun1, by = join_by(Year, Site, Habitat))
merged_clean <- merged[complete.cases(merged[c("Browser", "Brusher", "Cropper", "Concealed.Cropper", "Excavator", "Scraper")]), ]

species.data.transform = wisconsin(sqrt(merged_clean[c(6:75)]))
non_zero_rows <- rowSums(species.data.transform == 0) != ncol(species.data.transform)
species.data.use <- species.data.transform[non_zero_rows, ]
merged_clean_filtered <- merged_clean[non_zero_rows, ]
```

```{r}
cont = capscale(species.data.use ~ Browser + Brusher + Cropper + Concealed.Cropper + Excavator + Scraper, dist = "bray", data = merged_clean_filtered)
```

```{r}
plot(cont)
```

```{r}
summary(cont)$cont
```

```{r}
par(mfrow = c(1, 3))  # 1 row, 3 columns

# Plot 1: Species driving the community composition
plot(cont, display = "species", main = "Species Driving Community Composition", type = "n")
text(cont, display = "species", col = "blue", cex = 1)

# Plot 2: Communities (sites) plotted
plot(cont, display = "sites", main = "Communities", type = "n")
points(cont, display = "sites", col = color_palette[colors], pch = 21)  # Sites colored by the same scheme

# Plot 3: Constrained ordination axes (environmental variables)
plot(cont, display = "bp", main = "Constrained Ordination Axes", cex = 1)
```

```{r}
anova(cont)
anova(cont, by = "margin")
```

# Mantel 
I have no idea what I'm doing here
```{r}
library(vegan)
dist.algae <- dist(scale(x = merged_clean_filtered[,c(5:74)], scale = T, center = T), method = "euclidean")
dist.fish <- dist(scale(x = merged_clean_filtered[,c(75:88)], scale = T, center = T), method = "euclidean")
response_by_explan <- mantel(dist.algae, dist.fish, method = "spearman", permutations = 900, na.rm = TRUE)
response_by_explan
```

### Part 3: Model-Based Clustering