### 2024-08-20 noam_beta_diversity.Rmd
New beta diversity markdown for ordinations and dissimilarity. NOT FINISHED YET.

---
title: "noam_beta_diversity"
author: "Noam Altman-Kurosaki"
date: "2024-08-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
# library(ecofolio) # possible other method for examining portfolio effects 
```

```{r import and format data}
# starting with all data now because of 0 algae transects/quadrats
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sacrophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

data <- data[ , -which(names(data) %in% not_algae)]

# adding island side - adapted from Julianna's code
data <- data %>% mutate(island_side = case_when(site == "lter_1" ~ "north",
                                 site == "lter_2" ~ "north",
                                 site == "lter_3" ~ "east",
                                 site == "lter_4" ~ "east",
                                 site == "lter_5" ~ "west",
                                 site == "lter_6" ~ "west"))
data$island_side_habitat <- paste(data$island_side, data$habitat)

str(data)

meta_cols <- c(colnames(data[1:7] ), "island_side", "island_side_habitat", "stratum")
species_cols <- colnames(data)[!colnames(data) %in% meta_cols]

meta_df <- data[ , which(names(data) %in% meta_cols)]

site_colours <- c("lter_1" = "cornflowerblue",
                  "lter_2" = "darkblue",
                  "lter_3" = "yellowgreen",
                  "lter_4" = "forestgreen",
                  "lter_5" = "mediumorchid",
                  "lter_6" = "orange")
island_side_colours <- c("north" = "darkblue",
                         "east" = "mediumorchid",
                         "west" = "orange")
habitat_linetypes <- c("fringing" = "solid",
                       "backreef" = "longdash",
                       "outer_10" = "twodash",
                       "outer_17" = "dotted")
habitat_shapes <- c("fringing" = 16,
                     "backreef" = 17,
                     "outer_10" = 1,
                     "outer_17" = 2)

#pivot long 
long_data <- pivot_longer(
  data = data,
  cols = 8:77,
  names_to = "taxa",
  values_to = "percent_cover")

 transect_level <- long_data %>%
   dplyr::group_by(year, island_side, site, habitat, transect, taxa) %>%
   dplyr::summarise(mean_cover = mean(percent_cover), 
             SE = sd(percent_cover)/sqrt(n()))

site_level <- long_data %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, taxa) %>%
  dplyr::summarise(mean_site_cover = mean(percent_cover), 
            site_se = sd(percent_cover)/sqrt(n()))

# subset based on habitat
fringing <- subset(data, habitat == "fringing")
backreef <- subset(data, habitat == "backreef")
outer10 <- subset(data, habitat == "outer_10")
outer17 <- subset(data, habitat == "outer_17")

```

```{r fringing bray curtis}
fringing_species_matrix <- as.matrix(fringing[ , -which(names(fringing) %in% meta_cols)])
fringing_species_matrix <- fringing_species_matrix/100 # convert from percent to proportion
rownames(fringing_species_matrix) <- paste(fringing$year,"-", fringing$location, sep = "")
fringing_rows_to_remove <- rowSums(fringing_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
fringing_species_matrix <- fringing_species_matrix[!fringing_rows_to_remove,] 
fringing_species_matrix <- asin(sqrt(fringing_species_matrix)) # arcsin squareroot transform
fringing_subset <- fringing[!fringing_rows_to_remove,] # make sure rows are also removed from dataframe
fringing_subset$stratum <- paste(fringing_subset$year,"-", fringing_subset$location, sep = "") # for merging with bray matrix later
brayfringing <- vegdist(fringing_species_matrix, "bray") # create bray dissimilarity matrix
# View(brayfringing)
fringing_braymat <- as.matrix(brayfringing) # conver to matrix object
# convert to pairwise comparison
fringing_comp <- fringing_braymat[lower.tri(fringing_braymat)] # extract lower triangle
fringing_row_indices <- row(fringing_braymat)[lower.tri(fringing_braymat)] # extract row indices for first quad in comparison
fringing_col_indices <- col(fringing_braymat)[lower.tri(fringing_braymat)] # extract col indices = second quad in comp
fringing_loc1 <- rownames(fringing_braymat)[fringing_row_indices] # populate appropriate vector
fringing_loc2 <- colnames(fringing_braymat)[fringing_col_indices] # populate appropriate vector
fringing_pairwise_df <- data.frame(loc1 = fringing_loc1, loc2 = fringing_loc2, dissimilarity = fringing_comp)
fringing_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", fringing_pairwise_df$loc1)) # extract timepoint from first quad in comp
fringing_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", fringing_pairwise_df$loc2)) # repeat for second quad
fringing_pairwise_df$time <- abs(fringing_pairwise_df$year2 - fringing_pairwise_df$year1) # extract length of differences
fringing_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", fringing_pairwise_df$loc1) # extract quadrat id from first quad in comp
fringing_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", fringing_pairwise_df$loc2) # repeat for second quad

```

```{r backreef bray curtis}
backreef_species_matrix <- as.matrix(backreef[ , -which(names(backreef) %in% meta_cols)])
backreef_species_matrix <- backreef_species_matrix/100 # convert from percent to proportion
rownames(backreef_species_matrix) <- paste(backreef$year,"-", backreef$location, sep = "")
backreef_rows_to_remove <- rowSums(backreef_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
backreef_species_matrix <- backreef_species_matrix[!backreef_rows_to_remove,] 
backreef_species_matrix <- asin(sqrt(backreef_species_matrix)) # arcsin squareroot transform
backreef_subset <- backreef[!backreef_rows_to_remove,] # make sure rows are also removed from dataframe
backreef_subset$stratum <- paste(backreef_subset$year,"-", backreef_subset$location, sep = "") # for merging with bray matrix later
braybackreef <- vegdist(backreef_species_matrix, "bray") # create bray dissimilarity matrix
# View(braybackreef)
backreef_braymat <- as.matrix(braybackreef) # conver to matrix object
# convert to pairwise comparison
backreef_comp <- backreef_braymat[lower.tri(backreef_braymat)] # extract lower triangle
backreef_row_indices <- row(backreef_braymat)[lower.tri(backreef_braymat)] # extract row indices for first quad in comparison
backreef_col_indices <- col(backreef_braymat)[lower.tri(backreef_braymat)] # extract col indices = second quad in comp
backreef_loc1 <- rownames(backreef_braymat)[backreef_row_indices] # populate appropriate vector
backreef_loc2 <- colnames(backreef_braymat)[backreef_col_indices] # populate appropriate vector
backreef_pairwise_df <- data.frame(loc1 = backreef_loc1, loc2 = backreef_loc2, dissimilarity = backreef_comp)
backreef_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", backreef_pairwise_df$loc1)) # extract timepoint from first quad in comp
backreef_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", backreef_pairwise_df$loc2)) # repeat for second quad
backreef_pairwise_df$time <- abs(backreef_pairwise_df$year2 - backreef_pairwise_df$year1) # extract length of differences
backreef_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", backreef_pairwise_df$loc1) # extract quadrat id from first quad in comp
backreef_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", backreef_pairwise_df$loc2) # repeat for second quad

```

```{r outer10 bray curtis}
outer10_species_matrix <- as.matrix(outer10[ , -which(names(outer10) %in% meta_cols)])
outer10_species_matrix <- outer10_species_matrix/100 # convert from percent to proportion
rownames(outer10_species_matrix) <- paste(outer10$year,"-", outer10$location, sep = "")
outer10_rows_to_remove <- rowSums(outer10_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
outer10_species_matrix <- outer10_species_matrix[!outer10_rows_to_remove,] 
outer10_species_matrix <- asin(sqrt(outer10_species_matrix)) # arcsin squareroot transform
outer10_subset <- outer10[!outer10_rows_to_remove,] # make sure rows are also removed from dataframe
outer10_subset$stratum <- paste(outer10_subset$year,"-", outer10_subset$location, sep = "") # for merging with bray matrix later
brayouter10 <- vegdist(outer10_species_matrix, "bray") # create bray dissimilarity matrix
# View(brayouter10)
outer10_braymat <- as.matrix(brayouter10) # conver to matrix object
# convert to pairwise comparison
outer10_comp <- outer10_braymat[lower.tri(outer10_braymat)] # extract lower triangle
outer10_row_indices <- row(outer10_braymat)[lower.tri(outer10_braymat)] # extract row indices for first quad in comparison
outer10_col_indices <- col(outer10_braymat)[lower.tri(outer10_braymat)] # extract col indices = second quad in comp
outer10_loc1 <- rownames(outer10_braymat)[outer10_row_indices] # populate appropriate vector
outer10_loc2 <- colnames(outer10_braymat)[outer10_col_indices] # populate appropriate vector
outer10_pairwise_df <- data.frame(loc1 = outer10_loc1, loc2 = outer10_loc2, dissimilarity = outer10_comp)
outer10_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer10_pairwise_df$loc1)) # extract timepoint from first quad in comp
outer10_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer10_pairwise_df$loc2)) # repeat for second quad
outer10_pairwise_df$time <- abs(outer10_pairwise_df$year2 - outer10_pairwise_df$year1) # extract length of differences
outer10_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", outer10_pairwise_df$loc1) # extract quadrat id from first quad in comp
outer10_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", outer10_pairwise_df$loc2) # repeat for second quad

```

```{r outer17 bray curtis}
outer17_species_matrix <- as.matrix(outer17[ , -which(names(outer17) %in% meta_cols)])
outer17_species_matrix <- outer17_species_matrix/100 # convert from percent to proportion
rownames(outer17_species_matrix) <- paste(outer17$year,"-", outer17$location, sep = "")
outer17_rows_to_remove <- rowSums(outer17_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
outer17_species_matrix <- outer17_species_matrix[!outer17_rows_to_remove,] 
outer17_species_matrix <- asin(sqrt(outer17_species_matrix)) # arcsin squareroot transform
outer17_subset <- outer17[!outer17_rows_to_remove,] # make sure rows are also removed from dataframe
outer17_subset$stratum <- paste(outer17_subset$year,"-", outer17_subset$location, sep = "") # for merging with bray matrix later
brayouter17 <- vegdist(outer17_species_matrix, "bray") # create bray dissimilarity matrix
# View(brayouter17)
outer17_braymat <- as.matrix(brayouter17) # conver to matrix object
# convert to pairwise comparison
outer17_comp <- outer17_braymat[lower.tri(outer17_braymat)] # extract lower triangle
outer17_row_indices <- row(outer17_braymat)[lower.tri(outer17_braymat)] # extract row indices for first quad in comparison
outer17_col_indices <- col(outer17_braymat)[lower.tri(outer17_braymat)] # extract col indices = second quad in comp
outer17_loc1 <- rownames(outer17_braymat)[outer17_row_indices] # populate appropriate vector
outer17_loc2 <- colnames(outer17_braymat)[outer17_col_indices] # populate appropriate vector
outer17_pairwise_df <- data.frame(loc1 = outer17_loc1, loc2 = outer17_loc2, dissimilarity = outer17_comp)
outer17_pairwise_df$year1 <- as.numeric(sub("(.*)-.*", "\\1", outer17_pairwise_df$loc1)) # extract timepoint from first quad in comp
outer17_pairwise_df$year2 <- as.numeric(sub("(.*)-.*", "\\1", outer17_pairwise_df$loc2)) # repeat for second quad
outer17_pairwise_df$time <- abs(outer17_pairwise_df$year2 - outer17_pairwise_df$year1) # extract length of differences
outer17_pairwise_df$location_1 <- sub(".*-(.*)", "\\1", outer17_pairwise_df$loc1) # extract quadrat id from first quad in comp
outer17_pairwise_df$location_2 <- sub(".*-(.*)", "\\1", outer17_pairwise_df$loc2) # repeat for second quad

```

```{r bray dissimilarity over time, fig.height=15, fig.width=15}
# for the purposes of this first set of analyses, I'm interested in dissimilarity of a given quad over time
# not necessarily interested in how quads compare to each other over time/at each timepoint
fringing_pairwise_subset <- fringing_pairwise_df[fringing_pairwise_df$location_1 == fringing_pairwise_df$location_2,]
backreef_pairwise_subset <- backreef_pairwise_df[backreef_pairwise_df$location_1 == backreef_pairwise_df$location_2,]
outer10_pairwise_subset <- outer10_pairwise_df[outer10_pairwise_df$location_1 == outer10_pairwise_df$location_2,]
outer17_pairwise_subset <- outer17_pairwise_df[outer17_pairwise_df$location_1 == outer17_pairwise_df$location_2,]

comb_subset <- rbind(fringing_pairwise_subset,
                     backreef_pairwise_subset,
                     outer10_pairwise_subset,
                     outer17_pairwise_subset)
# location_1 and location_2 columns are all the same now
# create new col just called "location" for merge with other metadata
comb_subset$location <- comb_subset$location_1
bray_time_series <- merge(meta_df, comb_subset, by = "location")

# model

# some quadrats have dissimilarity = 0 or 1

# Smithson verkuilen transformation for beta-regression
sv_trans = function(prop, N, s = 0.000005){
  (prop*(nrow(N) - 1) + s)/nrow(N)
  # where prop is the proportional value you're transforming, 
  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
  # and s is a small offset 
}

# transformed dissimilarities
bray_time_series$dissimilarity_trans <- sv_trans(bray_time_series$dissimilarity, bray_time_series)

# bray_mod_universal <- glmmTMB(dissimilarity_trans ~ time*habitat*island_side + (1 | island_side/site/transect/location),
#                               na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series) #
# summary(bray_mod_universal) # a fair bit of variance w/i quads/transects
# 
# bray_mod_reduced <- glmmTMB(dissimilarity_trans ~ time*habitat + time*island_side + (1 | island_side/site/transect/location),
#                               na.action = na.pass, family = beta_family(link = "logit"), data = bray_time_series)

```

```{r dissimilarity over time plot, fig.height=15, fig.width=15}
 ggplot(bray_time_series,
        aes(x = time, y = dissimilarity, colour = island_side) )+
   facet_wrap(~habitat) + 
   geom_point() +
   scale_colour_manual(values = island_side_colours) +
   stat_smooth(se = T, linewidth = 2) +
   labs(y = "Dissimilarity\n") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") # make legend horizontal
         )

```

```{r fringing nmds set up}
# identifying species w/ > 10% cover for plotting later

#### FRINGING
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
# fringing_columns_to_keep <- apply(fringing_species_matrix, 2, function(x) any(x >= 0.1))
# fringing_abundant_species_matrix <- as.data.frame(fringing_species_matrix)[ , fringing_columns_to_keep]
# fringing_abundant_species_names <- colnames(fringing_abundant_species_matrix)
# 
# fringing_subset_meta <- fringing_subset[ , which(names(fringing_subset) %in% meta_cols)]
# 
# # for mean centroid:
# fringing_subset_meta$island_side_year <- paste(fringing_subset_meta$year, "-", fringing_subset_meta$island_side_habitat, sep = "")
# 
# # takes several hours to run - do so at your own peril
# set.seed(1032) 
# # ord3 <- metaMDS(fringing_braymat, dist	= "bray", k = 3, trymax = 100) 
# fringing_ord3 # stress = 0.018
# 
# 
# sppscores(fringing_ord3) <- fringing_species_matrix # add back in species scores
# fringing.species.scores <- as.data.frame(scores(fringing_ord3, "species")) # some NaN for absent species
# fringing.species.scores <- na.omit(fringing.species.scores)
# fringing.species.scores$species <- rownames(fringing.species.scores) 
# 
# 
# fringing.site.scores <- as.data.frame(vegan::scores(fringing_ord3, "sites"))  # extract the site scores and convert to a data.frame
# # add back in metadata
# fringing.site.scores$stratum <- rownames(fringing.site.scores)
# fringing.site.scores <- merge(fringing_subset_meta, fringing.site.scores, by = "stratum")
# 
# 
# 
# 
# write_csv(site.scores, here("data", "fringing_site_scores.csv"))
# write_csv(species.scores, here("data", "fringing_species_scores.csv"))

fringing.species.scores <- read.csv(here::here("data", "fringing_species_scores.csv"), stringsAsFactors = F) 
fringing.site.scores <- read.csv(here::here("data", "fringing_site_scores.csv"), stringsAsFactors = F) 
# fringing.cent <- read.csv(here::here("data", "fringing_mean_centroid.csv"), stringsAsFactors = F)

fringing.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = fringing.site.scores, FUN = mean) 
# fringing.cent.meta <- fringing_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# fringing.cent <- merge(fringing.cent, fringing_subset_meta[,-which(names(fringing_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
fringing.abundant.species.scores <- fringing.species.scores[fringing.species.scores$species %in% fringing_abundant_species_names,]

#############
# BACKREEF #
############
backreef_columns_to_keep <- apply(backreef_species_matrix, 2, function(x) any(x >= 0.1))
backreef_abundant_species_matrix <- as.data.frame(backreef_species_matrix)[ , backreef_columns_to_keep]
backreef_abundant_species_names <- colnames(backreef_abundant_species_matrix)

backreef_subset_meta <- backreef_subset[ , which(names(backreef_subset) %in% meta_cols)]

# for mean centroid:
backreef_subset_meta$island_side_year <- paste(backreef_subset_meta$year, "-", backreef_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(backreef_braymat, dist	= "bray", k = 3, trymax = 100) 
backreef_ord3 # stress = 0.018

sppscores(backreef_ord3) <- backreef_species_matrix # add back in species scores
backreef.species.scores <- as.data.frame(scores(backreef_ord3, "species")) # some NaN for absent species
backreef.species.scores <- na.omit(backreef.species.scores)
backreef.species.scores$species <- rownames(backreef.species.scores)

backreef.site.scores <- as.data.frame(vegan::scores(backreef_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
backreef.site.scores$stratum <- rownames(backreef.site.scores)
backreef.site.scores <- merge(backreef_subset_meta, backreef.site.scores, by = "stratum")

write_csv(site.scores, here("data", "backreef_site_scores.csv"))
write_csv(species.scores, here("data", "backreef_species_scores.csv"))

backreef.species.scores <- read.csv(here::here("data", "backreef_species_scores.csv"), stringsAsFactors = F) 
backreef.site.scores <- read.csv(here::here("data", "backreef_site_scores.csv"), stringsAsFactors = F) 
# backreef.cent <- read.csv(here::here("data", "backreef_mean_centroid.csv"), stringsAsFactors = F)

backreef.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = backreef.site.scores, FUN = mean) 
# backreef.cent.meta <- backreef_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# backreef.cent <- merge(backreef.cent, backreef_subset_meta[,-which(names(backreef_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
backreef.abundant.species.scores <- backreef.species.scores[backreef.species.scores$species %in% backreef_abundant_species_names,]

#######
# OUTER10
#######

outer10_columns_to_keep <- apply(outer10_species_matrix, 2, function(x) any(x >= 0.1))
outer10_abundant_species_matrix <- as.data.frame(outer10_species_matrix)[ , outer10_columns_to_keep]
outer10_abundant_species_names <- colnames(outer10_abundant_species_matrix)

outer10_subset_meta <- outer10_subset[ , which(names(outer10_subset) %in% meta_cols)]

# for mean centroid:
outer10_subset_meta$island_side_year <- paste(outer10_subset_meta$year, "-", outer10_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(outer10_braymat, dist	= "bray", k = 3, trymax = 100) 
outer10_ord3 # stress = 0.018

sppscores(outer10_ord3) <- outer10_species_matrix # add back in species scores
outer10.species.scores <- as.data.frame(scores(outer10_ord3, "species")) # some NaN for absent species
outer10.species.scores <- na.omit(outer10.species.scores)
outer10.species.scores$species <- rownames(outer10.species.scores)

outer10.site.scores <- as.data.frame(vegan::scores(outer10_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
outer10.site.scores$stratum <- rownames(outer10.site.scores)
outer10.site.scores <- merge(outer10_subset_meta, outer10.site.scores, by = "stratum")

write_csv(site.scores, here("data", "outer10_site_scores.csv"))
write_csv(species.scores, here("data", "outer10_species_scores.csv"))

outer10.species.scores <- read.csv(here::here("data", "outer10_species_scores.csv"), stringsAsFactors = F) 
outer10.site.scores <- read.csv(here::here("data", "outer10_site_scores.csv"), stringsAsFactors = F) 
# outer10.cent <- read.csv(here::here("data", "outer10_mean_centroid.csv"), stringsAsFactors = F)

outer10.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = outer10.site.scores, FUN = mean) 
# outer10.cent.meta <- outer10_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# outer10.cent <- merge(outer10.cent, outer10_subset_meta[,-which(names(outer10_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
outer10.abundant.species.scores <- outer10.species.scores[outer10.species.scores$species %in% outer10_abundant_species_names,]

#############
# OUTER 17 ##
#############

outer17_columns_to_keep <- apply(outer17_species_matrix, 2, function(x) any(x >= 0.1))
outer17_abundant_species_matrix <- as.data.frame(outer17_species_matrix)[ , outer17_columns_to_keep]
outer17_abundant_species_names <- colnames(outer17_abundant_species_matrix)

outer17_subset_meta <- outer17_subset[ , which(names(outer17_subset) %in% meta_cols)]

# for mean centroid:
outer17_subset_meta$island_side_year <- paste(outer17_subset_meta$year, "-", outer17_subset_meta$island_side_habitat, sep = "")

# takes several hours to run - do so at your own peril
set.seed(1032) 
# ord3 <- metaMDS(outer17_braymat, dist	= "bray", k = 3, trymax = 100) 
outer17_ord3 # stress = 0.018

sppscores(outer17_ord3) <- outer17_species_matrix # add back in species scores
outer17.species.scores <- as.data.frame(scores(outer17_ord3, "species")) # some NaN for absent species
outer17.species.scores <- na.omit(outer17.species.scores)
outer17.species.scores$species <- rownames(outer17.species.scores)

outer17.site.scores <- as.data.frame(vegan::scores(outer17_ord3, "sites"))  # extract the site scores and convert to a data.frame
# add back in metadata
outer17.site.scores$stratum <- rownames(outer17.site.scores)
outer17.site.scores <- merge(outer17_subset_meta, outer17.site.scores, by = "stratum")

write_csv(site.scores, here("data", "outer17_site_scores.csv"))
write_csv(species.scores, here("data", "outer17_species_scores.csv"))

outer17.species.scores <- read.csv(here::here("data", "outer17_species_scores.csv"), stringsAsFactors = F) 
outer17.site.scores <- read.csv(here::here("data", "outer17_site_scores.csv"), stringsAsFactors = F) 
# outer17.cent <- read.csv(here::here("data", "outer17_mean_centroid.csv"), stringsAsFactors = F)

outer17.cent <- aggregate(cbind(NMDS1, NMDS2) ~ island_side_year, data = outer17.site.scores, FUN = mean) 
# outer17.cent.meta <- outer17_subset_meta %>%
#    dplyr::group_by(year, island_side, island_side_habitat, island_side_year, site, habitat, transect, quadrat) %>%
#    dplyr::summarise(year = unique(year), 
#                     island_side = unique(island_side),
#                     island_side_habitat = unique(island_side_habitat),
#                     island_side_year = unique(island_side_year),
#                     site = unique(site),
#                     habitat = unique(habitat),
#                     transect = unique(transect),
#                     quadrat = unique(quadrat))
# outer17.cent <- merge(outer17.cent, outer17_subset_meta[,-which(names(outer17_subset_meta) %in% c("stratum"))],  by = "island_side_year")

# subset only the species w/ > 10% cover
outer17.abundant.species.scores <- outer17.species.scores[outer17.species.scores$species %in% outer17_abundant_species_names,]
```

```{r plot fringing}
ggplot() + 
  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, colour = island_side),
             alpha = 0.25, size=2) + # add the point markers
  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = species), size = 4) +  # add the species labels

  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = island_side_habitat, colour = island_side),
            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
#  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
  scale_linetype_manual(values = habitat_linetypes) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        strip.text.x = element_text(size = 25),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(1.5)), # increase legend text
        legend.title = element_text(size = rel(1.5)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) + # add box around legend
  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
                                     title.hjust =0.5, nrow = 2),
         shape=guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5,
                            nrow = 2),
         path = guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5)
         )

```

```{r full bray curtis set up, eval = FALSE}
### I DON'T THINK ANY COMPUTER CAN RUN THIS
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- paste(data$year,"-", data$location, sep = "")
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
data_subset$stratum <- paste(data_subset$year,"-", data_subset$location, sep = "") # for merging with bray matrix later
brayfull <- vegdist(species_matrix, "bray") # create bray dissimilarity matrix
# View(brayfull)
braymat <- as.matrix(brayfull) # conver to matrix object
# convert to pairwise comparison
comp <- braymat[lower.tri(braymat)] # extract lower triangle
row_indices <- row(braymat)[lower.tri(braymat)] # extract row indices for first quad in comparison
col_indices <- col(braymat)[lower.tri(braymat)] # 
loc1 <- rownames(braymat)[row_indices]
loc2 <- colnames(braymat)[col_indices]
pairwise_df <- data.frame(loc1 = loc1, loc2 = loc2, dissimilarity = comp)
# pairwise_df$year1 <- sub("(.*)-.*", "\\1", pairwise_df$loc1)
year1 <- strsplit(pairwise_df$loc1, "-")[[1]]

```

```{r old nmds code, fig.height = 15, fig.width = 15, eval = FALSE}
species_matrix <- as.matrix(data[ , -which(names(data) %in% meta_cols)])
species_matrix <- species_matrix/100 # convert from percent to proportion
rownames(species_matrix) <- paste(data$year, data$site_habitat, data$transect)
rows_to_remove <- rowSums(species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
species_matrix <- species_matrix[!rows_to_remove,] 
data_subset <- data[!rows_to_remove,] # make sure rows are also removed from dataframe
data_subset$stratum <- paste(data_subset$year, "-", data_subset$site, "-", data_subset$habitat )
# identifying species w/ > 10% cover for plotting later
# NOTE: Used ChatGPT to generate these next few lines of code but it looked good so I went with it
columns_to_keep <- apply(species_matrix, 2, function(x) any(x >= 0.1))
abundant_species_matrix <- as.data.frame(species_matrix)[ , c(TRUE, columns_to_keep)]
abundant_species_names <- colnames(abundant_species_matrix)

# My computer can't run this at the quadrat level

# ord2 # stress = 0.22
# stressplot(ord2)

set.seed(1032)
ord3 <- metaMDS(species_matrix, dist	= "bray",	trymax	= 100,	k	= 3)
ord3 # 0.166
stressplot(ord3)

# pro <- procrustes(ord2, ord3) # visually compare w procrustes rotation

# plot(pro, cex=1.5)

site.scores <- as.data.frame(vegan::scores(ord3)$sites)  # extract the site scores and convert to a data.frame
# add back in metadata
site.scores$year <- as.factor(data_subset$year)
site.scores$site <- as.factor(data_subset$site)
site.scores$habitat <- as.factor(data_subset$habitat)
site.scores$site_habitat <- as.factor(data_subset$site_habitat)
site.scores$transect <- as.factor(data_subset$transect)
site.scores$stratum <- as.factor(data_subset$stratum)

cent <- aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean) # compute centroids for NMDS1 and NMDS2

# Need to split stratum up for plotting later
cent<- 
  cent %>%
    separate(stratum, into = c("year", "site", "habitat"), sep = "-")
cent$site_habitat <- paste(cent$site, cent$habitat)
# Add back in stratum column
cent$stratum <-  aggregate(cbind(NMDS1, NMDS2) ~ stratum, data = site.scores, FUN = mean)$stratum
# A lot of extra spaces

cent <- cent %>%
  mutate(across(everything(), str_trim))
cent$year <- as.factor(cent$year)
cent$habitat <- as.factor(cent$habitat)
cent$site <- as.factor(cent$site)
cent$site_habitat <- as.factor(cent$site_habitat)
cent$NMDS1 <- as.numeric(cent$NMDS1) # read in as characters for some reason
cent$NMDS2 <- as.numeric(cent$NMDS2)
species.scores <- as.data.frame(scores(ord3)$species)  # extract the species scores and convert to a data.frame
species.scores$Species <- rownames(species.scores)

# subset only the species w/ > 10% cover
abundant.species.scores <- species.scores[species.scores$Species %in% abundant_species_names,]

ggplot() + 
  geom_point(data= site.scores, aes(x=NMDS1,y=NMDS2, shape = habitat, colour = site),
             alpha = 0.25, size=2) + # add the point markers
  geom_text(data= abundant.species.scores,aes(x=NMDS1,y=NMDS2, label = Species), size = 4) +  # add the species labels

  geom_path(data = cent, aes(x = NMDS1, y = NMDS2, group = site_habitat, colour = site),
            arrow = arrow(length=unit(0.5,"cm"), ends="last", type = "closed"), size = 1) +
  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17'))) +
  scale_linetype_manual(values = habitat_linetypes) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        strip.text.x = element_text(size = 25),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(1.5)), # increase legend text
        legend.title = element_text(size = rel(1.5)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) + # add box around legend
  guides(colour=guide_legend(title.position="top", # make color legend title go above legend
                                     title.hjust =0.5, nrow = 2),
         shape=guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5,
                            nrow = 2),
         path = guide_legend(title.position="top",  # make shape legend title go above legend
                                     title.hjust =0.5)
         )

# looks like stochasticity increases from outer_17 --> fringing
# Even in lagoon though it looks like communities return to initial state... except for north Shore lter_sites
# ways to test stochasticity vs. determinism?
# seems like deterministic trajectories, although we're not starting from a clean slate
# accounting for founder effects?

```

```{r beta dispersion, fig.height = 15, fig.width=20}
# Want to plot mean distance to centroid, but not sure how to do that with so many factors/interactions...
 bray_matrix <- vegdist(species_matrix, method = "bray")
# year_beta_disper <- vegan::betadisper(bray_matrix, data_subset$year)
# year_beta_disper #

# plot of dispersion test
# plot(year_beta_disper, sub = "") # lol
# boxplot(year_beta_disper, main = "year", xlab = "") 
# 
# site_beta_disper <- vegan::betadisper(bray_matrix, data_subset$site)
# plot(site_beta_disper, sub = "") # site five stands out
# boxplot(site_beta_disper, main = "site", xlab = "") 
# 
# 
# habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$habitat)
# plot(habitat_beta_disper, sub = "") 
# boxplot(habitat_beta_disper, main = "habitat", xlab = "") # greatest variance in fringing it seems (as expected)
# 
# site_habitat_beta_disper <- vegan::betadisper(bray_matrix, data_subset$site_habitat)
# plot(site_habitat_beta_disper, sub = "") 
# boxplot(site_habitat_beta_disper, main = "site_habitat", xlab = "") 
# 
stratum_beta_disper <- vegan::betadisper(bray_matrix, data_subset$stratum)
# plot(site_habitat_beta_disper, sub = "") 
boxplot(stratum_beta_disper, main = "site_habitat", xlab = "")
stratum_distances <- as.data.frame(stratum_beta_disper$distances)
colnames(stratum_distances)[1] <- "Distance"
stratum_distances$stratum <- data_subset$stratum
stratum_distances$transect <- data_subset$transect
stratum_distances$year <- data_subset$year
stratum_distances$site <- data_subset$site
stratum_distances$habitat <- data_subset$habitat
stratum_distances$site_habitat <- data_subset$site_habitat

beta_disper_means <- stratum_distances %>%
  dplyr::group_by(year, site, habitat, site_habitat) %>%
  dplyr::summarise(Distance_mean = mean(Distance), 
            Distance_se = sd(Distance)/sqrt(n()))

(Distance_plot <-
  ggplot(data = beta_disper_means) +
  geom_point(aes(x = year, y = Distance_mean, colour = site, shape = habitat, group = site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = year, y = Distance_mean, colour = site, group = site_habitat), size = 1.2) +
  geom_errorbar(aes(x = year, ymin = Distance_mean - Distance_se , ymax = Distance_mean + Distance_se, colour = site,
                    group = site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~factor(habitat, levels=c('fringing','backreef','outer_10','outer_17')))+
  labs(title = "Distance to centroid\n", y = "Distance\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
# beta dispersion increases as we move from forereef to lagoon, lagoon to shore

```