### 2024-08-30 julianna_coral_macroalg_wrangling.Rmd

---
title: "Wrangling coral and macroalgae data to be joinable"
output: html_document
date: "2024-08-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
```

This script summarizes the coral and macroalgae subsets to be joinable with the full algae dataset

# Bring in the data

```{r}
# algal cover 
macroalgae <- read_csv(here("data", "algaeonly_cleaned_colnames.csv")) %>% 
  mutate(island_side = case_when(site == "lter_1" ~ "north",
                                 site == "lter_2" ~ "north",
                                 site == "lter_3" ~ "east",
                                 site == "lter_4" ~ "east",
                                 site == "lter_5" ~ "west",
                                 site == "lter_6" ~ "west",
                                 TRUE ~ "UdINGLEbUTT")) %>% 
  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
         -neogoniolithon_brassica_florida, -sporolithon_sp) # on Aug 15 Noam and I decided to keep Peysonellia, etc. in

# sand cover and coral cover
sand_coral <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)
```

# Calculate macroalgal alpha diversity

## Make matrices

Make a macroalgal spp matrix to use with `vegan`

```{r make species matrix}
# matrix of just species and their relative percent covers in the same order as the algae tibble
macroalgae %>% 
  select(-year, -location, -site, -island_side, -habitat, -site_habitat, -transect, -quadrat) -> sppMat
```

```{r make a metadata matrix in the same order}
# get ranges to calculate over (this is just for indexing purposes in the next bit of code)
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

# metadata matrix
macroalgae %>% 
  # bring in sand data
  full_join(sand_coral) %>% 
  rename(coral_cover = coral,
         sand_cover = sand) %>% 
  mutate(hard_cover = 100-sand_cover) %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg]), # percent cover of macroalgae
         macroalgae_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (macroalgae_cover/hard_cover)*100 
           ),
         # also for corals
         coral_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (coral_cover/hard_cover)*100 
           )
         ) %>% # percent cover of hard substrate
   # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         island_side_hab_year = paste0(island_side, "_", hab_year)) %>% 
  select(year, location, site, habitat, island_side, site_habitat, 
         hab_year, island_side_hab_year, transect, quadrat, macroalgae_cover, 
         macroalgae_cover_hard, sand_cover, hard_cover, coral_cover, coral_cover_hard)  -> sppMeta
```

## Calculate diversity metrics

```{r get richness and Shannon diversity}
sppMeta %>% 
  # these functions use the vegan package
  mutate(macroalgae_richness = specnumber(sppMat),
         macroalgae_shannon = diversity(sppMat, "shannon"),
         macroalgae_richness_hard = macroalgae_richness/hard_cover,
         macroalgae_shannon_hard = macroalgae_shannon/hard_cover
         ) -> macroAlgDiv
```

# Calculate algal beta diversity

## Distance matrix: habitat-year

We get an error if we run `vegdist()` with quadrats that have no algae. We can see there are 672 of these that we need to filter out:

```{r look for zeros}
sppMat %>% 
  mutate(macroalgae_cover = rowSums(across(where(is.numeric)))) %>% 
  filter(macroalgae_cover == 0) %>% dim() # 7208 with no algae at all (that's like over 1/3!!)
```

Now we can exclude quadrats with no algae and make a species matrix again

```{r get spp mat df}
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

# spp matrix where 0 observations are filtered out
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>% 
  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0
```

```{r metadata matix}
# matching metadata matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>%
  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
  # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         island_side_hab_year = paste0(island_side, hab_year)) -> sppMeta0
```

Now we'll calculate this by habitat because we're compute limited

### Calculate fringing

```{r get fringe mat}
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

# spp matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>% 
  filter(habitat == "fringing") %>% 
  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0fring

# metadata matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>%
   filter(habitat == "fringing") %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
  # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         island_side_hab_year = paste0(island_side, hab_year),
         island_side_year = paste0(island_side, year)) -> sppMeta0fring
```

Calculate Bray Curtis distances

```{r calculate distances}
brayMatFringMacro <- vegdist(sppMat0fring, "bray") # 1 is maximum dissimilarity
```

Calculate dispersion (this is using Noam's code)

```{r calculate beta dispersion}
# yrBetaDisperFringMacro <- vegan::betadisper(brayMatFringMacro, sppMeta0fring$year) # warning: some squared distances are negative and changed to zero (ask Jamie about this??)
```

### Backreef

```{r get backreef df}
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

# spp matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>% 
  filter(habitat == "backreef") %>% 
  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0back

# metadata matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>%
   filter(habitat == "backreef") %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
  # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         island_side_hab_year = paste0(island_side, hab_year),
         island_side_year = paste0(island_side, year)) -> sppMeta0back
```

Calculate Bray Curtis distances

```{r calculate distance}
brayMatBackMacro <- vegdist(sppMat0back, "bray") # 1 is maximum dissimilarity
```

Calculate beta disperion

```{r calculate beta dispersion}
# yrBetaDisperBackMacro <- vegan::betadisper(brayMatBackMacro, sppMeta0back$year) 
```

### Outer 10

```{r get outer 10 df}
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

# spp matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>% 
  filter(habitat == "outer_10") %>% 
  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0outer10

# metadata matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>%
   filter(habitat == "outer_10") %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
  # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         island_side_hab_year = paste0(island_side, hab_year),
         island_side_year = paste0(island_side, year)) -> sppMeta0outer10
```

Calculate Bray Curtis distances

```{r calculate distance}
brayMatOuter10Macro <- vegdist(sppMat0outer10, "bray") # 1 is maximum dissimilarity
```

Calculate beta disperion

```{r calculate beta dispersion}
# yrBetaDisperOuter10Macro <- vegan::betadisper(brayMatOuter10Macro, sppMeta0outer10$year) # error: some squared distances are negative and changed to zero
```

### Outer 17

```{r get outer 17 df}
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

# spp matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>% 
  filter(habitat == "outer_17") %>% 
  select(-year, -location, -island_side, -site, -habitat, -site_habitat, -transect, -quadrat, -macroalgae_cover) -> sppMat0outer17

# metadata matrix
macroalgae %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg])) %>% 
  filter(macroalgae_cover != 0) %>%
   filter(habitat == "outer_17") %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, island_side) %>% 
  # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         island_side_hab_year = paste0(island_side, hab_year),
         island_side_year = paste0(island_side, year)) -> sppMeta0outer17
```

Calculate Bray Curtis distances

```{r calculate distance}
brayMatOuter17Macro <- vegdist(sppMat0outer17, "bray") # 1 is maximum dissimilarity
```

Calculate beta disperion

```{r calculate beta dispersion}
# yrBetaDisperOuter17Macro <- vegan::betadisper(brayMatOuter17Macro, sppMeta0outer17$year) # error: some squared distances are negative and changed to zero
```

### Save dispersions

```{r}
# save(yrBetaDisperFringMacro, file = here("data/fringingyrdispersionMacro.Rdata"))
# save(yrBetaDisperBackMacro, file = here("data/backyrdispersionMacro.Rdata"))
# save(yrBetaDisperOuter10Macro, file = here("data/outer10yrdispersionMacro.Rdata"))
# save(yrBetaDisperOuter17Macro, file = here("data/outer17yrdispersionMacro.Rdata"))
```

### Load dispersions

This is for if you're not running betadisper() for the first time (loading old runs)

```{r}
load(here("data/fringingyrdispersionMacro.Rdata"))
load(here("data/backyrdispersionMacro.Rdata"))
load(here("data/outer10yrdispersionMacro.Rdata"))
load(here("data/outer17yrdispersionMacro.Rdata"))
```

## Join together

Attach back to metadata

```{r join distances with metadata}
sppMeta0fring %>% 
  mutate(distance_centriod = yrBetaDisperFringMacro$distances) -> fringeDist

sppMeta0back %>% 
  mutate(distance_centriod = yrBetaDisperBackMacro$distances) -> backDist

sppMeta0outer10 %>% 
  mutate(distance_centriod = yrBetaDisperOuter10Macro$distances) -> outer10Dist

sppMeta0outer17 %>% 
  mutate(distance_centriod = yrBetaDisperOuter17Macro$distances) -> outer17Dist
```

```{r join together}
# make sure this looks good
colnames(fringeDist) == colnames(backDist) 
colnames(backDist) == colnames(outer10Dist) 
colnames(outer10Dist) == colnames(outer17Dist)

# join together and tidy up
betaDHabYr <- rbind(fringeDist, backDist, outer10Dist, outer17Dist) %>% 
  mutate(hab_yr_distcentroid = distance_centriod) %>% 
  select(year, location, site, habitat, transect, quadrat, hab_yr_distcentroid)
```

# Join all quadrat-level datasets together

```{r join}
macroAlgDiv %>% 
  full_join(betaDHabYr) %>% 
  rename(macroalgae_centroid_dist_habyear_mean = hab_yr_distcentroid) %>% 
  # get just the columns we need
  select(-site_habitat, -hab_year, -island_side_hab_year) -> quadData

```

```{r save}
 write_csv(quadData, here("data/macroalgae_coral_diversity_quadrat.csv"))
```

# Site/habitat/year-level mean

We'll also average these data at the site/habitat/year-level to be more compatible with other data types. This will be done at the *quadrat-level*, meaning we won't take the transect-level summary first, since they aren't meant to be factors in a statistical design

```{r get averages}
quadData %>% 
  group_by(year, site, habitat, island_side) %>% 
  summarize(N = n(),
            coral_cover_mean = mean(coral_cover),
            coral_cover_se = sd(coral_cover)/sqrt(N),
            coral_cover_hard_mean = mean(coral_cover_hard, na.rm = TRUE),
            coral_cover_hard_se = sd(coral_cover_hard, na.rm = TRUE)/sqrt(N),
            hard_cover_mean = mean(hard_cover),
            hard_cover_se = sd(hard_cover)/sqrt(N),

            macroalgae_cover_mean = mean(macroalgae_cover),
            macroalgae_cover_se = sd(macroalgae_cover)/sqrt(N),
            macroalgae_cover_hard_mean = mean(macroalgae_cover_hard, na.rm = TRUE),
            macroalgae_cover_hard_se = sd(macroalgae_cover_hard, na.rm = TRUE)/sqrt(N),
            macroalgae_richness_mean = mean(macroalgae_richness),
            macroalgae_richness_se = sd(macroalgae_richness)/sqrt(N),
            macroalgae_shannon_mean = mean(macroalgae_shannon),
            macroalgae_shannon_se = sd(macroalgae_shannon)/sqrt(N),
            # of hard cover
            macroalgae_richness_hard_mean = mean(macroalgae_richness_hard, na.rm = TRUE),
            macroalgae_richness_hard_se = sd(macroalgae_richness_hard, na.rm = TRUE)/sqrt(N),
            macroalgae_shannon_hard_mean = mean(macroalgae_shannon_hard, na.rm = TRUE),
            macroalgae_shannon_hard_se = sd(macroalgae_shannon_hard, na.rm = TRUE)/sqrt(N),
            macroalgae_centroid_dist_habyear_mean = mean(macroalgae_centroid_dist_habyear_mean, na.rm = TRUE)
            ) -> siteData

```

```{r save dataframe}
 write_csv(siteData, here("data/macroalgae_coral_diversity_site.csv"))
```

# All time metrics

## Calculate the coefficient of variation

This is basically just copying/pasting Noam's code

```{r cv function}
# coefficient of variation
CV <- function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}
```

```{r calculate CV and summarize}
quadData %>% 
  # don't include year this time
  group_by(site, habitat, island_side) %>%
  summarize(N = n(),

            # coral
            coral_cover_mean = mean(coral_cover),
            coral_cover_se = sd(coral_cover)/sqrt(N),
            coral_cover_hard_mean = mean(coral_cover_hard, na.rm = TRUE),
            coral_cover_hard_se = sd(coral_cover_hard, na.rm = TRUE)/sqrt(N),
            coral_cover_cv = CV(coral_cover),
            coral_cover_hard_cv = CV(coral_cover_hard),
            hard_cover_mean = mean(hard_cover),
            hard_cover_se = sd(hard_cover)/sqrt(N),
            hard_cover_cv = CV(hard_cover),

            # macroalgae cover
            macroalgae_cover_mean = mean(macroalgae_cover),
            macroalgae_cover_se = sd(macroalgae_cover)/sqrt(N),
            macroalgae_cover_hard_mean = mean(macroalgae_cover_hard, na.rm = TRUE),
            macroalgae_cover_hard_se = sd(macroalgae_cover_hard, na.rm = TRUE)/sqrt(N),
            macroalgae_cover_cv = CV(macroalgae_cover),
            macroalgae_cover_hard_cv = CV(macroalgae_cover_hard),

            # macroalgae richness
            macroalgae_richness_mean = mean(macroalgae_richness),
            macroalgae_richness_se = sd(macroalgae_richness)/sqrt(N),
            macroalgae_shannon_mean = mean(macroalgae_shannon),
            macroalgae_shannon_se = sd(macroalgae_shannon)/sqrt(N),

            # of hard
            macroalgae_richness_hard_mean = mean(macroalgae_richness_hard, na.rm = TRUE),
            macroalgae_richness_hard_se = sd(macroalgae_richness_hard, na.rm = TRUE)/sqrt(N),
            macroalgae_shannon_hard_mean = mean(macroalgae_shannon_hard, na.rm = TRUE),
            macroalgae_shannon_hard_se = sd(macroalgae_shannon_hard, na.rm = TRUE)/sqrt(N),

            macroalgae_centroid_dist_habyear_mean = mean(macroalgae_centroid_dist_habyear_mean, na.rm = TRUE)
            )  -> alltimeCV
```

```{r save}
write_csv(alltimeCV, here(here("data/macroalgae_coral_diversity_stability_metrics_alltime.csv")))
```