### 2024-09-04 noam_final_analyses.Rmd

---
title: "noam_final_analyses"
author: "Noam Altman-Kurosaki"
date: "2024-08-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(ecofolio) # possible other method for examining portfolio effects 
```

```{r import and format data}
# coefficient of variation
CV<-function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}

# for figures
island_side_colours <- c("north" = "darkblue",
                         "east" = "mediumorchid",
                         "west" = "orange")
habitat_colours <- c("fringing" = "#D81B60",
                     "backreef" = "#FFC107",
                     "outer_10" = "#1E88E5",
                     "outer_17" = "#004D40")

# starting with all data now because of 0 algae transects/quadrats
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

data <- data[ , -which(names(data) %in% not_algae)]

# adding island side - adapted from Julianna's code
data <- data %>% mutate(island_side = case_when(site == "lter_1" ~ "north",
                                 site == "lter_2" ~ "north",
                                 site == "lter_3" ~ "east",
                                 site == "lter_4" ~ "east",
                                 site == "lter_5" ~ "west",
                                 site == "lter_6" ~ "west"))
data$island_side_habitat <- paste(data$island_side, data$habitat)

macroalgae <- data %>% 
  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
         -neogoniolithon_brassica_florida, -sporolithon_sp)

sand_coral <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)

# get ranges to calculate over (this is just for indexing purposes in the next bit of code)
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

macroalgae %>% 
  # bring in sand data
  full_join(sand_coral) %>% 
  rename(coral_cover = coral,
         sand_cover = sand) %>% 
  mutate(hard_cover = 100-sand_cover) %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg]), # percent cover of macroalgae

         macroalgae_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (macroalgae_cover/hard_cover)*100 
           ),
         # also for corals
         coral_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (coral_cover/hard_cover)*100 
           )
         ) %>% # percent cover of hard substrate
   # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         site_hab_year = paste0(site, "_", hab_year)) %>% 
  select(year, location, site, habitat, island_side, site_habitat, island_side_habitat,
         hab_year, site_hab_year, transect, quadrat, macroalgae_cover, 
         macroalgae_cover_hard, sand_cover, hard_cover, coral_cover, coral_cover_hard)  -> meta_df

meta_df$all_algae_cover <- rowSums(data[startAlg:endAlg]) # percent cover of all algae
meta_df$all_algae_cover_hard <- case_when(
           meta_df$sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (meta_df$all_algae_cover/meta_df$hard_cover)*100   
         )

data <- merge(meta_df, data )
macroalgae <- merge(meta_df, macroalgae)

meta_cols <- colnames(meta_df)

full_species_cols <- colnames(data)[!colnames(data) %in% meta_cols]
macro_species_cols <- colnames(macroalgae)[!colnames(macroalgae) %in% meta_cols]

data[,full_species_cols] <- data[,full_species_cols]/data$hard_cover # transform cover to proportion of hardcover in quadrat

macroalgae[,macro_species_cols] <- macroalgae[,macro_species_cols]/macroalgae$hard_cover # repeat for macro dataframe

# summarize at site level
site_full <- data %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, hab_year, site_hab_year) %>%
  dplyr::summarize(across(all_of(
    c("macroalgae_cover", "macroalgae_cover_hard", "sand_cover", "hard_cover", "coral_cover",
      "coral_cover_hard","all_algae_cover", "all_algae_cover_hard", full_species_cols)), mean, na.rm = TRUE))

# summarize at species level
site_macro <- macroalgae %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, hab_year, site_hab_year) %>%
  dplyr::summarize(across(all_of(
    c("macroalgae_cover", "macroalgae_cover_hard", "sand_cover", "hard_cover", "coral_cover",
      "coral_cover_hard","all_algae_cover", "all_algae_cover_hard", macro_species_cols)), mean, na.rm = TRUE))

# extract site level meta_df for future merges
site_meta_df <- site_full[,which(names(site_full) %in% meta_cols)]

#pivot long 
long_data <- pivot_longer(
  data = data,
  cols = grep("acanthophora_spicifera", colnames(data)):grep("valonia_ventricosa", colnames(data)),
  names_to = "taxa",
  values_to = "percent_cover")

macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover")

site_long_data <- pivot_longer(
  data = site_full,
  cols = grep("acanthophora_spicifera", colnames(site_full)):grep("valonia_ventricosa", colnames(site_full)),
  names_to = "taxa",
  values_to = "percent_cover")

site_long_macro <- pivot_longer(
  data = site_macro,
  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
  names_to = "taxa",
  values_to = "percent_cover")

```

```{r create adiv dataframes}
alpha_diversity_quad <-
  data.frame(
    year = data$year,
    location = data$location,
    shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
    richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
  )

alpha_diversity_quad <- merge(meta_df, alpha_diversity_quad, by = c("location", "year") )

alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
  )

alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))

alpha_diversity_site <-
  data.frame(
    year = site_full$year,
    site_hab_year = site_full$site_hab_year,
    shannon = diversity(site_full[ , -which(names(site_full) %in% meta_cols)], "shannon"),
    richness = rowSums(site_full[ , -which(names(site_full) %in% meta_cols)] > 0)
  )

alpha_diversity_site <- merge(site_meta_df, alpha_diversity_site, by= c("year", "site_hab_year"))

alpha_diversity_site_macro <-
  data.frame(
    year = site_macro$year,
    site_hab_year = site_macro$site_hab_year,
    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
  )

alpha_diversity_site_macro <- merge(site_meta_df, alpha_diversity_site_macro, by= c("year", "site_hab_year"))

```

```{r read and merge secondary datasets, fig.height=15, fig.width=15}
wave <- read.csv(here::here("data", "algae_wave_dynamics.csv"), stringsAsFactors = F)
wave <- wave[,-which(names(wave) %in% "X")] # remove rownum columns
wave$site_hab_year <- paste0(wave$site,"_", wave$habitat,"_", wave$year) # for merge

nutrients <- read.csv(here::here("data", "lter_nuts.csv"), stringsAsFactors = F)
nutrients$site_hab_year <- paste0(nutrients$site,"_", nutrients$habitat,"_", nutrients$year)

temp <- read.csv(here::here("data", "temp_sum.csv"), stringsAsFactors = F)
temp$site_hab_year <- paste0(temp$site,"_", temp$habitat,"_", temp$year)

herb <- read.csv(here::here("data", "herbivore_biomass_and_diversity_site.csv"), stringsAsFactors = F)
# create column for total herbivore biomass
herb$herbivore_biomass <- rowSums(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))])
herb<-herb %>%  # rename columns for merge
  rename(
    year = Year,
    site = Site,
    habitat = Habitat,
    herbivore_richness = richness,
    herbivore_shannon = shannon,
    herbivore_evenness = evenness
    )
# fix site names and habitat names
herb <-
  herb %>% mutate(site = case_when(site == 1 ~ "lter_1",
                              site == 2 ~ "lter_2",
                              site == 3 ~ "lter_3",
                              site == 4 ~ "lter_4",
                              site == 5 ~ "lter_5",
                              site == 6 ~ "lter_6"))
herb <-
  herb %>% mutate(habitat = case_when(habitat == "Fringing" ~ "fringing",
                                      habitat == "Backreef" ~ "backreef",
                                      habitat == "Outer 10" ~ "outer_10")) # there is no outer 17 for herb data
herb$site_hab_year <- paste0(herb$site,"_", herb$habitat,"_", herb$year)
herb <- subset(herb, year != 2006)

### Merge with alpha diversity datasets

alpha_diversity_site <- full_join(alpha_diversity_site, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_site <- full_join(alpha_diversity_site, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site <- full_join(alpha_diversity_site,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site <- full_join(alpha_diversity_site, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_site, ~ all_algae_cover_hard + macroalgae_cover_hard + richness + shannon + coral_cover_hard + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_site_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_quad <- full_join(alpha_diversity_quad, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_quad <- full_join(alpha_diversity_quad, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad <- full_join(alpha_diversity_quad,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad <- full_join(alpha_diversity_quad, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_quad, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_quad_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

```

```{r all algae alpha diversity site level}

site_full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site) # need to include offset because of the way richness is calculated
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_full_richness_mod) # low variance in random effect
Anova(site_full_richness_mod) 
#                 Chisq Df Pr(>Chisq)    
# year           1.5505  1     0.2131    
# habitat      106.8578  3     <2e-16 ***
# year:habitat   5.3674  3     0.1468    
# habitat more important than time when time is numeric
# check_mod
hist(residuals(site_full_richness_mod)) # slight skew

# striping/patterining expected because of discrete response var.
plot(residuals(site_full_richness_mod) ~ fitted(site_full_richness_mod)) # slightly heteroskedastic

emtrends(site_full_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
site_full_richness_emm <- emmip(site_full_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 18, yend = 18, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_full_shannon_mod <- glmmTMB(log(shannon + 1) ~ year*habitat + (1|site), 
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed. No offset because calculation of shannon already incorporated differences in %hard
summary(site_full_shannon_mod) # low variance again
Anova(site_full_shannon_mod)
# Response: log(shannon + 1)
#                Chisq Df Pr(>Chisq)    
# year          0.0133  1   0.908348    
# habitat      97.6274  3  < 2.2e-16 ***
# year:habitat 14.0548  3   0.002831 **

hist(residuals(site_full_shannon_mod)) # looks good
plot(residuals(site_full_shannon_mod) ~ fitted(site_full_shannon_mod)) # looks good here too

emtrends(site_full_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ, backreef different fro outer17
site_full_shannon_emm <- emmip(site_full_shannon_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_shannon_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_shannon_emm, aes(x = year, y = yvar, colour = habitat), size = 2) +
  geom_ribbon(data = site_full_shannon_emm, aes(x = year, ymin = LCL, ymax = UCL, fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```

```{r macroalgae alpha diversity site level}

site_macro_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site_macro) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
summary(site_macro_richness_mod) # low variance again
Anova(site_macro_richness_mod)
#                Chisq Df Pr(>Chisq)    
# year          1.1856  1    0.27621    
# habitat      87.7765  3    < 2e-16 ***
# year:habitat  7.0855  3    0.06922 .  
hist(residuals(site_macro_richness_mod)) # slightly more skew
plot(residuals(site_macro_richness_mod) ~ fitted(site_macro_richness_mod)) # similar patterns

emtrends(site_macro_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
site_macro_richness_emm <- emmip(site_macro_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro,
             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  geom_segment(aes(y = 18, yend = 18, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_macro_shannon_mod <- glmmTMB(log(shannon +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed
summary(site_macro_shannon_mod)
Anova(site_macro_shannon_mod)
#                Chisq Df Pr(>Chisq)    
# year          0.0133  1   0.908348    
# habitat      97.6274  3  < 2.2e-16 ***
# year:habitat 14.0548  3   0.002831 ** 
hist(residuals(site_macro_shannon_mod)) # looks good
plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod)) # generally fine

emtrends(site_macro_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ
site_macro_shannon_emm <- emmip(site_macro_shannon_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_shannon_plot <- 
  ggplot() +
 geom_point(data = alpha_diversity_site,
            aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_shannon_emm, aes(x = year, y = yvar, colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_shannon_emm, aes(x = year, ymin = LCL, ymax = UCL, fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  geom_segment(aes(y = 2.5, yend = 2.5, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```
```{r stepwise models site level}

## FRINGING - exclude wave power
fringing_all_site <- na.omit(subset(alpha_diversity_site, habitat == "fringing")) # only use rows that contain all factors

fringing_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass) # universal model has convergence issues
fringing_all_richness_dredge <- dredge(fringing_all_richness_mod_universal) # most models have convergence issues
model.sel(fringing_all_richness_dredge) # top model just includes year
fringing_all_richness_top_mod <- glmmTMB(richness ~ year + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass)
summary(fringing_all_richness_top_mod)
hist(residuals(fringing_all_richness_top_mod)) # slight skew but not bad
plot(residuals(fringing_all_richness_top_mod) ~ fitted(fringing_all_richness_top_mod)) # looks fine

fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass)
fringing_all_shannon_dredge <- dredge(fringing_all_shannon_mod_universal) 
model.sel(fringing_all_shannon_dredge) # tie for the top model - best includes year and coral cover or just nitrogen
fringing_all_shannon_top_mod1 <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + (1|site), data = fringing_all_site, na.action = na.pass)
fringing_all_shannon_top_mod2 <- glmmTMB(log(shannon + 1) ~ mean_n + mean_temp_c + (1|site), data = fringing_all_site, na.action = na.pass)
hist(residuals(fringing_all_shannon_top_mod1)) # slightly bimodal
hist(residuals(fringing_all_shannon_top_mod2)) # also uniform - neither are great
plot(residuals(fringing_all_shannon_top_mod1) ~ fitted(fringing_all_shannon_top_mod1)) # slight patterning
plot(residuals(fringing_all_shannon_top_mod2) ~ fitted(fringing_all_shannon_top_mod2)) # similar here

## Backreef - includes all 
backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors

backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has convergence issues
backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
model.sel(backreef_all_richness_dredge) # top model includes year and coral 
backreef_all_richness_top_mod <- glmmTMB(richness ~ year + coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass)
summary(backreef_all_richness_top_mod)
hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine

backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass)
backreef_all_shannon_dredge <- dredge(backreef_all_shannon_mod_universal) 
model.sel(backreef_all_shannon_dredge) #  best includes just coral cover 
backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  coral_cover_hard + (1|site), data = backreef_all_site, na.action = na.pass)
hist(residuals(backreef_all_shannon_top_mod)) # pretty good
plot(residuals(backreef_all_shannon_top_mod) ~ fitted(backreef_all_shannon_top_mod)) # looks fine

## outer10 - no nitrogen 
outer10_all_site <- subset(alpha_diversity_site, habitat == "outer_10") # only use rows that contain all factors
# remove nitrogent columns
outer10_all_site <- outer10_all_site[,-which(names(outer10_all_site) %in% c("mean_n", "n_cv"))]
# remove other na rows
outer10_all_site <- na.omit(outer10_all_site)

outer10_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass) # universal model has convergence issues
outer10_all_richness_dredge <- dredge(outer10_all_richness_mod_universal) # most models have convergence issues
model.sel(outer10_all_richness_dredge) # top model includes just intercept... next best is coral 
outer10_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass)
summary(outer10_all_richness_top_mod)
hist(residuals(outer10_all_richness_top_mod)) #  not bad
plot(residuals(outer10_all_richness_top_mod) ~ fitted(outer10_all_richness_top_mod)) # looks fine

outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) # some convergence issues
model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness - first time there's been a decent gap in AICc
outer10_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
hist(residuals(outer10_all_shannon_top_mod)) # pretty good
plot(residuals(outer10_all_shannon_top_mod) ~ fitted(outer10_all_shannon_top_mod)) # looks fine

## outer17 - no nitrogen or herbivores
outer17_all_site <- subset(alpha_diversity_site, habitat == "outer_17") # only use rows that contain all factors
# remove nitrogent columns
outer17_all_site <- outer17_all_site[,-which(names(outer17_all_site) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
# remove other na rows
outer17_all_site <- na.omit(outer17_all_site)

outer17_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass) # universal model has convergence issues
outer17_all_richness_dredge <- dredge(outer17_all_richness_mod_universal) # most models have convergence issues
model.sel(outer17_all_richness_dredge) # top model includes just intercept... next best is coral 
outer17_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass)
summary(outer17_all_richness_top_mod)
hist(residuals(outer17_all_richness_top_mod)) #  not bad
plot(residuals(outer17_all_richness_top_mod) ~ fitted(outer17_all_richness_top_mod)) # looks fine

outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass)
outer17_all_shannon_dredge <- dredge(outer17_all_shannon_mod_universal) # some convergence issues
model.sel(outer17_all_shannon_dredge) #  best includes just year
outer17_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = outer17_all_site, na.action = na.pass)
hist(residuals(outer17_all_shannon_top_mod)) # pretty good
plot(residuals(outer17_all_shannon_top_mod) ~ fitted(outer17_all_shannon_top_mod)) # looks fine
```

```{r all algae alpha diversity quadrat level}

full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site/location) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_quad[!is.na(alpha_diversity_quad$macroalgae_cover_hard),]) 
summary(full_richness_mod)
hist(residuals(full_richness_mod)) # slightly skewed but not terrible

# striping/patterining expected because of discrete response var. otherwise fine
plot(residuals(full_richness_mod) ~ fitted(full_richness_mod))

emtrends(full_richness_mod, pairwise ~  habitat, var = "year")
full_richness_emm <- emmip(full_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad,
             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.05) + 
  geom_line(data = full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Richness") +
  scale_y_log10() +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```

```{r macroalgae diversity quadrat level}

```