### 2024-09-06 noam_final_analyses.Rmd
potential fix for error lauren caught in allalgae metric

---
title: "noam_final_analyses"
author: "Noam Altman-Kurosaki"
date: "2024-08-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(ecofolio) # possible other method for examining portfolio effects 
```

```{r import and format data}
# coefficient of variation
CV<-function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}

inv_logit <- function(x){
  exp(x)/(1+exp(x))
}

sv_trans <- function(prop, s = 0.000005){
  (prop*(length(prop) - 1) + s)/length(prop)
  # where prop is the vector of the proportional value you're transforming, 
  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
  # and s is a small offset 
}

# for figures
island_side_colours <- c("north" = "darkblue",
                         "east" = "mediumorchid",
                         "west" = "orange")
habitat_colours <- c("fringing" = "#D81B60",
                     "backreef" = "#FFC107",
                     "outer_10" = "#1E88E5",
                     "outer_17" = "#004D40")

# starting with all data now because of 0 algae transects/quadrats
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F)

not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")

data <- data[ , -which(names(data) %in% not_algae)]

# adding island side - adapted from Julianna's code
data <- data %>% mutate(island_side = case_when(site == "lter_1" ~ "north",
                                 site == "lter_2" ~ "north",
                                 site == "lter_3" ~ "east",
                                 site == "lter_4" ~ "east",
                                 site == "lter_5" ~ "west",
                                 site == "lter_6" ~ "west"))
data$island_side_habitat <- paste(data$island_side, data$habitat)

macroalgae <- data %>% 
  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
         -neogoniolithon_brassica_florida, -sporolithon_sp)

sand_coral <- read_csv(here("data", "allbenthicdata_cleaned_colnames.csv")) %>% 
  select(year, location, site, habitat, site_habitat, transect, quadrat, sand, coral)

# get ranges to calculate over (this is just for indexing purposes in the next bit of code)
startAlg <- grep("acanthophora_spicifera", colnames(macroalgae))
endAlg <- grep("valonia_ventricosa", colnames(macroalgae))

macroalgae %>% 
  # bring in sand data
  full_join(sand_coral) %>% 
  rename(coral_cover = coral,
         sand_cover = sand) %>% 
  mutate(hard_cover = 100-sand_cover) %>% 
  mutate(macroalgae_cover = rowSums(macroalgae[startAlg:endAlg]), # percent cover of macroalgae

         macroalgae_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (macroalgae_cover/hard_cover)*100 
           ),
         # also for corals
         coral_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (coral_cover/hard_cover)*100 
           )
         ) %>% # percent cover of hard substrate
   # make a year-habitat column for betadisper
  mutate(hab_year = paste0(habitat, "_", year),
         site_hab_year = paste0(site, "_", hab_year)) %>% 
  select(year, location, site, habitat, island_side, site_habitat, island_side_habitat,
         hab_year, site_hab_year, transect, quadrat, macroalgae_cover, 
         macroalgae_cover_hard, sand_cover, hard_cover, coral_cover, coral_cover_hard)  -> meta_df

meta_df$all_algae_cover <- rowSums(data[grep("acanthophora_spicifera", colnames(data)):
                                        grep("valonia_ventricosa", colnames(data))]) # percent cover of all algae
meta_df$all_algae_cover_hard <- case_when(
           meta_df$sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (meta_df$all_algae_cover/meta_df$hard_cover)*100   
         )

data <- merge(meta_df, data )
macroalgae <- merge(meta_df, macroalgae)

meta_cols <- colnames(meta_df)

full_species_cols <- colnames(data)[!colnames(data) %in% meta_cols]
macro_species_cols <- colnames(macroalgae)[!colnames(macroalgae) %in% meta_cols]

data[,full_species_cols] <- data[,full_species_cols]/data$hard_cover # transform cover to proportion of hardcover in quadrat

macroalgae[,macro_species_cols] <- macroalgae[,macro_species_cols]/macroalgae$hard_cover # repeat for macro dataframe

# summarize at site level
site_full <- data %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, hab_year, site_hab_year) %>%
  dplyr::summarize(across(all_of(
    c("macroalgae_cover", "macroalgae_cover_hard", "sand_cover", "hard_cover", "coral_cover",
      "coral_cover_hard","all_algae_cover", "all_algae_cover_hard", full_species_cols)), mean, na.rm = TRUE))

# summarize at species level
site_macro <- macroalgae %>%
  dplyr::group_by(year, island_side, site, habitat, site_habitat, hab_year, site_hab_year) %>%
  dplyr::summarize(across(all_of(
    c("macroalgae_cover", "macroalgae_cover_hard", "sand_cover", "hard_cover", "coral_cover",
      "coral_cover_hard","all_algae_cover", "all_algae_cover_hard", macro_species_cols)), mean, na.rm = TRUE))

# extract site level meta_df for future merges
site_meta_df <- site_full[,which(names(site_full) %in% meta_cols)]

#pivot long 
long_data <- pivot_longer(
  data = data,
  cols = grep("acanthophora_spicifera", colnames(data)):grep("valonia_ventricosa", colnames(data)),
  names_to = "taxa",
  values_to = "percent_cover")

macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover")

site_long_data <- pivot_longer(
  data = site_full,
  cols = grep("acanthophora_spicifera", colnames(site_full)):grep("valonia_ventricosa", colnames(site_full)),
  names_to = "taxa",
  values_to = "percent_cover")

site_long_macro <- pivot_longer(
  data = site_macro,
  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
  names_to = "taxa",
  values_to = "percent_cover")

```

```{r create adiv dataframes}
alpha_diversity_quad <-
  data.frame(
    year = data$year,
    location = data$location,
    shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
    richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
  )

alpha_diversity_quad <- merge(meta_df, alpha_diversity_quad, by = c("location", "year") )

alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
  )

alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))

alpha_diversity_site <-
  data.frame(
    year = site_full$year,
    site_hab_year = site_full$site_hab_year,
    shannon = diversity(site_full[ , -which(names(site_full) %in% meta_cols)], "shannon"),
    richness = rowSums(site_full[ , -which(names(site_full) %in% meta_cols)] > 0)
  )

alpha_diversity_site <- merge(site_meta_df, alpha_diversity_site, by= c("year", "site_hab_year"))

alpha_diversity_site_macro <-
  data.frame(
    year = site_macro$year,
    site_hab_year = site_macro$site_hab_year,
    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
  )

alpha_diversity_site_macro <- merge(site_meta_df, alpha_diversity_site_macro, by= c("year", "site_hab_year"))
alpha_diversity_site_macro$richness_area <- alpha_diversity_site_macro$richness/(alpha_diversity_site_macro$hard_cover/100)

```

```{r beta dispersion}
full_site_species_matrix <- as.matrix(site_full[ , -which(names(site_full) %in% meta_cols)])
full_site_species_matrix <- full_site_species_matrix/site_full$hard_cover # convert from percent to proportion of hard cover
rownames(full_site_species_matrix) <- paste(site_full$year,"-", site_full$site_habitat, sep = "")

full_site_species_matrix <- asin(sqrt(full_site_species_matrix)) # arcsin squareroot transform

full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% colnames(site_meta_df))])
full_site_macro_species_matrix <- full_site_macro_species_matrix/site_macro$hard_cover # convert from percent to proportion of hard cover
rownames(full_site_macro_species_matrix) <- paste(site_macro$year,"-", site_macro$site_habitat, sep = "")

full_site_macro_species_matrix <- asin(sqrt(full_site_macro_species_matrix)) # arcsin squareroot transform

full_site_bray <- vegdist(full_site_species_matrix, "bray") 
full_site_macro_bray <- vegdist(full_site_macro_species_matrix, "bray")

full_site_betadisper <- vegan::betadisper(full_site_bray, site_full$hab_year)
site_full$betadispersion <- full_site_betadisper$distances[match(site_full$stratum, names(full_site_betadisper$distances))]
site_full$betadispersion_trans <- sv_trans(site_full$betadispersion, site_full)

betadisper_site_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = site_full)
summary(betadisper_site_mod) # fair bit of within site variance - inv_logit(0.04) = 0.5
hist(residuals(betadisper_site_mod)) # pretty good
plot(residuals(betadisper_site_mod) ~ fitted(betadisper_site_mod)) # slight heteroskedasticity
Anova(betadisper_site_mod)
#                Chisq Df Pr(>Chisq)    
# year         12.9561  1  0.0003189 ***
# habitat      27.2191  3  5.296e-06 ***
# year:habitat  6.8472  3  0.0769311 .

emtrends(betadisper_site_mod, pairwise ~ habitat , "year") # fringing slope significantly different from outer17
betadisper_site_emm <- emmip(betadisper_site_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)

ggplot()+
   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
   geom_line(data = betadisper_site_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
   geom_ribbon(data = betadisper_site_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
   scale_fill_manual(values = habitat_colours) +
   geom_segment(aes(y = 0.6, yend = 0.6, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   labs(y = "Beta dispersion\n", x = "Year") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") )# make legend horizontal

full_site_macro_betadisper <- vegan::betadisper(full_site_macro_bray, site_macro$hab_year)
full_subset_site_macro$betadispersion <- full_site_macro_betadisper$distances[match(full_subset_site_macro$stratum, names(full_site_macro_betadisper$distances))]
full_subset_site_macro$betadispersion_trans <- sv_trans(full_subset_site_macro$betadispersion, full_subset_site_macro)

betadisper_site_macro_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = full_subset_site_macro)
summary(betadisper_site_macro_mod) # similar site variance - inv_logit(0.01) = 0.5
hist(residuals(betadisper_site_macro_mod)) # pretty good
plot(residuals(betadisper_site_macro_mod) ~ fitted(betadisper_site_macro_mod)) # slight heteroskedasticity and negative trend
Anova(betadisper_site_macro_mod)
#               Chisq Df Pr(>Chisq)    
# year         36.591  1  1.457e-09 ***
# habitat      46.029  3  5.590e-10 ***
# year:habitat 31.232  3  7.597e-07 ***

emtrends(betadisper_site_macro_mod, pairwise ~ habitat , "year") # backreef distinct from all other habitats
betadisper_site_macro_emm <- emmip(betadisper_site_macro_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)

ggplot()+
   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
   geom_line(data = betadisper_site_macro_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
   geom_ribbon(data = betadisper_site_macro_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
   scale_fill_manual(values = habitat_colours) +
   geom_segment(aes(y = 0.75, yend = 0.75, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   labs(y = "Beta dispersion\n", x = "Year") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") )# make legend horizontal

```

```{r read and merge secondary datasets, fig.height=15, fig.width=15}
wave <- read.csv(here::here("data", "algae_wave_dynamics.csv"), stringsAsFactors = F)
wave <- wave[,-which(names(wave) %in% "X")] # remove rownum columns
wave$site_hab_year <- paste0(wave$site,"_", wave$habitat,"_", wave$year) # for merge

nutrients <- read.csv(here::here("data", "lter_nuts.csv"), stringsAsFactors = F)
nutrients$site_hab_year <- paste0(nutrients$site,"_", nutrients$habitat,"_", nutrients$year)

temp <- read.csv(here::here("data", "temp_sum.csv"), stringsAsFactors = F)
temp$site_hab_year <- paste0(temp$site,"_", temp$habitat,"_", temp$year)

herb <- read.csv(here::here("data", "herbivore_biomass_and_diversity_site.csv"), stringsAsFactors = F)
# create column for total herbivore biomass
herb$herbivore_biomass <- rowSums(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))])
herb<-herb %>%  # rename columns for merge
  rename(
    year = Year,
    site = Site,
    habitat = Habitat,
    herbivore_richness = richness,
    herbivore_shannon = shannon,
    herbivore_evenness = evenness
    )
# fix site names and habitat names
herb <-
  herb %>% mutate(site = case_when(site == 1 ~ "lter_1",
                              site == 2 ~ "lter_2",
                              site == 3 ~ "lter_3",
                              site == 4 ~ "lter_4",
                              site == 5 ~ "lter_5",
                              site == 6 ~ "lter_6"))
herb <-
  herb %>% mutate(habitat = case_when(habitat == "Fringing" ~ "fringing",
                                      habitat == "Backreef" ~ "backreef",
                                      habitat == "Outer 10" ~ "outer_10")) # there is no outer 17 for herb data
herb$site_hab_year <- paste0(herb$site,"_", herb$habitat,"_", herb$year)
herb <- subset(herb, year != 2006)

### Merge with alpha diversity datasets

alpha_diversity_site <- full_join(alpha_diversity_site, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_site <- full_join(alpha_diversity_site, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site <- full_join(alpha_diversity_site,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site <- full_join(alpha_diversity_site, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_site, ~ all_algae_cover_hard + macroalgae_cover_hard + richness + shannon + coral_cover_hard + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_site_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_quad <- full_join(alpha_diversity_quad, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_quad <- full_join(alpha_diversity_quad, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad <- full_join(alpha_diversity_quad,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad <- full_join(alpha_diversity_quad, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_quad, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_quad_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

```

```{r all algae alpha diversity site level}

site_full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site) # need to include offset because of the way richness is calculated
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_full_richness_mod) # low variance in random effect
Anova(site_full_richness_mod) 
#                 Chisq Df Pr(>Chisq)    
# year           1.5505  1     0.2131    
# habitat      106.8578  3     <2e-16 ***
# year:habitat   5.3674  3     0.1468    
# habitat more important than time when time is numeric
# check_mod
hist(residuals(site_full_richness_mod)) # slight skew

# striping/patterining expected because of discrete response var.
plot(residuals(site_full_richness_mod) ~ fitted(site_full_richness_mod)) # slightly heteroskedastic

emtrends(site_full_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
site_full_richness_emm <- emmip(site_full_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 18, yend = 18, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_full_shannon_mod <- glmmTMB(log(shannon + 1) ~ year*habitat + (1|site), 
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed. No offset because calculation of shannon already incorporated differences in %hard
summary(site_full_shannon_mod) # low variance again
Anova(site_full_shannon_mod)
# Response: log(shannon + 1)
#                Chisq Df Pr(>Chisq)    
# year          0.0133  1   0.908348    
# habitat      97.6274  3  < 2.2e-16 ***
# year:habitat 14.0548  3   0.002831 **

hist(residuals(site_full_shannon_mod)) # looks good
plot(residuals(site_full_shannon_mod) ~ fitted(site_full_shannon_mod)) # looks good here too

emtrends(site_full_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ, backreef different fro outer17
site_full_shannon_emm <- emmip(site_full_shannon_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_shannon_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_shannon_emm, aes(x = year, y = yvar, colour = habitat), size = 2) +
  geom_ribbon(data = site_full_shannon_emm, aes(x = year, ymin = LCL, ymax = UCL, fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```

```{r macroalgae alpha diversity site level}

site_macro_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site_macro) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
summary(site_macro_richness_mod) # low variance again
Anova(site_macro_richness_mod)
#                Chisq Df Pr(>Chisq)    
# year          1.1856  1    0.27621    
# habitat      87.7765  3    < 2e-16 ***
# year:habitat  7.0855  3    0.06922 .  
hist(residuals(site_macro_richness_mod)) # slightly more skew
plot(residuals(site_macro_richness_mod) ~ fitted(site_macro_richness_mod)) # similar patterns

emtrends(site_macro_richness_mod, pairwise ~  habitat, var = "year") # habitat trends don't differ from each other
site_macro_richness_emm <- emmip(site_macro_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro,
             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  geom_segment(aes(y = 18, yend = 18, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_macro_shannon_mod <- glmmTMB(log(shannon +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed
summary(site_macro_shannon_mod)
Anova(site_macro_shannon_mod)
#                Chisq Df Pr(>Chisq)    
# year          0.0133  1   0.908348    
# habitat      97.6274  3  < 2.2e-16 ***
# year:habitat 14.0548  3   0.002831 ** 
hist(residuals(site_macro_shannon_mod)) # looks good
plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod)) # generally fine

emtrends(site_macro_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ
site_macro_shannon_emm <- emmip(site_macro_shannon_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_shannon_plot <- 
  ggplot() +
 geom_point(data = alpha_diversity_site,
            aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_shannon_emm, aes(x = year, y = yvar, colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_shannon_emm, aes(x = year, ymin = LCL, ymax = UCL, fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  geom_segment(aes(y = 2.5, yend = 2.5, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```

```{r stepwise models site level, fig.height = 8, fig.width=12}

## FRINGING - exclude wave power
fringing_all_site <- na.omit(subset(alpha_diversity_site, habitat == "fringing")) # only use rows that contain all factors
# 
# fringing_all_richness_mod_universal <- glmmTMB(richness ~ year + mean_n + mean_temp_c +  herbivore_evenness +
# offset(log(hard_cover)) #+ (1|site), family = poisson, data = fringing_all_site, na.action = na.pass) 
# fringing_all_richness_dredge <- dredge(fringing_all_richness_mod_universal) 
# model.sel(fringing_all_richness_dredge) # top model just includes year
# fringing_all_richness_top_mod <- glmmTMB(richness ~ year + offset(log(hard_cover)) + (1|site), family = poisson, data = # fringing_all_site, na.action = na.pass)
# summary(fringing_all_richness_top_mod)
# hist(residuals(fringing_all_richness_top_mod)) # slight skew but not bad
# plot(residuals(fringing_all_richness_top_mod) ~ fitted(fringing_all_richness_top_mod)) # looks fine

fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass)
fringing_all_shannon_dredge <- dredge(fringing_all_shannon_mod_universal) 
model.sel(fringing_all_shannon_dredge) # tie for the top model just nitrogen
fringing_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ mean_n + (1|site), data = fringing_all_site, na.action = na.pass)

qqnorm(residuals(fringing_all_shannon_top_mod))# not bad
qqline(residuals(fringing_all_shannon_top_mod))
plot(residuals(fringing_all_shannon_top_mod) ~ fitted(fringing_all_shannon_top_mod)) # slight patterning
summary(fringing_all_shannon_top_mod)

fringing_emm <-as.data.frame(emmeans(fringing_all_shannon_top_mod, ~mean_n,
                                     at = list(mean_n = seq(min(fringing_all_site$mean_n),
                                                            max(fringing_all_site$mean_n),
                                                            0.01)))) 
emmip(fringing_all_shannon_top_mod, 1~ mean_n)
ggplot()+
  geom_point(data = fringing_all_site, aes(x = mean_n, y = shannon))+
  geom_line(data = fringing_emm, aes(x = mean_n, y = emmean)) +
  geom_ribbon(data = fringing_emm, aes(x = mean_n, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
  scale_y_log10() + 
  labs(x = "Mean N", y = "Fringing Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

## Backreef - includes all 
# backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
# 
# backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + # offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has # convergence issues
# backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
# model.sel(backreef_all_richness_dredge) # top model includes year and coral 
# backreef_all_richness_top_mod <- glmmTMB(richness ~ year  + offset(log(hard_cover)) + (1|site), family = poisson, data = # backreef_all_site, na.action = na.pass)
# summary(backreef_all_richness_top_mod)
# hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
# plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine

backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass)
backreef_all_shannon_dredge <- dredge(backreef_all_shannon_mod_universal) 
model.sel(backreef_all_shannon_dredge) #  best includes just year, next best also includes herbivore evenness
backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = backreef_all_site, na.action = na.pass)
qqnorm(residuals(backreef_all_shannon_top_mod)) #  a little wonky on the tail but not bad
plot(residuals(backreef_all_shannon_top_mod) ~ fitted(backreef_all_shannon_top_mod)) # looks fine
summary(backreef_all_shannon_top_mod)

## outer10 - no nitrogen 
outer10_all_site <- subset(alpha_diversity_site, habitat == "outer_10") # only use rows that contain all factors
# remove nitrogent columns
outer10_all_site <- outer10_all_site[,-which(names(outer10_all_site) %in% c("mean_n", "n_cv"))]
# remove other na rows
outer10_all_site <- na.omit(outer10_all_site)

# outer10_all_richness_mod_universal <- glmmTMB(richness ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + # offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass) # universal model has convergence # issues
# outer10_all_richness_dredge <- dredge(outer10_all_richness_mod_universal) # most models have convergence issues
# model.sel(outer10_all_richness_dredge)

# 
outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) 
model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness 
outer10_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
qqnorm(residuals(outer10_all_shannon_top_mod)) # pretty good
plot(residuals(outer10_all_shannon_top_mod) ~ fitted(outer10_all_shannon_top_mod)) # looks fine
summary(outer10_all_shannon_top_mod)

outer10_emm <-as.data.frame(emmeans(outer10_all_shannon_top_mod, ~herbivore_evenness,
                                     at = list(herbivore_evenness = seq(min(outer10_all_site$herbivore_evenness),
                                                            max(outer10_all_site$herbivore_evenness),
                                                            0.05))))

ggplot()+
  geom_point(data = outer10_all_site, aes(x = herbivore_evenness, y = shannon))+
  geom_line(data = outer10_emm, aes(x = herbivore_evenness, y = emmean)) +
  geom_ribbon(data = outer10_emm, aes(x = herbivore_evenness, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
  scale_y_log10() + 
  labs(x = "Herbivore evenness", y = "Outer 10 Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

## outer17 - no nitrogen or herbivores
outer17_all_site <- subset(alpha_diversity_site, habitat == "outer_17") # only use rows that contain all factors
# remove nitrogent columns
outer17_all_site <- outer17_all_site[,-which(names(outer17_all_site) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
# remove other na rows
outer17_all_site <- na.omit(outer17_all_site)

# outer17_all_richness_mod_universal <- glmmTMB(richness ~ year  + wave_power_adcp  + mean_temp_c +  offset(log(hard_cover)) + # (1|site), family = poisson, data = outer17_all_site, na.action = na.pass) # universal model has convergence issues
# outer17_all_richness_dredge <- dredge(outer17_all_richness_mod_universal) # most models have convergence issues
# model.sel(outer17_all_richness_dredge) # top model includes just intercept...

outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass)
outer17_all_shannon_dredge <- dredge(outer17_all_shannon_mod_universal) # some convergence issues
model.sel(outer17_all_shannon_dredge) #  best includes just year. next best also includes temperature and is almost equivalent AICc
outer17_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = outer17_all_site, na.action = na.pass)
qqnorm(residuals(outer17_all_shannon_top_mod)) # pretty good
plot(residuals(outer17_all_shannon_top_mod) ~ fitted(outer17_all_shannon_top_mod)) # looks fine
summary(outer17_all_shannon_top_mod)

```

```{r stepwise models site level macroalgae}

## FRINGING - exclude wave power
fringing_macro_site <- na.omit(subset(alpha_diversity_site_macro, habitat == "fringing")) # only use rows that contain all factors

fringing_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_macro_site, na.action = na.pass)
fringing_macro_shannon_dredge <- dredge(fringing_macro_shannon_mod_universal) 
model.sel(fringing_macro_shannon_dredge) # top model just year, next best includes temp
fringing_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year + (1|site), data = fringing_macro_site, na.action = na.pass)
qqnorm(residuals(fringing_macro_shannon_top_mod))# not bad
plot(residuals(fringing_macro_shannon_top_mod) ~ fitted(fringing_macro_shannon_top_mod)) # slight negative trend

## Backreef - includes all 
backreef_macro_site <- na.omit(subset(alpha_diversity_site_macro, habitat == "backreef")) # only use rows that contain all factors

backreef_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_macro_site, na.action = na.pass)
backreef_macro_shannon_dredge <- dredge(backreef_macro_shannon_mod_universal) 
model.sel(backreef_macro_shannon_dredge) #  best includes just wave power but estimate is tiny
backreef_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  wave_power_adcp + (1|site), data = backreef_macro_site, na.action = na.pass)
qqnorm(residuals(backreef_macro_shannon_top_mod)) #  a little wonky but not bad
plot(residuals(backreef_macro_shannon_top_mod) ~ fitted(backreef_macro_shannon_top_mod)) # looks fine
summary(backreef_macro_shannon_top_mod)
Anova(backreef_macro_shannon_top_mod) # lol X2 = 2.56, P = 0.1

backreef_emm <-as.data.frame(emmeans(backreef_macro_shannon_top_mod, ~wave_power_adcp,
                                     at = list(wave_power_adcp = seq(min(backreef_macro_site$wave_power_adcp),
                                                            max(backreef_macro_site$wave_power_adcp),
                                                            1000)))) 
ggplot()+
  geom_point(data = backreef_macro_site, aes(x = wave_power_adcp, y = shannon))+
  geom_line(data = backreef_emm, aes(x = wave_power_adcp, y = emmean)) +
  geom_ribbon(data = backreef_emm, aes(x = wave_power_adcp, ymin = lower.CL, ymax = upper.CL, alpha = 0.2), show.legend = F) +
  scale_y_log10() + 
  labs(x = "Wave power", y = "Backreef Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

## outer10 - no nitrogen 
outer10_macro_site <- subset(alpha_diversity_site_macro, habitat == "outer_10") # only use rows that contain all factors
# remove nitrogent columns
outer10_macro_site<- outer10_macro_site[,-which(names(outer10_macro_site) %in% c("mean_n", "n_cv"))]
# remove other na rows
outer10_macro_site <- na.omit(outer10_macro_site)

outer10_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_macro_site, na.action = na.pass)
outer10_macro_shannon_dredge <- dredge(outer10_macro_shannon_mod_universal) 
model.sel(outer10_macro_shannon_dredge) #  best includes wave power and year 
outer10_macro_shannon_top_mod <- glmmTMB(log(shannon + 1) ~ year  + (1|site), data = outer10_macro_site, na.action = na.pass)
qqnorm(residuals(outer10_macro_shannon_top_mod)) # pretty good, tails maybe off
plot(residuals(outer10_macro_shannon_top_mod) ~ fitted(outer10_macro_shannon_top_mod)) # looks fine
summary(outer10_macro_shannon_top_mod) # NaN?

## outer17 - no nitrogen or herbivores
outer17_macro_site<- subset(alpha_diversity_site_macro, habitat == "outer_17") # only use rows that contain all factors
# remove nitrogent columns
outer17_macro_site<- outer17_macro_site[,-which(names(outer17_macro_site) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
# remove other na rows
outer17_macro_site <- na.omit(outer17_macro_site)

outer17_macro_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year  + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_macro_site, na.action = na.pass)
outer17_macro_shannon_dredge <- dredge(outer17_macro_shannon_mod_universal) 
model.sel(outer17_macro_shannon_dredge) #  best includes all terms!
qqnorm(residuals(outer17_macro_shannon_mod_universal)) # a little hump shaped
plot(residuals(outer17_macro_shannon_mod_universal) ~ fitted(outer17_macro_shannon_mod_universal)) # looks fine
summary(outer17_macro_shannon_mod_universal) # also NaNs...

```

```{r synchrony diversity stability setup site level, fig.height = 8, fig.width= 8}

lm_synchrony_site <- codyn::synchrony(df = site_long_data,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "percent_cover",
                                   metric = "Loreau",
                                   replicate.var = "site_habitat") 
colnames(lm_synchrony_site) = c("site_habitat", "loreau_synchrony") # rename cols for merge later

lm_synchrony_site_macro <- codyn::synchrony(df = site_long_macro,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "percent_cover",
                                   metric = "Loreau",
                                   replicate.var = "site_habitat") 
colnames(lm_synchrony_site_macro) = c("site_habitat", "loreau_synchrony") # rename cols for merge later

diversity_cv_site <- alpha_diversity_site %>%
  dplyr::group_by(site, habitat, island_side, site_habitat) %>%
  dplyr::summarise(
            algae_richness_mean = mean(richness), 
            algae_richness_cv = CV(richness),
            algae_richness_stability = 1/CV(richness),
            algae_shannon_mean = mean(shannon), 
            algae_shannon_cv = CV(shannon),
            algae_shannon_stability = 1/CV(shannon), 
            algae_cover_mean = mean(all_algae_cover_hard),
            algae_cover_cv = CV(all_algae_cover_hard),
            algae_cover_stability = 1/CV(all_algae_cover_hard)) 
# an extra outer_10 was introduced for every site for some reason that is all NAs
# remove NAs and merge with synchrony
diversity_stability_site <- merge(lm_synchrony_site, na.omit(diversity_cv_site), by = "site_habitat")

diversity_cv_site_macro <- alpha_diversity_site_macro %>%
  dplyr::group_by(site, habitat, island_side, site_habitat) %>%
  dplyr::summarise(
            algae_richness_mean = mean(richness), 
            algae_richness_cv = CV(richness),
            algae_richness_stability = 1/CV(richness),
            algae_shannon_mean = mean(shannon), 
            algae_shannon_cv = CV(shannon),
            algae_shannon_stability = 1/CV(shannon), 
            algae_cover_mean = mean(macroalgae_cover_hard),
            algae_cover_cv = CV(macroalgae_cover_hard),
            algae_cover_stability = 1/CV(macroalgae_cover_hard)) 
# an extra outer_10 was introduced for every site for some reason that is all NAs
# remove NAs and merge with synchrony
diversity_stability_site_macro <- merge(lm_synchrony_site_macro, na.omit(diversity_cv_site_macro), by = "site_habitat")

# for plotting later
diversity_ranges_site <- diversity_stability_site %>%
  group_by(habitat) %>%
  summarise(min_richness = min(algae_richness_mean),
            max_richness = max(algae_richness_mean),
            min_shannon = min(algae_shannon_mean),
            max_shannon = max(algae_shannon_mean),
            min_loreau = min(loreau_synchrony),
            max_loreau = max(loreau_synchrony))

diversity_ranges_site_macro <- diversity_stability_site_macro %>%
  group_by(habitat) %>%
  summarise(min_richness = min(algae_richness_mean),
            max_richness = max(algae_richness_mean),
            min_shannon = min(algae_shannon_mean),
            max_shannon = max(algae_shannon_mean),
            min_loreau = min(loreau_synchrony),
            max_loreau = max(loreau_synchrony))

```

```{r site all diversity stability synchrony, fig.height = 8, fig.width=8}
richness_cv_site <- glmmTMB(algae_cover_cv ~ algae_richness_mean*habitat, family = Gamma(link = "log"),
                          data = diversity_stability_site)
r.squaredGLMM(richness_cv_site) # 0.49
summary(richness_cv_site)
hist(residuals(richness_cv_site)) # symmetrical
plot(residuals(richness_cv_site) ~ fitted(richness_cv_site)) # slightly heteroskedastic
Anova(richness_cv_site)
#                               Chisq Df Pr(>Chisq)   
# algae_richness_mean          5.9078  1   0.015074 * 
# habitat                     14.5560  3   0.002238 **
# algae_richness_mean:habitat  3.8220  3   0.281336

emtrends(richness_cv_site, pairwise ~ habitat, var = "algae_richness_mean") # no significant differences
richness_cv_emm_site <- emmip(richness_cv_site,
                        habitat ~ algae_richness_mean,
                        at = list(algae_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)

filtered_richness_effects_site <- richness_cv_emm_site %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(algae_richness_mean >= min_richness & algae_richness_mean <= max_richness)

(richness_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site,
             aes(x = algae_richness_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_richness_effects_site, aes(x = algae_richness_mean, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_richness_effects_site, aes(x = algae_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

## REPEAT WITH SHANNON
shannon_cv_site <- glmmTMB(algae_cover_cv ~ algae_shannon_mean*habitat, family = Gamma(link = "log"),
                          data = diversity_stability_site)
r.squaredGLMM(shannon_cv_site) # 0.42
summary(shannon_cv_site)
hist(residuals(shannon_cv_site)) # symmetrical
plot(residuals(shannon_cv_site) ~ fitted(shannon_cv_site)) # slightly heteroskedastic
Anova(shannon_cv_site)
#                              Chisq Df Pr(>Chisq)   
# algae_shannon_mean          1.5291  1   0.216254   
# habitat                    12.1017  3   0.007043 **
# algae_shannon_mean:habitat  2.8675  3   0.412512

emtrends(shannon_cv_site, pairwise ~ habitat, var = "algae_shannon_mean") # no significant differences
shannon_cv_emm_site <- emmip(shannon_cv_site,
                        habitat ~ algae_shannon_mean,
                        at = list(algae_shannon_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_shannon_effects_site <- shannon_cv_emm_site %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)

(shannon_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site,
             aes(x = algae_shannon_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_shannon_effects_site, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_shannon_effects_site, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

synchrony_cv_site <- glmmTMB(algae_cover_cv ~ loreau_synchrony*habitat, family = Gamma(link = "log"),
                          data = diversity_stability_site)
r.squaredGLMM(synchrony_cv_site) # 0.76
summary(synchrony_cv_site)
hist(residuals(synchrony_cv_site)) # symmetrical
plot(residuals(synchrony_cv_site) ~ fitted(synchrony_cv_site)) # slightly heteroskedastic
Anova(synchrony_cv_site)
#                           Chisq Df Pr(>Chisq)    
# loreau_synchrony         45.857  1  1.272e-11 ***
# habitat                  14.018  3    0.00288 ** 
# loreau_synchrony:habitat  2.456  3    0.48330

emtrends(synchrony_cv_site, pairwise ~ habitat, var = "loreau_synchrony") # no differences between habitats
synchrony_cv_emm_site <- emmip(synchrony_cv_site,
                        habitat ~ loreau_synchrony,
                        at = list(loreau_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synchrony_cv_effects_site <- synchrony_cv_emm_site %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(loreau_synchrony >= min_loreau & loreau_synchrony <= max_loreau)

(synchrony_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site,
             aes(x = loreau_synchrony, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_cv_effects_site, aes(x = loreau_synchrony, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_synchrony_cv_effects_site, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

synchrony_richness_site <- glmmTMB(loreau_synchrony  ~ algae_richness_mean*habitat, family = beta_family(link = "logit"),
                          data = diversity_stability_site)
# r.squaredLR(synchrony_richness_site)
summary(synchrony_richness_site)
hist(residuals(synchrony_richness_site)) # symmetrical
plot(residuals(synchrony_richness_site) ~ fitted(synchrony_richness_site)) # slightly heteroskedastic
Anova(synchrony_richness_site)
# 
#                               Chisq Df Pr(>Chisq)    
# algae_richness_mean         12.6670  1  0.0003722 ***
# habitat                      5.0218  3  0.1702104    
# algae_richness_mean:habitat  6.0890  3  0.1073579

emtrends(synchrony_richness_site, pairwise ~ habitat, var = "algae_richness_mean") # no differences between habitats
synchrony_richness_emm_site <- emmip(synchrony_richness_site,
                        habitat ~ algae_richness_mean,
                        at = list(algae_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)

filtered_synchrony_richness_effects_site <- synchrony_richness_emm_site %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(algae_richness_mean >= min_richness & algae_richness_mean <= max_richness)

(synchrony_richness_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site,
             aes(x = algae_richness_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_richness_effects_site, aes(x = algae_richness_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_synchrony_richness_effects_site, aes(x = algae_richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

synchrony_shannon_site <- glmmTMB(loreau_synchrony  ~ algae_shannon_mean*habitat, family = beta_family(link = "logit"),
                          data = diversity_stability_site)
# r.squaredLR(synchrony_shannon_site)
summary(synchrony_shannon_site)
hist(residuals(synchrony_shannon_site)) # symmetrical
plot(residuals(synchrony_shannon_site) ~ fitted(synchrony_shannon_site)) # slightly heteroskedastic
Anova(synchrony_shannon_site)
# 
#                              Chisq Df Pr(>Chisq)   
# algae_shannon_mean         10.5506  1   0.001161 **
# habitat                     4.9410  3   0.176171   
# algae_shannon_mean:habitat  6.1573  3   0.104205

emtrends(synchrony_shannon_site, pairwise ~ habitat, var = "algae_shannon_mean") # no differences between habitats
synchrony_shannon_emm_site <- emmip(synchrony_shannon_site,
                        habitat ~ algae_shannon_mean,
                        at = list(algae_shannon_mean = seq(0,1.5,0.01)), plotit = F, CIs = T)

filtered_synchrony_shannon_effects_site <- synchrony_shannon_emm_site %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)

(synchrony_shannon_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site,
             aes(x = algae_shannon_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_shannon_effects_site, aes(x = algae_shannon_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_synchrony_shannon_effects_site, aes(x = algae_shannon_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

```

```{r site macroalgae diversity stability synchrony, fig.height = 8, fig.width=8}
richness_cv_site_macro <- glmmTMB(algae_cover_cv ~ algae_richness_mean*habitat, family = Gamma(link = "log"),
                          data = diversity_stability_site_macro)
r.squaredGLMM(richness_cv_site_macro) # 0.65
summary(richness_cv_site_macro)
hist(residuals(richness_cv_site_macro)) # symmetrical
plot(residuals(richness_cv_site_macro) ~ fitted(richness_cv_site_macro)) # not bad
Anova(richness_cv_site_macro)
#                               Chisq Df Pr(>Chisq)   
# algae_richness_mean         15.772  1  7.144e-05 ***
# habitat                     21.689  3  7.572e-05 ***
# algae_richness_mean:habitat 12.244  3   0.006591 **

emtrends(richness_cv_site_macro, pairwise ~ habitat, var = "algae_richness_mean") # backreef differs from two forereef sites
richness_cv_emm_site_macro <- emmip(richness_cv_site_macro,
                        habitat ~ algae_richness_mean,
                        at = list(algae_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)

filtered_richness_effects_site_macro <- richness_cv_emm_site_macro %>%
  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
  filter(algae_richness_mean >= min_richness & algae_richness_mean <= max_richness)

(richness_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site_macro,
             aes(x = algae_richness_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_richness_effects_site_macro, aes(x = algae_richness_mean, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_richness_effects_site_macro, aes(x = algae_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

## REPEAT WITH SHANNON
shannon_cv_site_macro <- glmmTMB(algae_cover_cv ~ algae_shannon_mean*habitat, family = Gamma(link = "log"),
                          data = diversity_stability_site_macro)
r.squaredGLMM(shannon_cv_site_macro) # 0.37
summary(shannon_cv_site_macro)
hist(residuals(shannon_cv_site_macro)) # a little chunky
plot(residuals(shannon_cv_site_macro) ~ fitted(shannon_cv_site_macro)) # pretty good
Anova(shannon_cv_site_macro)
#                             Chisq Df Pr(>Chisq)  
# algae_shannon_mean         1.1533  1    0.28287  
# habitat                    7.9691  3    0.04666 *
# algae_shannon_mean:habitat 4.1867  3    0.24199

emtrends(shannon_cv_site_macro, pairwise ~ habitat, var = "algae_shannon_mean") # no significant differences
shannon_cv_emm_site_macro <- emmip(shannon_cv_site_macro,
                        habitat ~ algae_shannon_mean,
                        at = list(algae_shannon_mean = seq(0,2,0.01)), plotit = F, CIs = T)

filtered_shannon_effects_site_macro <- shannon_cv_emm_site_macro %>%
  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)

(shannon_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site_macro,
             aes(x = algae_shannon_mean, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_shannon_effects_site_macro, aes(x = algae_shannon_mean, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_shannon_effects_site_macro, aes(x = algae_shannon_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

synchrony_cv_site_macro <- glmmTMB(algae_cover_cv ~ loreau_synchrony*habitat, family = Gamma(link = "log"),
                          data = diversity_stability_site_macro)
r.squaredGLMM(synchrony_cv_site_macro) # 0.71
summary(synchrony_cv_site_macro)
hist(residuals(synchrony_cv_site_macro)) # symmetrical
plot(residuals(synchrony_cv_site_macro) ~ fitted(synchrony_cv_site_macro)) # slightly heteroskedastic
Anova(synchrony_cv_site_macro)
#                           Chisq Df Pr(>Chisq)    
# loreau_synchrony         22.575  1  2.021e-06 ***
# habitat                  11.611  3   0.008841 ** 
# loreau_synchrony:habitat 11.389  3   0.009799 **

emtrends(synchrony_cv_site_macro, pairwise ~ habitat, var = "loreau_synchrony") # backreef and fringe different from outer17
synchrony_cv_emm_site_macro <- emmip(synchrony_cv_site_macro,
                        habitat ~ loreau_synchrony,
                        at = list(loreau_synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synchrony_cv_effects_site_macro <- synchrony_cv_emm_site_macro %>%
  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
  filter(loreau_synchrony >= min_loreau & loreau_synchrony <= max_loreau)

(synchrony_cv_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site_macro,
             aes(x = loreau_synchrony, y = algae_cover_cv, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_cv_effects_site_macro, aes(x = loreau_synchrony, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_synchrony_cv_effects_site_macro, aes(x = loreau_synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

synchrony_richness_site_macro <- glmmTMB(loreau_synchrony  ~ algae_richness_mean*habitat, family = beta_family(link = "logit"),
                          data = diversity_stability_site_macro)
# r.squaredGLMM(synchrony_richness_site_macro)
summary(synchrony_richness_site_macro)
hist(residuals(synchrony_richness_site_macro)) # kind of chunky
plot(residuals(synchrony_richness_site_macro) ~ fitted(synchrony_richness_site_macro)) # slightly heteroskedastic
Anova(synchrony_richness_site_macro)
# 
#                               Chisq Df Pr(>Chisq)    
# algae_richness_mean         0.2299  1    0.63161  
# habitat                     5.4261  3    0.14312  
# algae_richness_mean:habitat 9.1650  3    0.02718 *

emtrends(synchrony_richness_site_macro, pairwise ~ habitat, var = "algae_richness_mean") # backreef almost diff from fing and outer17
synchrony_richness_emm_site_macro <- emmip(synchrony_richness_site_macro,
                        habitat ~ algae_richness_mean,
                        at = list(algae_richness_mean = seq(0,10,0.1)), plotit = F, CIs = T)

filtered_synchrony_richness_effects_site_macro <- synchrony_richness_emm_site_macro %>%
  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
  filter(algae_richness_mean >= min_richness & algae_richness_mean <= max_richness)

(synchrony_richness_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site_macro,
             aes(x = algae_richness_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_richness_effects_site_macro, aes(x = algae_richness_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_synchrony_richness_effects_site_macro, aes(x = algae_richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

synchrony_shannon_site_macro <- glmmTMB(loreau_synchrony  ~ algae_shannon_mean*habitat, family = beta_family(link = "logit"),
                          data = diversity_stability_site_macro)
summary(synchrony_shannon_site_macro)
hist(residuals(synchrony_shannon_site_macro)) # a little chunky
plot(residuals(synchrony_shannon_site_macro) ~ fitted(synchrony_shannon_site_macro)) # slightly heteroskedastic
Anova(synchrony_shannon_site_macro)
# 
#                              Chisq Df Pr(>Chisq)   
# algae_shannon_mean         0.6238  1     0.4296
# habitat                    4.8196  3     0.1855
# algae_shannon_mean:habitat 2.2517  3     0.5218

emtrends(synchrony_shannon_site_macro, pairwise ~ habitat, var = "algae_shannon_mean") # no differences between habitats
synchrony_shannon_emm_site_macro <- emmip(synchrony_shannon_site_macro,
                        habitat ~ algae_shannon_mean,
                        at = list(algae_shannon_mean = seq(0,1.5,0.01)), plotit = F, CIs = T)

filtered_synchrony_shannon_effects_site_macro <- synchrony_shannon_emm_site_macro %>%
  left_join(diversity_ranges_site_macro, by = c("habitat")) %>%
  filter(algae_shannon_mean >= min_shannon & algae_shannon_mean <= max_shannon)

(synchrony_shannon_plot = 
  ggplot() +
  geom_point(data = diversity_stability_site_macro,
             aes(x = algae_shannon_mean, y = loreau_synchrony, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = filtered_synchrony_shannon_effects_site_macro, aes(x = algae_shannon_mean, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = filtered_synchrony_shannon_effects_site_macro, aes(x = algae_shannon_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

```

```{r site all turnover}
total_turnover_site<-
  turnover(
  df = site_long_data,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "site_habitat",
  metric = "total"
)
total_turnover_site$site_hab_year <- paste0(total_turnover_site$site_habitat, "_", total_turnover_site$year) # for merge
total_turnover_site <- merge(total_turnover_site, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )

total_turnover_site$turnover_trans <- sv_trans(total_turnover_site$total)

total_turnover_mod_site <- glmmTMB(turnover_trans ~ year*habitat + (1 | site), family = beta_family(), data = total_turnover_site)
summary(total_turnover_mod_site)
hist(residuals(total_turnover_mod_site)) # a little stumpy nut not bad
plot(residuals(total_turnover_mod_site) ~ fitted(total_turnover_mod_site)) # negative trend
Anova(total_turnover_mod_site)
emtrends(total_turnover_mod_site, pairwise ~ habitat, var = "year") # differences between backreef and fringing
total_turnover_emm_site <- emmip(total_turnover_mod_site,
                        habitat ~ year,
                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)

(turnover_site_plot <- 
  ggplot() +
  geom_point(data = total_turnover_site,
             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = total_turnover_emm_site, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = total_turnover_emm_site, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
                                                  fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "Total turnover\n") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

```
```{r site macro turnover}
total_turnover_site_macro<-
  turnover(
  df = site_long_macro,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "site_habitat",
  metric = "total"
)
total_turnover_site_macro$site_hab_year <- paste0(total_turnover_site_macro$site_habitat, "_", total_turnover_site_macro$year) # for merge
total_turnover_site_macro <- merge(total_turnover_site_macro, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )

total_turnover_site_macro$turnover_trans <- sv_trans(total_turnover_site_macro$total)

total_turnover_mod_site_macro <- glmmTMB(turnover_trans ~ year*habitat + (1 | site), family = beta_family(), data = total_turnover_site_macro)
summary(total_turnover_mod_site_macro)
hist(residuals(total_turnover_mod_site_macro)) # slight negative skew
plot(residuals(total_turnover_mod_site_macro) ~ fitted(total_turnover_mod_site_macro)) # negative trend
Anova(total_turnover_mod_site_macro)
#               Chisq Df Pr(>Chisq)    
# year         73.936  1  < 2.2e-16 ***
# habitat      19.201  3  0.0002485 ***
# year:habitat 12.099  3  0.0070512 ** 
emtrends(total_turnover_mod_site_macro, pairwise ~ habitat, var = "year") # differences between forereef sites and fringing
total_turnover_emm_site_macro <- emmip(total_turnover_mod_site_macro,
                        habitat ~ year,
                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)

(turnover_site_macro_plot <- 
  ggplot() +
  geom_point(data = total_turnover_site_macro,
             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = total_turnover_emm_site_macro, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = total_turnover_emm_site_macro, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
                                                  fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "Total turnover\n") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

```

```{r site all appearance}
appearance_site<-
  turnover(
  df = site_long_data,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "site_habitat",
  metric = "appearance"
)
appearance_site$site_hab_year <- paste0(appearance_site$site_habitat, "_", appearance_site$year) # for merge
appearance_site <- merge(appearance_site, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )

appearance_site$turnover_trans <- sv_trans(appearance_site$appearance)

appearance_mod_site <- glmmTMB(turnover_trans ~ year*habitat , family = beta_family(), data = appearance_site)
# convergence issue from inclusion of random effect
summary(appearance_mod_site)
hist(residuals(appearance_mod_site)) # a little stumpy nut not bad
plot(residuals(appearance_mod_site) ~ fitted(appearance_mod_site)) # negative trend
Anova(appearance_mod_site)
#                Chisq Df Pr(>Chisq)    
# year         58.4417  1  2.094e-14 ***
# habitat       9.5344  3  0.0229685 *  
# year:habitat 18.4862  3  0.0003491 ***
emtrends(appearance_mod_site, pairwise ~ habitat, var = "year") # lagoons both diff from forereefs but not each other
appearance_emm_site <- emmip(appearance_mod_site,
                        habitat ~ year,
                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)

(appearance_site_plot <- 
  ggplot() +
  geom_point(data = appearance_site,
             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = appearance_emm_site, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = appearance_emm_site, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
                                                  fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "Appearances\n") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

```
```{r site macro disappearance}
appearance_site_macro<-
  turnover(
  df = site_long_macro,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "site_habitat",
  metric = "appearance"
)
appearance_site_macro$site_hab_year <- paste0(appearance_site_macro$site_habitat, "_", appearance_site_macro$year) # for merge
appearance_site_macro <- merge(appearance_site_macro, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )

appearance_site_macro$turnover_trans <- sv_trans(appearance_site_macro$appearance)

appearance_mod_site_macro <- glmmTMB(turnover_trans ~ year*habitat, family = beta_family(), data = appearance_site_macro) # same convergence issue
summary(appearance_mod_site_macro)
hist(residuals(appearance_mod_site_macro)) # stumpy
plot(residuals(appearance_mod_site_macro) ~ fitted(appearance_mod_site_macro)) # negative trend
Anova(appearance_mod_site_macro)
#               Chisq Df Pr(>Chisq)    
# year         45.8213  1  1.296e-11 ***
# habitat       5.8339  3   0.119979    
# year:habitat 13.7486  3   0.003268 ** 
emtrends(appearance_mod_site_macro, pairwise ~ habitat, var = "year") # fringing different from fore, backreef from outer10
appearance_emm_site_macro <- emmip(appearance_mod_site_macro,
                        habitat ~ year,
                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)

(turnover_site_macro_plot <- 
  ggplot() +
  geom_point(data = appearance_site_macro,
             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = appearance_emm_site_macro, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = appearance_emm_site_macro, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
                                                  fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "Appearances\n") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

```

```{r site all disappearance}
disappearance_site<-
  turnover(
  df = site_long_data,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "site_habitat",
  metric = "disappearance"
)
disappearance_site$site_hab_year <- paste0(disappearance_site$site_habitat, "_", disappearance_site$year) # for merge
disappearance_site <- merge(disappearance_site, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )

disappearance_site$turnover_trans <- sv_trans(disappearance_site$disappearance)

disappearance_mod_site <- glmmTMB(turnover_trans ~ year*habitat + (1|site), family = beta_family(), data = disappearance_site)
# no convergence issue
summary(disappearance_mod_site)
hist(residuals(disappearance_mod_site)) # positive skew
plot(residuals(disappearance_mod_site) ~ fitted(disappearance_mod_site)) # negative trend
Anova(disappearance_mod_site)
#                Chisq Df Pr(>Chisq)    
# year         10.402  1   0.001259 ** 
# habitat      22.685  3  4.698e-05 ***
# year:habitat 12.099  3   0.007053 ** 
emtrends(disappearance_mod_site, pairwise ~ habitat, var = "year") # backrreef different from outer 10
disappearance_emm_site <- emmip(disappearance_mod_site,
                        habitat ~ year,
                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)

(disappearance_site_plot <- 
  ggplot() +
  geom_point(data = disappearance_site,
             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = disappearance_emm_site, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = disappearance_emm_site, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
                                                  fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "Disappearances\n") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

```
```{r site all turnover}
disappearance_site_macro<-
  turnover(
  df = site_long_macro,
  time.var = "year",
  species.var = "taxa",
  abundance.var = "percent_cover",
  replicate.var = "site_habitat",
  metric = "disappearance"
)
disappearance_site_macro$site_hab_year <- paste0(disappearance_site_macro$site_habitat, "_", disappearance_site_macro$year) # for merge
disappearance_site_macro <- merge(disappearance_site_macro, site_meta_df, by = c("year", "site_habitat", "site_hab_year") )

disappearance_site_macro$turnover_trans <- sv_trans(disappearance_site_macro$disappearance)

disappearance_mod_site_macro <- glmmTMB(turnover_trans ~ year*habitat + (1|site), family = beta_family(), data = disappearance_site_macro) # same here
summary(disappearance_mod_site_macro)
hist(residuals(disappearance_mod_site_macro)) # not bad
plot(residuals(disappearance_mod_site_macro) ~ fitted(disappearance_mod_site_macro)) # negative trend
Anova(disappearance_mod_site_macro)
#               Chisq Df Pr(>Chisq)    
#                Chisq Df Pr(>Chisq)    
# year         33.3709  1  7.616e-09 ***
# habitat      11.5079  3   0.009274 ** 
# year:habitat  1.5899  3   0.661677    
emtrends(disappearance_mod_site_macro, pairwise ~ habitat, var = "year") # no sig difference
disappearance_emm_site_macro <- emmip(disappearance_mod_site_macro,
                        habitat ~ year,
                        at = list(year = seq(2008,2023,1)), plotit = F, CIs = T)

(turnover_site_macro_plot <- 
  ggplot() +
  geom_point(data = disappearance_site_macro,
             aes(x = year, y = turnover_trans, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = disappearance_emm_site_macro, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = disappearance_emm_site_macro, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL),
                                                  fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "Disappearances\n") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

```

```{r all algae alpha diversity quadrat level}

full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site/location) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_quad[!is.na(alpha_diversity_quad$macroalgae_cover_hard),]) 
summary(full_richness_mod)
hist(residuals(full_richness_mod)) # slightly skewed but not terrible

# striping/patterining expected because of discrete response var. otherwise fine
plot(residuals(full_richness_mod) ~ fitted(full_richness_mod))

emtrends(full_richness_mod, pairwise ~  habitat, var = "year")
full_richness_emm <- emmip(full_richness_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad,
             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.05) + 
  geom_line(data = full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Richness") +
  scale_y_log10() +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```

```{r macroalgae diversity quadrat level}

```