### 2024-09-13 julianna_gut_check.Rmd
Added a script to gut check site vs. quad level comparisons

---
title: "Gut check: site-level analyses"
output: html_document
date: "2024-09-13"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
```

# Load Callie's data

```{r}
synchrony <- read_csv(here("data/diversity_stability_synchrony.csv"))
alpha <- read_csv(here("data/alpha_diversity_quad_macro_with_cover.csv"))
```

# Summarize

```{r synchrony summary}
synchrony %>% 
  group_by(habitat, site) %>% 
  summarize(richness_mean_mean = mean(richness_mean), 
            cover_mean_mean = mean(cover_mean), 
            cover_cv_mean = mean(cover_cv), 
            cover_stability_mean = mean(cover_stability), 
            synchrony_trans_mean = mean(synchrony_trans)) -> synchrony_mean
```

```{r alpha summary}
alpha %>% 
  group_by(year, habitat, site) %>% 
  summarize(richness_mean = mean(richness), 
            macroalgal_prop_cover_mean = mean(macroalgal_prop_cover), 
            cover_trans_mean = mean(cover_trans)) -> alpha_mean
```

# Plot synchrony

```{r}
synchrony_mean
```