### 2024-09-24 Personal_Prelim_Code/dataexploration.Rmd
I HAVE MOVED YOUR PERSONAL CODE.
I did not delete it, but it is in a new folder. your file paths may not be happy, but the code is still there.

---
title: "data_exploration"
author: "L. Enright"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(dplyr)
library(tidyverse)
#install.packages("fishualize")
library(fishualize)

data <- read.csv(here("data", "allbenthicdata_cleaned.csv"), stringsAsFactors = F) # replacements don't work when strings are factors

str(data)

#pivot long

long_data <- pivot_longer(
  data = data,
  cols = 8:92,
  names_to = "taxa",
  values_to = "percent_cover")

```

#look at sites by location (back, fringing, forereef)
```{r backreef}

transect_level <- long_data %>%
  #dplyr::filter(Habitat == "Backreef") %>%
  dplyr::group_by(Year, Site, Habitat, Transect, taxa) %>%
  summarise(mean_cover = mean(percent_cover), 
            se = sd(percent_cover)/sqrt(n()))

site_level <- transect_level %>%
  dplyr::group_by(Year, Site, Habitat, taxa) %>%
  summarise(mean_site_cover = mean(mean_cover), 
            site_se = sd(mean_cover)/sqrt(n()))

#backreef
site_level %>%
  dplyr::filter(taxa %in% c("Coral", "Algal.Turf", "Crustose.Corallines", "Lobophora.variegata", 
                            "Asparagopsis.taxiformis")) %>%
  dplyr::filter(Habitat == "Backreef") %>%
  #dplyr::filter(Site == "LTER 1") %>%
  ggplot(aes(x = Year, y = mean_site_cover, color = taxa, group = taxa)) +
  geom_point() +
  scale_color_fish_d(option = "Coryphaena_hippurus")+
  geom_line() +
  geom_errorbar(aes(ymax = mean_site_cover + site_se,
                ymin = mean_site_cover - site_se)) +
  facet_wrap(~Site)

site_level %>%
  dplyr::filter(taxa %in% c("Algal.Turf")) %>%
  dplyr::filter(Habitat == "Backreef") %>%
   dplyr::filter(Site == "LTER 1")
```

```{r fringing}

#fringing

site_level %>%
  dplyr::filter(taxa %in% c("Coral", "Algal.Turf", "Crustose.Corallines", "Lobophora.variegata", 
                            "Asparagopsis.taxiformis")) %>%
  dplyr::filter(Habitat == "Fringing") %>%
  #dplyr::filter(Site == "LTER 1") %>%
  ggplot(aes(x = Year, y = mean_site_cover, color = taxa, group = taxa)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = mean_site_cover + site_se,
                ymin = mean_site_cover - site_se)) +
  facet_wrap(~Site) +
  theme_classic()
```

```{r forereef 10}

#forereef

site_level %>%
  dplyr::filter(taxa %in% c("Coral", "Algal.Turf", "Crustose.Corallines", "Lobophora.variegata", 
                            "Asparagopsis.taxiformis")) %>%
  dplyr::filter(Habitat == "Outer 10") %>%
  #dplyr::filter(Site == "LTER 1") %>%
  ggplot(aes(x = Year, y = mean_site_cover, color = taxa, group = taxa)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = mean_site_cover + site_se,
                ymin = mean_site_cover - site_se)) +
  facet_wrap(~Site)

```

```{r forereef 17}

#forereef

site_level %>%
  dplyr::filter(taxa %in% c("Coral", "Algal.Turf", "Crustose.Corallines", "Lobophora.variegata", 
                            "Asparagopsis.taxiformis", "Bare.Space")) %>%
  dplyr::filter(Habitat == "Outer 17") %>%
  #dplyr::filter(Site == "LTER 1") %>%
  ggplot(aes(x = Year, y = mean_site_cover, color = taxa, group = taxa)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = mean_site_cover + site_se,
                ymin = mean_site_cover - site_se)) +
  facet_wrap(~Site)

```