### 2024-09-24 Personal_Prelim_Code/julianna_checkLTER1fringe.Rmd

---
title: "julianna_check2017fringeLTER1"
output: html_document
date: "2024-09-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load packages
require(here) # relative file paths
require(tidyverse) # data wrangling
require(vegan) # community/diversity analyses
```

This script summarizes the coral and macroalgae subsets to be joinable with the full algae dataset

# Bring in the data

```{r}
# algal cover 
algae_raw <- read_csv(here("data/MCR benthic algae data cover RAW.csv"))

```

```{r}
algae_raw %>% 
  filter(Taxonomy_Substrate_Functional_Group %in% c("Ascidian", "Sponge", "Millepora platyphylla", "Shell Debris", "Soft Coral", "Corallimorpharia", "Cyanophyta", "Coral Rubble", "No data", "Heteractis sp.", "Tridacna sp.", "Bare Space", "Sarcophyton sp.", "Zooxanthellae")) %>% 
  group_by(Year, Site, Location, Taxonomy_Substrate_Functional_Group) %>% 
  summarize(Total_other_cover = sum(Percent_Cover)) %>% 
  mutate(Habitat = case_when(str_detect(string = Location, pattern = "Fring") ~ "Fringing",
                              TRUE ~ "nawh")) %>% 
  filter(Year == 2017,
         Habitat == "Fringing",
         Site == "LTER 1"
         ) %>% view()

```