### 2024-09-24 Personal_Prelim_Code/julianna_figs_Aug29.Rmd

---
title: "Juilanna's figures for August 29"
output: html_document
date: "2024-08-16"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean up
rm(list=ls())

# load librarian (for repeatable package loading/installation)
if (!require("librarian")) install.packages("librarian")

# load packages
librarian::shelf(
  here, # relative file paths
  tidyverse, # data wrangling
  vegan, # community/diversity analyses
  MoMAColors # color palette: can install using devtools::install_github("BlakeRMills/MoMAColors")
                 )
```

# Read in data

Want data at the quadrat-level data for the trends through time plot.

```{r load data}
# raw algae information (e.g., for grabbing CCA for Fig 1.)
algae <- read_csv(here("data/algaeonly_cleaned_colnames.csv"))

# algae standardized dataset with diversity information at the quadrat-level (generated using julianna_algalMetric_wrangling.Rmd)
algaeQuad <- read_csv(here("data/algae_diversity_quadrat.csv"))

# macroalgae & coral dataset in the ancillary data Drive (generated using julianna_coral_macroalg_wrangling.Rmd)
macroalgQuad <- read_csv(here("data/macroalgae_coral_diversity_quadrat.csv"))

# all time metrics for stability
algaeAT <- read_csv(here("data/algae_diversity_stability_metrics_alltime.csv")) 
macroalgAT <- read_csv(here("data/macroalgae_coral_diversity_stability_metrics_alltime.csv")) 
```

# Patterns of 4 key spaceholders through time

Deron suggested we make this figure: macroalgae, CCA, turf, and coral through time

```{r get a dataset with the main four spaceholders}
macroalgQuad %>% 
  select(year, location, site, habitat, transect, quadrat, island_side, macroalgae_cover, macroalgae_cover_hard, coral_cover, coral_cover_hard, sand_cover) %>% 
  full_join(algae) %>%
  rowwise() %>% 
  # get turf and CCA cover
  mutate(turf_cover = sum(algal_turf, damselfish_turf),
         turf_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (turf_cover/(100-sand_cover))*100 
           ),
         CCA_cover = sum(lithophyllum_kotschyanum, crustose_corallines, neogoniolithon_brassica_florida, sporolithon_sp),
         CCA_cover_hard = case_when(
           sand_cover == 100 ~ NA, # since we can't divide by a zero
           TRUE ~ (CCA_cover/(100-sand_cover))*100 
           )
         ) %>% 
  select(year, location, site, habitat, transect, quadrat, island_side, 
         macroalgae_cover, macroalgae_cover_hard, coral_cover, coral_cover_hard, sand_cover, 
         turf_cover, turf_cover_hard, CCA_cover, CCA_cover_hard, sand_cover
         ) %>% 
  # relevel for graphing
  mutate(island_side = str_to_sentence(island_side),
         island_side = factor(island_side, levels = c("North", "East", "West")),
         habitat = case_when(
           str_detect(habitat, pattern = "_") ~ str_to_sentence(paste0(str_split(habitat, "_")[[1]][1], " ", str_split(habitat, "_")[[1]][2])),
           TRUE ~ str_to_sentence(habitat)),
         habitat = factor(habitat, levels = c("Fringing", "Backreef", "Outer 10", "Outer 17"))) %>% 
  mutate(site = case_when(
    site == "lter_1" ~ "LTER 1 (N)",
    site == "lter_2" ~ "LTER 2 (N)",
    site == "lter_3" ~ "LTER 3 (E)",
    site == "lter_4" ~ "LTER 4 (E)",
    site == "lter_5" ~ "LTER 5 (W)",
    site == "lter_6" ~ "LTER 6 (W)"
  )) -> main4quad
```

```{r plot for absolute cover}
main4quad %>% 
  select(year, location, site, habitat, transect, quadrat, island_side, 
         macroalgae_cover, turf_cover, CCA_cover, coral_cover, sand_cover) %>% 
  # pivot key columns longer
  pivot_longer(cols = c(macroalgae_cover, turf_cover, CCA_cover, coral_cover, sand_cover),
               names_to = "Substrate",
               values_to = "cover") %>% 
  # summarize and group quadrats by island/habitat
  group_by(year, habitat, site, island_side, Substrate) %>% 
  summarise(mean_cov = mean(cover),
            sd_cov = sd(cover, na.rm = TRUE),
            n = n(),
            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
  # plot
  ggplot(aes(x = year, y = mean_cov, group = Substrate, fill = Substrate)) +
  geom_ribbon(aes(ymin = mean_cov -CI95, 
                    ymax = mean_cov + CI95),
              alpha  = 0.7) +
  geom_line() +
  theme_bw() +
  # scale_fill_manual(values = c(
  #  moma.colors("Klein", type = "discrete")[2], # CCA
  #  moma.colors("Klein", type = "discrete")[1], # coral
  #  moma.colors("Klein", type = "discrete")[5], # macro
  #  moma.colors("Klein", type = "discrete")[4] # turf
  #),
  #labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
  facet_grid(site ~ habitat) + # not sure which direction is best here
  ylab("Mean absolute cover (%)") +
  xlab("Year") +
  # add lines for disturbances
  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
  geom_vline(xintercept = 2019, linetype = "dashed", color = "red") 
```

```{r plot for percent of hard cover}
main4quad %>% 
  select(year, location, site, habitat, transect, quadrat, island_side, 
         macroalgae_cover_hard, turf_cover_hard, CCA_cover_hard, coral_cover_hard) %>% 
  # pivot key columns longer
  pivot_longer(cols = c(macroalgae_cover_hard, turf_cover_hard, CCA_cover_hard, coral_cover_hard),
               names_to = "Substrate",
               values_to = "cover") %>% 
  # summarize and group quadrats by island/habitat
  group_by(year, habitat, site, island_side, Substrate) %>% 
  summarise(mean_cov = mean(cover, na.rm = TRUE),
            sd_cov = sd(cover, na.rm = TRUE),
            n = n(),
            CI95 = 1.96*(sd_cov)/sqrt(n)) %>% 
  # plot
  ggplot(aes(x = year, y = mean_cov, group = Substrate, fill = Substrate)) +
  geom_ribbon(aes(ymin = mean_cov -CI95, 
                    ymax = mean_cov + CI95),
              alpha  = 0.7) +
  geom_line() +
  theme_bw() +
   scale_fill_manual(values = c(
    moma.colors("Klein", type = "discrete")[2], # CCA
    moma.colors("Klein", type = "discrete")[1], # coral
    moma.colors("Klein", type = "discrete")[5], # macro
    moma.colors("Klein", type = "discrete")[4] # turf
  ),
  labels = c("CCA", "Coral", "Macroalgae", "Turf algae")) +
  facet_grid(site ~ habitat) + # not sure which direction is best here
  ylab("Mean cover of hard substrate(%)") +
  xlab("Year") +
  # add lines for disturbances
  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
  geom_vline(xintercept = 2019, linetype = "dashed", color = "red") 
```

# Coral cover vs. algal metrics

```{r join coral and algae}
algaeQuad %>% 
  full_join(macroalgQuad) -> coralAlgaeQuad
```

```{r plot algal-richness vs. coral}
coralAlgaeQuad %>% 
  ggplot(aes(x = algae_richness, y = coral_cover_hard, color = island_side, group = algae_richness)) +
  geom_boxplot() + 
  facet_grid(habitat~site) +
  scale_color_manual(values = c("darkblue", "darkred", "darkgreen")) +
  theme_classic()
```

```{r plot macroalgal-richness vs. coral}
coralAlgaeQuad %>% 
  ggplot(aes(x = macroalgae_richness, y = coral_cover_hard, color = island_side, group = macroalgae_richness)) +
  geom_boxplot() + 
  facet_grid(habitat~site) +
  scale_color_manual(values = c("darkblue", "darkred", "darkgreen")) +
  theme_classic()
```

```{r plot macroalgal-richness vs. algal richness}
coralAlgaeQuad %>% 
  ggplot(aes(x = algae_richness, y = macroalgae_richness, color = island_side, group = algae_richness)) +
  geom_point() + 
  facet_grid(habitat~site) +
  scale_color_manual(values = c("darkblue", "darkred", "darkgreen")) +
  theme_classic()
```

# Beta diversity

```{r beta diversity with disturbances}
coralAlgaeQuad %>% 
  # first get average coral cover per year
  group_by(year, habitat, site) %>% 
  summarize(coral_cover_hard_mean = mean(coral_cover_hard, na.rm = TRUE)) %>% 
  full_join(coralAlgaeQuad) %>% 
  ggplot(aes(x = year, y = hab_yr_distcentroid, group = year)) +
  geom_boxplot() +
  ylab("Distance to centroid") +
  xlab("Year") + 
  theme_bw() +
  facet_grid(site~habitat) +
  # add lines for disturbances
  geom_vline(xintercept = 2006, linetype = "dotdash") + # start of COTS
  geom_vline(xintercept = 2010,  color = "darkgray") + # cyclone Oli
  geom_vline(xintercept = 2019, linetype = "dashed", color = "red")  + 
  # Custom the Y scales:
  scale_y_continuous(
    # Features of the first axis
    name = "Algal beta diversity",
    # Add a second axis and specify its features
    sec.axis = sec_axis( trans=~.*100, name="Coral (% hard)")
  ) +
  geom_line(aes(group = 1, x = year, y = coral_cover_hard_mean/100), size = 1, color = "coral")
```

# Algal stability vs. coral cover

Join both data frames

```{r join all time data frames}
macroalgAT %>% 
  full_join(algaeAT) -> alltime
```

Plot

```{r cover CV vs. coral cover}
alltime %>% 
  ggplot(aes(x = coral_cover_hard_mean, y = macroalgae_cover_hard_cv)) +
  geom_point(aes(color = island_side)) +
  facet_wrap(~habitat) +
  geom_smooth(method = "lm") +
  theme_bw()
```

```{r cover CV vs. coral cover CV}
alltime %>% 
  ggplot(aes(x = coral_cover_hard_cv, y = macroalgae_cover_hard_cv)) +
  geom_point(aes(color = island_side)) +
  facet_wrap(~habitat) +
  geom_smooth(method = "lm") +
  theme_bw()
```

```{r richness CV vs. coral cover CV}
alltime %>% 
  ggplot(aes(x = coral_cover_hard_cv, y = algae_richness_cv)) +
  geom_point(aes(color = island_side)) +
  facet_wrap(~habitat) +
  geom_smooth(method = "lm") +
  theme_bw()
```

```{r synchrony vs. coral cover}
alltime %>% 
  #filter(habitat == "outer_10" | habitat ==  "outer_17") %>% 
  ggplot(aes(x = coral_cover_hard_mean, y = loreau_synchrony)) +
  geom_point(aes(color = island_side)) +
  #facet_wrap(~habitat) +
  geom_smooth(method = "lm") +
  theme_bw()
```

```{r synchrony vs. coral cover CV}
alltime %>% 
  ggplot(aes(x = coral_cover_hard_cv, y = loreau_synchrony)) +
  geom_point(aes(color = island_side)) +
  facet_wrap(~habitat) +
  geom_smooth(method = "lm") +
  theme_bw()
```

```{r portfolio effects vs. coral cover CV}
alltime %>% 
  ggplot(aes(x = coral_cover_hard_cv, y = portfolio_effect)) +
  geom_point(aes(color = island_side)) +
  facet_wrap(~habitat) +
  geom_smooth(method = "lm") +
  theme_bw()
```