### 2024-10-02 functional_groups_v3.Rmd
Add files via upload

---
title: "functional_groups"
author: "L. Enright"
date: "2024-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(janitor)
library(stringr)
library(vegan)
library(car)
library(glmmTMB)
library(MuMIn)
library(emmeans)
library(performance)
library(lme4)
library(DHARMa)
library(ggplot2)
```

```{r}
#read in groupings
#these are the OG groupings 
algae_fg_raw <- read_csv(here("data","algae_functional_groupings.csv"))

#these are our best guesses based on Bob's feedback (subject to change)
algae_fg_raw2 <- read_csv(here("data","algae_functional_groupings_092724.csv"))

#clean column names
algae_fg <- algae_fg_raw2 %>%
  clean_names()

#change species names to be lowercase & have underscore 
algae_fg$species <- gsub(" ", "_", tolower(algae_fg$species))
algae_fg$morphology <- trimws(algae_fg$morphology)
algae_fg$morphology <- gsub(" ", "_", algae_fg$morphology)

#rename species column to taxa

algae_fg <- algae_fg %>%
  select(species, morphology) %>%
  dplyr::rename(taxa = species) %>%
   dplyr::rename(functional_group = morphology)

unique(algae_fg$functional_group)

#9 fg in v1 (8 excluding turf)
#12 fg in v2 (10 excluding turf), added spherical and branched calcareous

```

```{r species distribution across functional groups}
algae_fg_raw %>%
  group_by(Morphology) %>%
  summarise(count = n())

algae_fg_raw2 %>%
  group_by(Morphology) %>%
  summarise(count = n())
```

```{r functions}
#this is needed for figures

inv_logit <- function(x){
  exp(x)/(1+exp(x))
}
```

```{r read in macroalgal data}
#read in long macro data 
macro_long <- read_csv(here("data", "macro_long_data.csv"))

#merge functional groups with 
macro_functional_groups <- merge(macro_long, algae_fg, all.x = TRUE)

unique(macro_functional_groups$taxa)

#macro_functional_groups$functional_group[NA]

#look for mistakes in merging 
macro_functional_groups %>% filter(is.na(functional_group))
#none, should be fine!

```

```{r summarizing long macroalgae data}

taxa_sum_check <- macro_functional_groups %>%
  group_by(location, year) %>%
  summarise(sum_macrocover = sum(prop_cover))

macro_functional_groups %>%
  filter(location == "lter_2_backreef_algae_transect_1_quad_1") %>%
  filter(year == "2022")

#summarizing functional groups at the quad level
fg_summary <- macro_functional_groups %>%
  group_by(location, year, functional_group, habitat, site) %>%
  summarise(sum_fg_cover = sum(prop_cover))

max(fg_summary$sum_fg_cover)

#summarizing functional groups at the site level
#NOTE that this is ADDED and not averaged. This data is only used to find richness at the site level. LE did this as adding and as a sum, both end up the same from a richness perspective. 
fg_summary_site_SUM <- macro_functional_groups %>%
  group_by(site, year, habitat, functional_group) %>%
  summarise(sum_fg_cover = sum(prop_cover))

#don't need to use this, was just to check both methods
#fg_summary_site_AVG <- macro_functional_groups %>%
  #group_by(site, year, habitat, functional_group) %>%
  #summarise(mean_fg_cover = mean(prop_cover))

#fg_summary %>%
  #filter(sum_fg_cover==1)

#QAQC
fg_summary_qaqc <- macro_functional_groups %>%
  group_by(location, year, functional_group) %>%
  summarise(sum_fg_cover = sum(prop_cover))

fg_qaqc <- fg_summary_qaqc %>%
  group_by(location, year) %>%
  summarise(sum_fg_check = sum(sum_fg_cover))

check_them <- merge(taxa_sum_check, fg_qaqc)

check_them %>%
  filter(all.equal(sum_macrocover, sum_fg_check) != TRUE)

#should be 0 :)

```

```{r info for themes and colors for figures}
mytheme <- theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))

habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")
```

```{r functional group data, pivot wide for quad}
#pivot wide at quad level
fg_summary_wide <- fg_summary %>%
  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")

#richness at every quad/year combination 
fg_summary_wide$fg_richness_quad <- vegan::specnumber(fg_summary_wide[5:14])

unique(fg_summary_wide$fg_richness_quad)
#0 to 4

#pull out the habitat and site, still at quad level 
fg_summary_wide_meta <- fg_summary_wide %>%
  mutate(habitat = str_extract(location, 
                                 "(fringing|outer_10|outer_17|backreef)")) %>%
  mutate(site = str_extract(location, "lter_[123456]"))
```

```{r functional group data, pivot wide for site}
#SUM
#pivot wide at site level
fg_summary_site_wide_SUM <- fg_summary_site_SUM %>%
  pivot_wider(names_from = "functional_group", values_from = "sum_fg_cover")

#richness at every site/year combination ---> NOT AVERAGED,
fg_summary_site_wide_SUM$fg_richness_site <- vegan::specnumber(fg_summary_site_wide_SUM[4:13])

fg_summary_site_wide <- fg_summary_site_wide_SUM %>%
   mutate(site_habitat = paste(site, habitat, sep = "_"))

#AVG -- don't need to use this, just was checking both methods 
#fg_summary_site_wide_AVG <- fg_summary_site_AVG %>%
  #pivot_wider(names_from = "functional_group", values_from = "mean_fg_cover")

#richness at every site/year combination ---> NOT AVERAGED 
#fg_summary_site_wide_AVG$fg_richness_site <- vegan::specnumber(fg_summary_site_wide_AVG[4:13])

#setdiff(fg_summary_site_wide_SUM$fg_richness_site, fg_summary_site_wide_AVG$fg_richness_site)

```

```{r functional richness through time at quad level}
fg_summary_wide_meta_through_time <- fg_summary_wide_meta %>%
  group_by(location, habitat, site) %>%
  summarise(fg_richness_mean = mean(fg_richness_quad),
            fg_richness_se = sd(fg_richness_quad)/n())

#each quad only has ONE value, averaged through time 
#1200 quads in this data set
#this is used in figure 4 things
```

# Figure 3

```{r figure 3b - setup dataframe quad}

#read in Noam's cover data 
alpha_div_cover <- read_csv(here("data", "alpha_diversity_quad_macro_with_cover_updated.csv"))

#merge Noam's data with functional group summary 
fig3df_quad <- merge(alpha_div_cover, fg_summary_wide, by = c("location", "year", "site"))

#choose habitat column & rename from Noam's cover data

fig3df_quad <- fig3df_quad %>%
  select(-habitat.y) %>%
  rename(habitat = habitat.x)

```

## Figure 3b model (functional group)

```{r figure 3b model - fg VS cover - quad}
cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_quad*habitat + (1|site/location) + (1|year), family = beta_family(), data = fig3df_quad)

hist(fig3df_quad$cover_trans)

summary(cover_mod_fg)
car::Anova(cover_mod_fg)
#                            Chisq Df Pr(>Chisq)    
#fg_richness_quad         31615.090  1  < 2.2e-16 ***
#habitat                     71.048  3  2.546e-15 ***
#fg_richness_quad:habitat   383.667  3  < 2.2e-16 ***
hist(residuals(cover_mod_fg)) # looks fine

# Let's check model diagnostics 
#step 1: linearity
#flat and horizontal --> linearity 
scatter.smooth(residuals(cover_mod_fg) ~ fitted(cover_mod_fg), col = "red")

#looks... fine?

#step 2: homosketadicity..

#standarized pearson's residualizuals

standard_mod3b <- (resid(cover_mod_fg, type = 'pearson')/sd(resid(cover_mod_fg, type = 'pearson')))

#took sqrt

sqrt_residuals_mod3b <-  sqrt(abs(standard_mod3b))

#plotted against fitted 
scatter.smooth(sqrt_residuals_mod3b ~ fitted(cover_mod_fg), col = "red")

#we have determined that this is ok, despite small hump and upward trend

#step 3: check uniformity of residuals plot

plot(performance::simulate_residuals(cover_mod_fg))

#gosh, these seems ... questionsable

r.squaredGLMM(cover_mod_fg) # 0.625, 0.645
emtrends(cover_mod_fg, pairwise ~ habitat, var = "fg_richness_quad") 
#all pairwise comparisons are different

```

## figure 3b plot setup
```{r figure 3b setup}
diversity_ranges_quad_fg <- fig3df_quad %>%
  group_by(habitat) %>%
  summarise(min_fg_richness_quad = min(fg_richness_quad),
            max_fg_richness_quad = max(fg_richness_quad),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

cover_emm <- emmip(cover_mod_fg,
                        habitat ~ fg_richness_quad,
                        at = list(fg_richness_quad = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects <- cover_emm %>%
  left_join(diversity_ranges_quad_fg, by = c("habitat")) %>%
  filter(fg_richness_quad >= min_fg_richness_quad & fg_richness_quad <= max_fg_richness_quad)

```

## figure 3b plot

```{r figure 3b plot}
(figure_3b <- 
  ggplot() +
  geom_point(data = fig3df_quad,
             aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Macroalgal Cover\n", x = "Functional Richness", title = "3b.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

#ggsave(filename = "output/figure_3b_quad_v2.png", figure_3b, height = 10, width = 14)
```

```{r figure 3e - setup dataframe site}

#read in Noam's site cover data

alpha_div_site_macro <- read_csv(here("data", "alpha_diversity_site_macro_with_cover.csv"))

#merge site level cover with site functional group richness

site_cover_fg_richness <- merge(alpha_div_site_macro, fg_summary_site_wide, by = c("year", "site_habitat", "site"))

#choose habitat column & rename from Noam's cover data

site_cover_fg_richness_df <- site_cover_fg_richness %>%
  select(-habitat.y) %>%
  rename(habitat = habitat.x)

```

### fig 3e site functional richness mod

```{r site functional richness mod - fig3e}
site_cover_mod_fg <- glmmTMB(cover_trans ~ fg_richness_site*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = site_cover_fg_richness_df)

summary(site_cover_mod_fg)
Anova(site_cover_mod_fg)

#                           Chisq Df Pr(>Chisq)    
#fg_richness_site         31.5959  1  1.898e-08 ***
#habitat                   4.9665  3     0.1743    
#fg_richness_site:habitat  4.3903  3     0.2223

hist(residuals(site_cover_mod_fg)) #looks bad
plot(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod_fg) # 0.1826, 0.5729
emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "fg_richness_site") # no difference between the habitats

#step 1: linearity

#flat and horizontal --> linearity 
scatter.smooth(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg), col = "red")
#scatter.smooth(resid(site_cover_mod_fg, type = 'pearson') ~ fitted(site_cover_mod_fg), col = "red")

#looks good

#step 2: homosketadicity..

#standarized residualized

standard_mod3e <- (resid(site_cover_mod_fg, type = 'pearson')/sd(resid(site_cover_mod_fg, type = 'pearson')))

#took sqrt
#https://boostedml.com/2019/03/linear-regression-plots-scale-location-plot.html

sqrt_residuals_mod3e <- sqrt(abs(standard_mod3e))

#plotted against fitted 
scatter.smooth(sqrt_residuals_mod3e ~ fitted(site_cover_mod_fg), col = "red")

#this is ok

#tells us these are uniform
plot(performance::simulate_residuals(site_cover_mod_fg))

#looks okay?

```

### 3e plot setup
```{r figure 3e setup }
diversity_ranges_site <- site_cover_fg_richness_df %>%
  group_by(habitat) %>%
  summarise(min_fg_richness_site = min(fg_richness_site),
            max_fg_richness_site = max(fg_richness_site),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

site_cover_emm <- emmip(site_cover_mod_fg,
                        habitat ~ fg_richness_site,
                        at = list(fg_richness_site = seq(0,17,1)), plotit = F, CIs = T)

filtered_site_cover_effects <- site_cover_emm %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(fg_richness_site >= min_fg_richness_site & fg_richness_site <= max_fg_richness_site)

```

### 3e plot

```{r figure 3e plot }
(figure_3e <- 
  ggplot() +
  geom_point(data = site_cover_fg_richness_df,
             aes(x = fg_richness_site, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
  geom_line(data = filtered_site_cover_effects, aes(x = fg_richness_site, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_site_cover_effects, aes(x = fg_richness_site, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "Functional Richness", title = "3e.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
#ggsave(filename = "output/figure_3e_v2.png", figure_3e, height = 10, width = 14)
```
#FIGURE 4

#Stability Quad Level

```{r figure 4b stabililty quad level}
#read in noam's stability data 
stability_data <- read_csv(here("data", "diversity_stability_synchrony.csv"))

#merge with fg data 
stability_quad_level <- merge(fg_summary_wide_meta_through_time, stability_data, 
                              by = 
                                c("location","site"))

#choose habitat column & rename from Noam's cover data

stability_quad_level_df <- stability_quad_level %>%
  select(-habitat.x) %>%
  rename(habitat = habitat.y)

#we lose 16 quads for this, 1184 instead of 1200 
```

```{r DSS - Stability}
dss_ranges_fg <- stability_quad_level_df %>%
  group_by(habitat) %>%
  summarise(min_richness_fg = min(fg_richness_mean),
            max_richness_fg = max(fg_richness_mean),
            min_stability= min(cover_stability, na.rm = TRUE),
            max_stability = max(cover_stability, na.rm = TRUE))
```

#figure 4b model
```{r fig4b fg richness stability mod}

rich_stab_mod_fg <- glmmTMB(cover_stability ~ fg_richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = stability_quad_level_df)

summary(rich_stab_mod_fg)
car::Anova(rich_stab_mod_fg)

#                            Chisq Df Pr(>Chisq)    
#fg_richness_mean         3025.235  1  < 2.2e-16 ***
#habitat                   122.635  3  < 2.2e-16 ***
#fg_richness_mean:habitat   20.721  3  0.0001203 ***

hist(residuals(rich_stab_mod_fg)) # ??
plot(residuals(rich_stab_mod_fg) ~ fitted(rich_synch_mod_fg)) #ugh 
r.squaredGLMM(rich_stab_mod_fg) # 0.815, 0.823
emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "fg_richness_mean") 
# backreef and fringing are different, fringing and outer10 are different, fringing and outer17 are different. (so, fringing is always differnt )

#step 1: linearity

#flat and horizontal --> linearity 
scatter.smooth(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg), col = "red")
#scatter.smooth(resid(site_cover_mod_fg, type = 'pearson') ~ fitted(site_cover_mod_fg), col = "red")

#looks good

#step 2: homosketadicity..

#standarized residualized

standard_mod4b <- (resid(rich_stab_mod_fg, type = 'pearson')/sd(resid(rich_stab_mod_fg, type = 'pearson')))

#took sqrt
#https://boostedml.com/2019/03/linear-regression-plots-scale-location-plot.html

sqrt_residuals_mod4b <- sqrt(abs(standard_mod4b))

#plotted against fitted 
scatter.smooth(sqrt_residuals_mod4b ~ fitted(rich_stab_mod_fg), col = "red")

#this is ok

#tells us these are uniform
plot(performance::simulate_residuals(rich_stab_mod_fg))

#yay, looks good!

performance::check_model(rich_stab_mod_fg)

```

```{r figure 4b setup}
rich_stab_emm_fg <- emmip(rich_stab_mod_fg,
                        habitat ~ fg_richness_mean,
                        at = list(fg_richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_fg <- rich_stab_emm_fg %>%
  left_join(dss_ranges_fg, by = c("habitat")) %>%
  filter(fg_richness_mean >= min_richness_fg & fg_richness_mean <= max_richness_fg)
```

```{r}
(figure_4b <- 
  ggplot() +
  geom_point(data = stability_quad_level_df,
             aes(x = fg_richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects_fg, aes(x = fg_richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Functional Richness", title = "4b.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18),
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

#ggsave(filename = "output/figure_4b_v2.png", figure_4b, height = 10, width = 14)
```

```{r functional group -- stability site}

stability_site <- read_csv(here("data", "diversity_stability_synchrony_site.csv"))
#n = 24

#need to get average functional richness through time! 
fg_summary_site_wide_summary <- fg_summary_site_wide %>%
  group_by(habitat, site, site_habitat) %>%
  summarize(avg_richness_fg = mean(fg_richness_site),
            se_richness_fg = sd(fg_richness_site)/sqrt(n()),
            n = n())

#n = 24!

stability_site_level_fg <- merge(stability_site, fg_summary_site_wide_summary, by = c("site_habitat", "site"))

#need to remove my habitat column and rename Noam's

stability_site_level_fg_df <- stability_site_level_fg %>%
  select(-habitat.y) %>%
  rename(habitat = habitat.x)

```

```{r}
dss_ranges_site_fg_stab <- stability_site_level_fg_df %>%
  group_by(habitat) %>%
  summarise(min_richness_fg = min(avg_richness_fg),
            max_richness_fg = max(avg_richness_fg),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability))
```

```{r fig 4e fg richness stability mod - site level }

rich_stab_mod_site_fg <- glmmTMB(cover_stability~ avg_richness_fg*habitat, family = Gamma(link = "log"), data = stability_site_level_fg_df)
summary(rich_stab_mod_site_fg)
Anova(rich_stab_mod_site_fg)

#                          Chisq Df Pr(>Chisq)   
#avg_richness_fg          3.8463  1   0.049856 * 
#habitat                 12.5935  3   0.005603 **
#avg_richness_fg:habitat  7.8799  3   0.048560 *

hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod_site_fg) # 0.671
emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "avg_richness_fg") 
# forereef 10 and fringing are different

#step 1: linearity

#flat and horizontal --> linearity 
scatter.smooth(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg), col = "red")
#scatter.smooth(resid(site_cover_mod_fg, type = 'pearson') ~ fitted(site_cover_mod_fg), col = "red")

#looks good

#step 2: homosketadicity..

#standarized residualized

standard_mod4e <- (resid(rich_stab_mod_site_fg, type = 'pearson')/sd(resid(rich_stab_mod_site_fg, type = 'pearson')))

#took sqrt
#https://boostedml.com/2019/03/linear-regression-plots-scale-location-plot.html

sqrt_residuals_mod4e <- sqrt(abs(standard_mod4e))

#plotted against fitted 
scatter.smooth(sqrt_residuals_mod4e ~ fitted(rich_stab_mod_site_fg), col = "red")

#this is ok

#tells us these are uniform
plot(performance::simulate_residuals(rich_stab_mod_site_fg))

#yay, looks good!

performance::check_model(rich_stab_mod_site_fg)

```

```{r figure 4e setup}
rich_stab_emm_site_fg <- emmip(rich_stab_mod_site_fg,
                        habitat ~ avg_richness_fg,
                        at = list(avg_richness_fg = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_site <- rich_stab_emm_site_fg %>%
  left_join(dss_ranges_site_fg_stab, by = c("habitat")) %>%
  filter((avg_richness_fg >= min_richness_fg) & (avg_richness_fg <= max_richness_fg))

```

```{r figure 4e}
(figure_4e <- 
  ggplot() +
  geom_point(data = stability_site_level_fg_df,
             aes(x = avg_richness_fg, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = avg_richness_fg, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Functional Richness", title = "4e.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

#ggsave(filename = "output/figure_4e_v2.png", figure_4e, height = 10, width = 14)
```

#STOP HERE