### 2025-01-21 SEM_site.Rmd

---
title: "SEM with site-level algae data"
output: html_document
date: "2025-01-21"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear workspace
rm(list = ls())

# load needed packages
require(tidyverse)
require(here)
require(piecewiseSEM)
require(lme4) # glmer() function
require(performance)
```

# Read in data

# Read in data

```{r}
# spatial synchrony data from Noam & functional richness--need to double check that FR is site-level
spatial_synch <- read_csv(here("data/spatial_synchrony.csv")) %>% 
  select(habitat, site, functional_richness_mean, spatial_synchrony)

# site-level data summarized with one point for all time per site
algae_site <- read_csv(here("data/diversity_stability_synchrony_site.csv")) %>% 
  full_join(spatial_synch)
```

# Scale data

Gamma families are currently not supported and we're not able to get glmmTMB to work for the time being.

```{r}
algae_site %>% 
  mutate(
    richness_mean_scale = scale(algae_site$richness_mean)[,1],
    functional_richness_mean_scale = scale(algae_site$functional_richness_mean)[,1],
    cover_mean_scale = scale(algae_site$cover_mean)[,1],
    cover_stability_scale = scale(algae_site$cover_stability)[,1],
    synchrony_scale = scale(algae_site$synchrony)[,1],
    spatial_synchrony_scale = scale(algae_site$spatial_synchrony)[,1]
  ) -> algae_site_scaled
```

# Make a simple SEM

First, fit individual models

```{r}
# effect of richness on synchrony
m.sprich.synch <- glmer(synchrony_mean_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_site_scaled)
  performance::check_model(m.sprich.synch)

# effect of synchrony on stability
m.synch.stab.cov <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale + spatial_synchrony_scale + (1|site) + (1|habitat),
                      data = algae_site_scaled)  
  performance::check_model(m.synch.stab.cov)

# effect of richness on cover
m.sprich.cov <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_site_scaled)
  performance::check_model(m.sprich.cov)
```

Make it with piecewiseSEM

```{r}
# Build SEM
sem <- psem(
  m.sprich.synch,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae_site_scaled)
)

summary(sem)
```

Plot

```{r}
plot(sem)
```