### 2025-01-24 SEM_quad_lavaan.Rmd

---
title: "SEM for quad-level data using lavaan"
output: html_document
date: "2025-01-22"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear workspace
rm(list = ls())

require(tidyverse)
require(here)
require(lavaan)
require(lavaanPlot)
```

# Read in data

```{r}
# functional richness from Noam
functrich <- read_csv(here("data/full_plot_level_dss.csv")) %>% 
  select(location, habitat, site, functional_richness_mean)

# quad-level data summarized with one point for all time
algae <- read_csv(here("data/diversity_stability_synchrony.csv")) %>% 
  inner_join(functrich)
```

# Scale data

```{r}
algae %>% 
  mutate(
    richness_mean_scale = scale(algae$richness_mean)[,1],
    functional_richness_mean_scale = scale(algae$functional_richness_mean)[,1],
    cover_mean_scale = scale(algae$cover_mean)[,1],
    cover_stability_scale = scale(algae$cover_stability)[,1],
    synchrony_scale = scale(algae$synchrony)[,1]
  ) %>% 
  select(site, habitat, richness_mean_scale, functional_richness_mean_scale,
         cover_mean_scale, cover_stability_scale, synchrony_scale
         ) -> algae_scaled
```

# Overall model

```{r define model}
# define the model
m.all <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale
  functional_richness_mean_scale ~ richness_mean_scale
  cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale
  cover_mean_scale ~ richness_mean_scale
'   
```

```{r examine fit}
fit.all <- lavaan::sem(m.all, cluster = "site", data = algae_scaled) 
  lavaan::summary(fit.all, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels1 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness", functional_richness_mean_scale = "Func. Richness", cover_stability_scale = "Stability", cover_mean_scale = "Cover")

# plot
lavaanPlot(model = fit.all,
           labels = labels1,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

# Overall modell: no functional richness

```{r define model}
# define the model
m.all.noFR <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale 
  cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale
  cover_mean_scale ~ richness_mean_scale
'   
```

```{r examine fit}
fit.noFR <- lavaan::sem(m.all.noFR, cluster = "site", data = algae_scaled) 
  lavaan::summary(fit.noFR, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels2 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness", cover_stability_scale = "Stability", cover_mean_scale = "Cover")

# plot
lavaanPlot(model = fit.noFR,
           labels = labels2,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

# Overall model: no cover

```{r define model}
# define the model
m.noC <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale
  functional_richness_mean_scale ~ richness_mean_scale
  cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + richness_mean_scale
'   
```

```{r examine fit}
fit.noC <- lavaan::sem(m.noC, cluster = "site", data = algae_scaled) 
  lavaan::summary(fit.noC, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels3 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness", functional_richness_mean_scale = "Func. Richness", cover_stability_scale = "Stability")

# plot
lavaanPlot(model = fit.noC,
           labels = labels3,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

# Overall model: no cover or functional richness

```{r define model}
# define the model
m.noCorFR <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale 
  cover_stability_scale ~ synchrony_scale + richness_mean_scale
'   
```

```{r examine fit}
fit.noCorF <- lavaan::sem(m.noCorFR, cluster = "site", data = algae_scaled) 
  lavaan::summary(fit.noCorF, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels4 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness",  cover_stability_scale = "Stability")

# plot
lavaanPlot(model = fit.noCorF,
           labels = labels4,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```

# Fringing model: no cover or functional richness

```{r define model}
# define the model
m.noCorFR.FR <- '
  # Structural model/regressions
  synchrony_scale ~ richness_mean_scale 
  cover_stability_scale ~  richness_mean_scale + synchrony_scale
'   
```

```{r examine fit}
fit.noCorF.FR <- lavaan::sem(m.noCorFR.FR, cluster = "site", data = algae_scaled) # [algae_scaled$habitat == "Fringing", ] 
  lavaan::summary(fit.noCorF.FR, standardized = TRUE, fit.measures = TRUE)
```

```{r}
# make labels
labels4 <- list(synchrony_scale = "Synchrony", richness_mean_scale = "Spp richness", cover_stability_scale = "Stability")

# plot
lavaanPlot(model = fit.noCorF.FR,
           labels = labels4,
           coefs = TRUE,
           stand = TRUE,
           sig = 0.05 # standardized regression paths, showing only paths with p<= .05
           )
```