### 2025-07-31 beta_div.Rmd
beta diversity code
This verison includes transformations and removal of rare species

---
title: "beta diversity"
author: "L. Enright"
date: "2025-06-10"
output: html_document
---

# read in the data. these are written out from previous scripts, need to rerun all of these scripts and then read back in the data, to make sure they are up to date.

if someone else other than lauren is running these, need to update data folder name
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(vegan)

macro_long_data <- read.csv(here::here("data_le", "macro_long_data.csv"), stringsAsFactors = F) 
diversity_stability_synchrony <- read.csv(here::here("data_le", "full_plot_level_dss.csv"), stringsAsFactors = F) 
diversity_stability_synchrony_site <- read.csv(here::here("data_le", "diversity_stability_synchrony_site.csv"), stringsAsFactors = F) 
spatial_synchrony <- read.csv(here::here("data_le", "spatial_synchrony.csv"), stringsAsFactors = F) 
mean_plot_level_dss <- read.csv(here::here("data_le", "mean_plot_level_dss.csv"), stringsAsFactors = F)

taxo <- read.csv(here::here("data_le", "MCR_Algal_Functional_Groups_02042025_updatedtax.csv"), stringsAsFactors = F)
```

# Data Prep
*** is it okay to leave the data in prop cover? instead of percent cover? leaving in prop cover for now. --> Noam says yes!

```{r}
# first, need to SUM the data across years for each quad 
macro_long_data_summed <- macro_long_data %>%
  group_by(location, site, habitat, site_habitat, transect, quadrat, taxa) %>%
  summarise(sum_taxa = sum(prop_cover)) %>%
  ungroup()
#should have 1 row for each site/habitat/transect/quad/taxa, summarized across years. 
```

# beta diversity - intital attempt
```{r}

#from long to wide
macro_wide_data_summed <- macro_long_data_summed %>%
   pivot_wider(names_from = taxa, values_from = sum_taxa, values_fill = 0)

#calculating beta diversity for individual sites/habitats so that I can check the foreloop is working.
#LTER 1 BR

macro_wide_data_summed_LTER1BR <- macro_wide_data_summed %>%
  filter(site_habitat == "lter_1_backreef") %>%
   filter(rowSums(.[, 7:68]) != 0)
# lost 1 :) --> this would be a quad that has no macroalgae in it 
#(lter_1_backreef_algae_transect_5_quad_7)

#for checking row sums
#macro_wide_data_summed_LTER1BR$row_sum <- rowSums(macro_wide_data_summed_LTER1BR[, 7:68])

sp_community_LTER1BR <- as.matrix(macro_wide_data_summed_LTER1BR[,c(7:68)])
meta_LTER1BR <-macro_wide_data_summed_LTER1BR[,c(1:6)]
sp_community_LTER1BR_transformed <- wisconsin(sp_community_LTER1BR)

rowSums(sp_community_LTER1BR)

#pairwise comparisons 
bray_LTER1BR <- vegdist(sp_community_LTER1BR, method = "bray")
bray_LTER1BR_transformed <- vegdist(sp_community_LTER1BR_transformed, method = "bray")

median_vegdist <- median(bray_LTER1BR)
#0.4913037
median_vegdist_tranformed <- median(bray_LTER1BR_transformed)
# 0.6590098
mean_vegdist <- mean(bray_LTER1BR)
#0.5195
mean_vegdist_tranformed <- mean(bray_LTER1BR_transformed)
# 0.6451
betadisp_LTER1BR <- betadisper(bray_LTER1BR, meta_LTER1BR$site_habitat)
#0.3596  
betadisp_LTER1BR_transformed <- betadisper(bray_LTER1BR_transformed, meta_LTER1BR$site_habitat)
#0.4521

#LTER 2 Outer 10

macro_wide_data_summed_LTER2OUT10 <- macro_wide_data_summed %>%
  filter(site_habitat == "lter_2_outer_10") %>%
   filter(rowSums(.[, 7:68]) != 0)
# lost 0  :)

sp_community_LTER2O10 <- as.matrix(macro_wide_data_summed_LTER2OUT10[,c(7:68)])
meta_LTER2O10 <-macro_wide_data_summed_LTER2OUT10[,c(1:6)]

rowSums(sp_community_LTER2O10)

#pairwise comparisons 
bray_LTER2O10 <- vegdist(sp_community_LTER2O10, method = "bray")
median_vegdist <- median(bray_LTER2O10)
#0.283019
betadisp_LTER2O10 <- betadisper(bray_LTER2O10, meta_LTER2O10$site_habitat)
#0.2041

```
## Good forloop for beta
this is good for calculating beta diversity for all 24 sites 
```{r}

# Initialize a results data frame
results_beta <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
   mean_bray = numeric(),
  mean_betadisp = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list <- list()

# Get unique site_habitat combinations
sites <- unique(macro_wide_data_summed$site_habitat)

for (sh in sites) {
  # Filter for one site/habitat
  df_sub <- macro_wide_data_summed %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:68]) != 0)

  # Skip if no rows left
  if (nrow(df_sub) < 2) {
    warning(paste("Skipping", sh, "- not enough data"))
    next
  }

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:68])
  meta <- df_sub[, 1:6]

  # Bray-Curtis distances
  bray <- vegdist(sp_community, method = "bray")
  median_bray <- median(bray) # i used median not mean and I am not sure why? should it be mean? 
  mean_bray <- mean(bray)

  # Beta dispersion
  betadisp <- betadisper(bray, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)

  betadisp_list[[sh]] <- betadisp

  # Store result
  results_beta <- rbind(results_beta, data.frame(
    site_habitat = sh,
    #betadisp_results = betadisp_results,
    median_bray = median_bray,
    mean_bray = mean_bray,
    mean_betadisp = mean_disp,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta)
betadisp_list

#betadisp list =       0.3596 LTER 1 BR
#betadisp list =       0.2041  LTER 2 FR10
#okay, these match with the single sites I did above, so I think the loop is working well. 
```

## plotting for beta 
```{r merge and plot}

# will eventually want to match Noam's nice themes, but just focusing on colors for now. 
habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

# merge together the beta div with the spatial synchrony output
beta_spatialsync <- merge(spatial_synchrony, results_beta, by = "site_habitat")

# some exploratory plots
beta_spatialsync %>%
  ggplot(aes(x = median_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

beta_spatialsync %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

beta_spatialsync %>%
  ggplot(aes(x = mean_betadisp, y = spatial_synchrony, color = site)) +
  geom_point() +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#patterns are similar whether we use median bray or mean bray or mean betadisp ... do we need to try with median bray?

#NO FACET
beta_spatialsync %>%
  ggplot(aes(x = median_bray, y = spatial_synchrony, color = habitat, shape = site)) +
  geom_point() +
   geom_smooth(method = "lm", se = TRUE, aes(group = habitat)) +
  scale_colour_manual(values = habitat_colours) +
  theme_minimal()
  #geom_line(method = lm)

# negative relationship between median bray and stability 
beta_spatialsync %>%
  ggplot(aes(x = median_bray, y = stability_mean, color = habitat, shape = site)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, aes(group = habitat)) +
  scale_colour_manual(values = habitat_colours) +
  theme_minimal()

   #geom_line(method = lm)

```

## probably the figure we want to use based on discussion 
shows a negative relationship, don't need to touch on habitat really 
```{r}
# need to make pretty
beta_spatialsync %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()
```
# Trying transformation with no removal of rare species

```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_transformed <- list()

# Get unique site_habitat combinations
sites <- unique(macro_wide_data_summed$site_habitat)

for (sh in sites) {
  # Filter for one site/habitat
  df_sub <- macro_wide_data_summed %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:68]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:68])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list[[sh]] <- betadisp
  betadisp_list_transformed[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed <- bind_rows(results_beta_transformed, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed)
betadisp_list_transformed

```

```{r}
beta_spatialsync_transformed <- merge(spatial_synchrony, results_beta_transformed, by = "site_habitat")

# some exploratory plots
# not transformed
beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#transformed
beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

beta_spatialsync_transformed %>%
  ggplot(aes(x = median_bray_trans, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_disp_transformed, y = spatial_synchrony, color = site)) +
  geom_point() +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#patterns are similar whether we use transformed median bray or transformed mean bray or transformed mean betadisp
# slightly different pattersn with the transformation

```

#Again, I think this is the figure we want to show. 
```{r}

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()
```
# Trying removing the species.

let's make a matrix for each habitat

## Fringing
```{r}

fringing_macro_wide_data_summed <- macro_wide_data_summed %>%
  filter(habitat == "Fringing")

#300 rows.. 
#make into a matrix
fringing_macro_wide_data_summed_matrix <- as.matrix(fringing_macro_wide_data_summed[, 7:68])

fringing_macro_wide_data_summed_meta <- fringing_macro_wide_data_summed[, 1:6]

#filter rare species
#how many quads out of 300 have a species
fringing_site_species_presence <- colSums(fringing_macro_wide_data_summed_matrix > 0) # Presence absence 
#divide by # of rows totla
fringing_site_species_percentage <- (fringing_site_species_presence/nrow(fringing_macro_wide_data_summed_matrix)) # proportion of rows where a species is present

fringing_site_common_species <- names(fringing_site_species_percentage[fringing_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...

fringing_site_species_percentage
fringing_site_common_species

#17 species

fringing_site_common_species_matrix <- fringing_macro_wide_data_summed_matrix[,colnames(fringing_macro_wide_data_summed_matrix) %in% fringing_site_common_species] # filter original species matrix
#okay this has 17

#now need to make it back into a dataframe for the forloop
fringing_site_common_species_df <- cbind(fringing_macro_wide_data_summed_meta, fringing_site_common_species_matrix)
#fringing_site_rows_to_remove <- rowSums(fringing_site_species_matrix) == 0 # remove rows with no algal species observed - thought this was done already
```

will need to do the forloops for each habitat - fringing
```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed_fringing_commonspecies <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_fringing_commonspecies <- list()
betadisp_list_transformed_fringing_commonspecies <- list()

# Get unique site_habitat combinations
sites_fring <- unique(fringing_site_common_species_df$site_habitat)

for (sh in sites_fring) {
  # Filter for one site/habitat
  df_sub <- fringing_site_common_species_df %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:23]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:23])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list_fringing_commonspecies[[sh]] <- betadisp
  betadisp_list_transformed_fringing_commonspecies[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed_fringing_commonspecies <- bind_rows(results_beta_transformed_fringing_commonspecies, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed_fringing_commonspecies)
betadisp_list_transformed_fringing_commonspecies

```
again, trying a one off to make sure foreloop is working 
```{r}
commonspecies_macro_LTER1FRI <- fringing_site_common_species_df %>%
  filter(site_habitat == "lter_1_fringing") %>%
   filter(rowSums(.[, 7:23]) != 0)

sp_community_LTER1FRI <- as.matrix(commonspecies_macro_LTER1FRI[,c(7:23)])
meta_LTER1FR <-commonspecies_macro_LTER1FRI[,c(1:6)]
sp_community_LTER1FRI_transformed <- wisconsin(sp_community_LTER1FRI)

rowSums(sp_community_LTER1BR)

#pairwise comparisons 
bray_LTER1FRI <- vegdist(sp_community_LTER1FRI, method = "bray")
bray_LTER1FRI_transformed <- vegdist(sp_community_LTER1FRI_transformed, method = "bray")

mean_vegdist <- mean(bray_LTER1FRI)
#0.3833
mean_vegdist_tranformed <- mean(bray_LTER1FRI_transformed)
#0.5812

#yay, they match with forloop output
```

## Backreef
```{r}

backreef_macro_wide_data_summed <- macro_wide_data_summed %>%
  filter(habitat == "Backreef")

#300 rows.. 
#make into a matrix
backreef_macro_wide_data_summed_matrix <- as.matrix(backreef_macro_wide_data_summed[, 7:68])

backreef_macro_wide_data_summed_meta <- backreef_macro_wide_data_summed[, 1:6]

#filter rare species
#how many quads out of 300 have a species
backreef_site_species_presence <- colSums(backreef_macro_wide_data_summed_matrix > 0) # Presence absence 
#divide by # of rows totla
backreef_site_species_percentage <- (backreef_site_species_presence/nrow(backreef_macro_wide_data_summed_matrix)) # proportion of rows where a species is present

backreef_site_common_species <- names(backreef_site_species_percentage[backreef_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...

backreef_site_species_percentage
backreef_site_common_species

#18 species

backreef_site_common_species_matrix <- backreef_macro_wide_data_summed_matrix[,colnames(backreef_macro_wide_data_summed_matrix) %in% backreef_site_common_species] # filter original species matrix
#okay this has 18

#now need to make it back into a dataframe for the forloop
backreef_site_common_species_df <- cbind(backreef_macro_wide_data_summed_meta, backreef_site_common_species_matrix)

```

will need to do the forloops for each habitat - backreef
```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed_backreef_commonspecies <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_backreef_commonspecies <- list()
betadisp_list_transformed_backreef_commonspecies <- list()

# Get unique site_habitat combinations
sites_back <- unique(backreef_site_common_species_df$site_habitat)

for (sh in sites_back) {
  # Filter for one site/habitat
  df_sub <- backreef_site_common_species_df %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:24]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:24])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list_backreef_commonspecies[[sh]] <- betadisp
  betadisp_list_transformed_backreef_commonspecies[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed_backreef_commonspecies <- bind_rows(results_beta_transformed_backreef_commonspecies, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed_backreef_commonspecies)
betadisp_list_transformed_backreef_commonspecies

```

## Forereef 10
```{r}

forereef10_macro_wide_data_summed <- macro_wide_data_summed %>%
  filter(habitat == "Forereef 10m")

#300 rows.. 
#make into a matrix
forereef10_macro_wide_data_summed_matrix <- as.matrix(forereef10_macro_wide_data_summed[, 7:68])

forereef10_macro_wide_data_summed_meta <- forereef10_macro_wide_data_summed[, 1:6]

#filter rare species
#how many quads out of 300 have a species
forereef10_site_species_presence <- colSums(forereef10_macro_wide_data_summed_matrix > 0) # Presence absence 
#divide by # of rows totla
forereef10_site_species_percentage <- (forereef10_site_species_presence/nrow(forereef10_macro_wide_data_summed_matrix)) # proportion of rows where a species is present

forereef10_site_common_species <- names(forereef10_site_species_percentage[forereef10_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...

forereef10_site_species_percentage
forereef10_site_common_species

#15 species

forereef10_site_common_species_matrix <- forereef10_macro_wide_data_summed_matrix[,colnames(forereef10_macro_wide_data_summed_matrix) %in% forereef10_site_common_species] # filter original species matrix
#okay this has 15

#now need to make it back into a dataframe for the forloop
forereef10_site_common_species_df <- cbind(forereef10_macro_wide_data_summed_meta, forereef10_site_common_species_matrix)

```

will need to do the forloops for each habitat - forereef10
```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed_forereef10_commonspecies <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_forereef10_commonspecies <- list()
betadisp_list_transformed_forereef10_commonspecies <- list()

# Get unique site_habitat combinations
sites_FO10 <- unique(forereef10_site_common_species_df$site_habitat)

for (sh in sites_FO10) {
  # Filter for one site/habitat
  df_sub <- forereef10_site_common_species_df %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:21]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:21])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list_forereef10_commonspecies[[sh]] <- betadisp
  betadisp_list_transformed_forereef10_commonspecies[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed_forereef10_commonspecies <- bind_rows(results_beta_transformed_forereef10_commonspecies, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed_forereef10_commonspecies)
betadisp_list_transformed_forereef10_commonspecies

```
## Forereef 17
```{r}

forereef17_macro_wide_data_summed <- macro_wide_data_summed %>%
  filter(habitat == "Forereef 17m")

#300 rows.. 
#make into a matrix
forereef17_macro_wide_data_summed_matrix <- as.matrix(forereef17_macro_wide_data_summed[, 7:68])

forereef17_macro_wide_data_summed_meta <- forereef17_macro_wide_data_summed[, 1:6]

#filter rare species
#how many quads out of 300 have a species
forereef17_site_species_presence <- colSums(forereef17_macro_wide_data_summed_matrix > 0) # Presence absence 
#divide by # of rows totla
forereef17_site_species_percentage <- (forereef17_site_species_presence/nrow(forereef17_macro_wide_data_summed_matrix)) # proportion of rows where a species is present

forereef17_site_common_species <- names(forereef17_site_species_percentage[forereef17_site_species_percentage > 0.05]) # filter for presence in more than 5% of quads - only 9 species...

forereef17_site_species_percentage
forereef17_site_common_species

#13 species

forereef17_site_common_species_matrix <- forereef17_macro_wide_data_summed_matrix[,colnames(forereef17_macro_wide_data_summed_matrix) %in% forereef17_site_common_species] # filter original species matrix
#okay this has 13

#now need to make it back into a dataframe for the forloop
forereef17_site_common_species_df <- cbind(forereef17_macro_wide_data_summed_meta, forereef17_site_common_species_matrix)

```

will need to do the forloops for each habitat - forereef17
```{r}

#since we have values greater than 1, cannot use arcsin sqrt, need to use Wisconsin transformation

# Initialize a results data frame
results_beta_transformed_forereef17_commonspecies <- data.frame(
  site_habitat = character(),
  median_bray = numeric(),
  median_bray_trans = numeric(),
  mean_bray = numeric(),
  mean_bray_trans = numeric(),
  mean_betadisp = numeric(),
   mean_disp_transformed = numeric(),
  stringsAsFactors = FALSE
)

betadisp_list_forereef17_commonspecies <- list()
betadisp_list_transformed_forereef17_commonspecies <- list()

# Get unique site_habitat combinations
sites_FO17 <- unique(forereef17_site_common_species_df$site_habitat)

for (sh in sites_FO17) {
  # Filter for one site/habitat
  df_sub <- forereef17_site_common_species_df %>%
    filter(site_habitat == sh) %>%
    filter(rowSums(.[, 7:19]) != 0)  # Remove rows with all zeros

  # Community and metadata
  sp_community <- as.matrix(df_sub[, 7:19])
  meta <- df_sub[, 1:6]

  # Arc-sine square root transformation for the community matrix (required before vegdist)
  sp_community_transformed <- wisconsin(sp_community)

  # Confirm it's working
  print(paste("Range after transform for", sh, ":", paste(range(sp_community_transformed, na.rm = TRUE), collapse = " - ")))

  # Bray-Curtis distances
   bray_raw <- vegdist(sp_community, method = "bray")
  bray_transformed <- vegdist(sp_community_transformed, method = "bray")

  correlation <- cor(as.numeric(bray_raw), as.numeric(bray_transformed))
  print(paste("Correlation between raw and transformed Bray-Curtis for", sh, ":", round(correlation, 4)))

  # Calculate median and mean Bray-Curtis distance
  median_bray <- median(bray_raw)
  median_bray_trans <- median(bray_transformed)
  mean_bray <- mean(bray_raw)
  mean_bray_trans <- mean(bray_transformed)

  # Beta dispersion
  betadisp <- betadisper(bray_raw, meta$site_habitat)
  betadisp_transformed <- betadisper(bray_transformed, meta$site_habitat)
  mean_disp <- mean(betadisp$distances)
  mean_disp_transformed <- mean(betadisp_transformed$distances)

  betadisp_list_forereef17_commonspecies[[sh]] <- betadisp
  betadisp_list_transformed_forereef17_commonspecies[[sh]] <- betadisp_transformed

  # Store results
  results_beta_transformed_forereef17_commonspecies <- bind_rows(results_beta_transformed_forereef17_commonspecies, data.frame(
    site_habitat = sh,
    median_bray = median_bray,
    median_bray_trans = median_bray_trans,
    mean_bray = mean_bray,
    mean_bray_trans = mean_bray_trans,
    mean_betadisp = mean_disp,
    mean_disp_transformed = mean_disp_transformed,
    stringsAsFactors = FALSE
  ))
}

# View results
print(results_beta_transformed_forereef17_commonspecies)
betadisp_list_transformed_forereef17_commonspecies

```
# add all habitats together
```{r}
commonspecies_allhabs <- 
  bind_rows(results_beta_transformed_fringing_commonspecies, 
            results_beta_transformed_backreef_commonspecies,
            results_beta_transformed_forereef10_commonspecies,
            results_beta_transformed_forereef17_commonspecies)

# yay, 24 rows!

# now combine with spatial synchrony data

beta_spatialsync_commonspecies <- merge(spatial_synchrony, commonspecies_allhabs, by = "site_habitat")
```
Plotting time
```{r }
# some exploratory plots
# not transformed, all species
beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#transformed, all species
beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

## not transformed, common species
beta_spatialsync_commonspecies %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

# transformed, common species
beta_spatialsync_commonspecies %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony, color = site)) +
  geom_point() +
  #scale_colour_manual(values = habitat_colours) +
  #geom_line(method = lm) +
  facet_wrap(~habitat)

#comparison between removing rare species and not removing them look very similar. looking below to make sure they not exactly the same, which would indicate I did something wrong

beta_spatialsync_transformed %>%
  select(site_habitat, site, mean_bray, mean_bray_trans) %>%
  filter(site == "lter_6")

beta_spatialsync_commonspecies %>%
   select(site_habitat, site, mean_bray, mean_bray_trans) %>%
  filter(site == "lter_6")

#nope, just look different but similar! good sign?

```

#Again, I think this is the figure we want to show. 
```{r}

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()

beta_spatialsync_transformed %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()

beta_spatialsync_commonspecies %>%
  ggplot(aes(x = mean_bray, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()

beta_spatialsync_commonspecies %>%
  ggplot(aes(x = mean_bray_trans, y = spatial_synchrony)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  #scale_colour_manual(values = habitat_colours) +
  theme_minimal()
```
# looking for an R2, just for kicks
```{r}
model_mean_bray_allspecies <- lm(spatial_synchrony ~ mean_bray, data = beta_spatialsync_transformed)
summary(model_mean_bray_allspecies)$r.squared
#.358

model_mean_bray_trans_allspecies <- lm(spatial_synchrony ~ mean_bray_trans, data = beta_spatialsync_transformed)
summary(model_mean_bray_trans_allspecies)$r.squared
#.28

model_mean_bray_common <- lm(spatial_synchrony ~ mean_bray, data = beta_spatialsync_commonspecies)
summary(model_mean_bray_common)$r.squared
#.354

model_mean_bray_trans_common <- lm(spatial_synchrony ~ mean_bray_trans, data = beta_spatialsync_commonspecies)
summary(model_mean_bray_trans_common)$r.squared
#.270
```