### 2025-09-03 supplemental_abundance_figures.Rmd
supplemental abundance figures for Figure S1

---
title: "supplemental abundance figures"
author: "L. Enright"
date: "2025-06-10"
output: html_document
---

# here we read in the data produced from other scripts. should rerun other scripts and make sure these are the most up to date csvs

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(codyn)

macro_long_data <- read.csv(here::here("data_le", "macro_long_data.csv"), stringsAsFactors = F) 
diversity_stability_synchrony <- read.csv(here::here("data_le", "full_plot_level_dss.csv"), stringsAsFactors = F) 
diversity_stability_synchrony_site <- read.csv(here::here("data_le", "diversity_stability_synchrony_site.csv"), stringsAsFactors = F) 
spatial_synchrony <- read.csv(here::here("data_le", "spatial_synchrony.csv"), stringsAsFactors = F) 
mean_plot_level_dss <- read.csv(here::here("data_le", "mean_plot_level_dss.csv"), stringsAsFactors = F)

taxo <- read.csv(here::here("data_le", "MCR_Algal_Functional_Groups_02042025_updatedtax.csv"), stringsAsFactors = F)
```

```{r set colors and labels}

colors_abundance <- c(
    "sargassum_pacificum" = "#543005",
    #"turbinaria_ornata" =  "#7B3F00",
    "turbinaria_ornata" = "#8C510A",
   # "lobophora_variegata" = "#8C510A",
     "lobophora_variegata" = "#BF812D",
    #"dictyota_bartayresiana" =  "#BF812D",
    "dictyota_bartayresiana" = "#DFC27D",
    "peyssonnelia_inamoena" = "#E6B88A",
    "peyssonnelia_sp" = "#F6E8C3",
    "halimeda_sp" = "#C7EAE5",
    "halimeda_minima" = "#80CDC1" ,
     "halimeda_opuntia" = "#35978F",
    "amansia_rhodantha" = "#01665E", 
    "asparagopsis_taxiformis" = "#003C30",
    "Other" = "grey70")

#other colors I considered
"#5AA6D1"
"#7BAFD4"
"#2B83BA"
"#F1D6A7"
"#D2C48D"
"#E6B88A"
"#E9A972"
"#704214"
"#8B4513"
"#6C3A08"
"#75420B"
"#64330A"

# Use a named LIST of expressions for italics
labs_fill <- list(
  "Other" = "Other",
  "sargassum_pacificum" = expression(italic("Sargassum pacificum")),
  "turbinaria_ornata"   = expression(italic("Turbinaria ornata")),
  "lobophora_variegata" = expression(italic("Lobophora variegata")),
  "dictyota_bartayresiana" = expression(italic("Dictyota bartayresiana")),
  "peyssonnelia_inamoena"  = expression(italic("Peyssonnelia inamoena")),
  "peyssonnelia_sp" = expression(italic("Peyssonnelia") ~ "sp."),
 # "peyssonnelia_sp"        = expression(italic("Peyssonnelia sp.")),
    "halimeda_sp" = expression(italic("Halimeda") ~ "sp."),
  "halimeda_minima"        = expression(italic("Halimeda minima")),
  "halimeda_opuntia"       = expression(italic("Halimeda opuntia")),
  "amansia_rhodantha"      = expression(italic("Amansia rhodantha")),
  "asparagopsis_taxiformis"= expression(italic("Asparagopsis taxiformis"))
)

```

# Clock Circle Absolute Abundance Figure 
```{r}

#pull out top 10 abundance, otherwise is messy with all ofthem 
aggdat_hab <- aggregate(prop_cover ~ taxa * year * habitat,  data = subset(macro_long_data,
      taxa == "sargassum_pacificum" |
      taxa == "turbinaria_ornata" | 
      taxa == "dictyota_bartayresiana"| 
      taxa == "peyssonnelia_inamoena"| 
      taxa == "peyssonnelia_sp"| 
      taxa == "halimeda_minima"| 
      taxa == "halimeda_sp"| 
      taxa == "lobophora_variegata"| 
      taxa == "halimeda_opuntia"| 
      taxa == "amansia_rhodantha"| 
      taxa == "asparagopsis_taxiformis"), FUN = mean)

# order the habitats so figure is correct
aggdat_hab$habitat <- factor(aggdat_hab$habitat, levels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

aggdat_hab$habitat

aggdat_hab_supp <- ggplot(aggdat_hab, aes(year, prop_cover, color = taxa)) + 
  # plot species lines
  geom_line(size = 3) + # 3 or 4 works well. 
  # faceted by species
  facet_wrap(~habitat) +
  # on polor coordinates
  coord_polar()  +
  scale_color_manual(values = colors_abundance, labels = labs_fill, drop = FALSE) +
  theme_minimal() +
  labs(x = "", y = "proportional cover", color = "Taxa") +
   theme(
    strip.text   = element_text(size = 25, face = "bold"),   # facet labels
    axis.title   = element_text(size = 25),                  # axis titles
    axis.text    = element_text(size = 20),                  # axis tick labels
    legend.title = element_text(size = 25, face = "bold"),   # legend title
    legend.text  = element_text(size = 20)                   # legend items
  )

#ggsave(filename = "output/supplement/abundanceclock_supplement_09032025v3.jpg", aggdat_hab_supp, height = 15, width = 20)

```
#Relative Abundance Bar Graph

```{r}
macro_long_data

TAXA_sum_TAXA_YR <- macro_long_data %>%
  group_by(habitat, taxa, year) %>%
   summarise(sum_TAXA = sum(prop_cover))

MA_sum_TAXA_YR <- macro_long_data %>%
  group_by(habitat, year) %>%
   summarise(sum_ALLMA = sum(prop_cover))

relative_abundance_TAXA_YR <- merge(TAXA_sum_TAXA_YR, MA_sum_TAXA_YR, by = c("habitat", "year"))

relative_abundance_cover_TAXA_YR <- relative_abundance_TAXA_YR %>%
  #lter 5 fringing does not have any MA in 2018 or 2023
  mutate(relative_abundance = ifelse(sum_ALLMA > 0, sum_TAXA/sum_ALLMA, 0))

highlight_species <- c("sargassum_pacificum", "turbinaria_ornata", "lobophora_variegata", 
                       "dictyota_bartayresiana", "peyssonnelia_inamoena", "peyssonnelia_sp",
                       "halimeda_minima", "halimeda_opuntia", "halimeda_sp",
                       "amansia_rhodantha", "asparagopsis_taxiformis" 
                        )

data_RA <- relative_abundance_cover_TAXA_YR %>%
  mutate(color_group = ifelse(taxa %in% highlight_species, taxa, "Other")) %>%
  group_by(year, habitat, color_group) %>%  # Group by year, habitat, and color category
  summarize(relative_abundance = sum(relative_abundance), .groups = "drop") %>%  # Sum abundance for "Other"
  mutate(color_group = factor(color_group, levels = c("Other", highlight_species)))

# order the habitats so figure is correct
data_RA$habitat <- factor(data_RA$habitat, levels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

#figure option 1

relative_abundance_supp <- data_RA %>%
ggplot(aes(x = year, y = relative_abundance, fill = color_group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = colors_abundance, labels = labs_fill, drop = FALSE) +
  labs(
    x = "Year",
    y = "Relative Abundance",
    fill = "Taxa"
  ) +
  scale_y_continuous(labels = scales::percent) +  
  theme_classic() +
  facet_wrap(~habitat) +
  theme(
    strip.text   = element_text(size = 25, face = "bold"),   # facet labels
    axis.title   = element_text(size = 25),                  # axis titles
    axis.text    = element_text(size = 20),                  # axis tick labels
    legend.title = element_text(size = 25, face = "bold"),   # legend title
    legend.text  = element_text(size = 18)                   # legend items
  )

#ggsave(filename = "output/supplement/relative_abundance_supplement_09032025.jpg", relative_abundance_supp , height = 15, width = 20)
```

#Absolute Abundance Bar Graph

```{r}

#11 highlight species

#average cover in each habitat/year

TAXA_mean_TAXA_YR <- macro_long_data %>%
  group_by(habitat, taxa, year) %>%
   summarise(mean_TAXA = mean(prop_cover))

#add together averages for "other species", avoids the little gray lines
TAXA_mean_TAXA_YR_plot <- TAXA_mean_TAXA_YR %>%
  mutate(color_group = if_else(taxa %in% highlight_species, taxa, "Other")) %>%
  group_by(year, habitat, color_group) %>%
  summarise(mean_TAXA_plot = sum(mean_TAXA, na.rm = TRUE), .groups = "drop") %>%
  mutate(color_group = factor(color_group, levels = c("Other", highlight_species)))

# order the habitats so figure is correct
TAXA_mean_TAXA_YR_plot$habitat <- factor(TAXA_mean_TAXA_YR_plot$habitat, levels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

absolute_abundance_supp <- TAXA_mean_TAXA_YR_plot %>%
ggplot(aes(x = year, y = mean_TAXA_plot, fill = color_group)) +
  geom_bar(stat = "identity") +
    scale_fill_manual(values = colors_abundance, labels = labs_fill, drop = FALSE) +
  labs(
    x = "Site",
    y = "Average Percent Cover",
    fill = "Taxa"
  ) +
  scale_y_continuous(labels = scales::percent) +  
  theme_classic() +
  facet_wrap(~habitat) +
  theme(
    strip.text   = element_text(size = 25, face = "bold"),   # facet labels
    axis.title   = element_text(size = 25),                  # axis titles
    axis.text    = element_text(size = 20),                  # axis tick labels
    legend.title = element_text(size = 25, face = "bold"),   # legend title
    legend.text  = element_text(size = 18)                   # legend items
  )

#ggsave(filename = "output/supplement/absolute_abundance_supplement_09032025.jpg", relative_abundance_supp , height = 15, width = 20)

```